### 2023-01-14 README.Rmd
placeholder

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gran

<!-- badges: start -->
<!-- badges: end -->

The goal of gran is to ...

## Installation

You can install the development version of gran like so:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(gran)
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

### 2023-01-14 R/resolve.R

@@ -1,0 +1,6 @@
+
+## one hr
+
+.search <- memoise::memoise(pkgsearch::cran_package_history, cache = cachem::cache_mem(max_age = 60 * 60))
+x <- .search("rtoot")
+

### 2023-01-14 tests/testthat.R

@@ -1,0 +1,12 @@
+# This file is part of the standard setup for testthat.
+# It is recommended that you do not modify it.
+#
+# Where should you do additional test configuration?
+# Learn more about the roles of various files in:
+# * https://r-pkgs.org/tests.html
+# * https://testthat.r-lib.org/reference/test_package.html#special-files
+
+library(testthat)
+library(gran)
+
+test_check("gran")

### 2023-01-14 README.Rmd
make `resolve` disable Enhances by default fix #3

@@
 
-The goal of gran is to ...
+The goal of gran (gran: R anachronism nullifier) [^gesis] is to obtain the dependency graph of an R package at a specific time point.
 
@@
 ``` r
-# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
+remotes::install_github("GESIS-Methods-Hub/gran")
 ```
@@
 
-This is a basic example which shows you how to solve a common problem:
+To obtain the dependency graph of an R package, use `resolve`
 
@@
 library(gran)
-## basic example code
+rtoot_graph <- resolve(pkg = "rtoot", snapshot_date = "2022-12-31")
+rtoot_graph
 ```
 
-What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:
+---
 
-```{r cars}
-summary(cars)
-```
-
-You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date.
-`devtools::build_readme()` is handy for this.
-You could also use GitHub Actions to re-render `README.Rmd` every time you push.
-An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.
-
-You can also embed plots, for example:
-
-```{r pressure, echo = FALSE}
-plot(pressure)
-```
-
-In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
+[^gesis]: Actually, it stands for "GESIS R Archive Network".

### 2023-01-14 R/resolve.R

@@ -43,15 +43,20 @@
 
 ## We should consider Imports, Depends, LinkingTo, and Enhances
 
-.keep_queryable_dependencies <- function(dep_df) {
+.keep_queryable_dependencies <- function(dep_df, no_enhances = FALSE) {
     if (!"y" %in% colnames(dep_df)) {
         return(NULL)
     }
-    res <- dep_df[dep_df$type != "Suggests" & dep_df$y != "R" & !(dep_df$y %in% c("datasets", "utils", "grDevices", "graphics", "stats", "methods", "tools", "grid", "splines", "Rgraphviz", "parallel", "stats4")),]
+    if (isTRUE(no_enhances)) {
+        disabled_types <- c("Suggests", "Enhances")
+    } else {
+        disabled_types <- c("Suggests")
+    }
+    res <- dep_df[!dep_df$type %in% disabled_types & dep_df$y != "R" & !(dep_df$y %in% c("datasets", "utils", "grDevices", "graphics", "stats", "methods", "tools", "grid", "splines", "Rgraphviz", "parallel", "stats4")),]
     if (nrow(res) == 0) {
         return(NULL)
     } else {
-        return(res$y)
+        return(unique(res$y))
     }
 }
 
@@ -63,7 +68,7 @@
 ## snapshot_date <- "2022-12-10"
 
 #' @export
-resolve <- function(pkg, snapshot_date) {
+resolve <- function(pkg, snapshot_date, no_enhances = TRUE) {
     snapshot_date <- anytime::anytime(snapshot_date, tz = "UTC", asUTC = TRUE)
     if (snapshot_date >= anytime::anytime(Sys.Date())) {
         stop("We don't know the future.", call. = FALSE)
@@ -75,7 +80,7 @@ resolve <- function(pkg, snapshot_date) {
     output[['original']] <- pkg_dep_df
     output$dep <- list()
     q <- dequer::deque()
-    for (dep in .keep_queryable_dependencies(pkg_dep_df)) {
+    for (dep in .keep_queryable_dependencies(pkg_dep_df, no_enhances)) {
         dequer::pushback(q, dep)
     }
     seen_deps <- c()
@@ -83,7 +88,7 @@ resolve <- function(pkg, snapshot_date) {
         current_pkg <- dequer::pop(q)
         pkg_dep_df <- .get_snapshot_dependencies(pkg = current_pkg, snapshot_date = snapshot_date)
         output$dep[[current_pkg]] <- pkg_dep_df
[…]

### 2023-01-15 R/resolve.R
allow exporting gran as edgelist

@@ -18,7 +18,7 @@
     if (nrow(snapshot_versions) == 0) {
         stop("No snapshot version exists for ", pkg, ".",  call. = FALSE)
     }
-    latest_version <- tail(snapshot_versions[order(snapshot_versions$pubdate),], n = 1)
+    latest_version <- utils::tail(snapshot_versions[order(snapshot_versions$pubdate),], n = 1)
     dependencies <- latest_version$dependencies[[1]]
     if (nrow(dependencies != 0)) {
         return(data.frame(snapshot_date = snapshot_date, x = pkg, x_version = latest_version$Version, x_pubdate = latest_version$pubdate,  y = dependencies$package, type = dependencies$type, y_raw_version = dependencies$version))
@@ -110,7 +110,8 @@ print.gran <- function(x, ...) {
     cat("GRAN: The latest version of `", x$pkg, "` at ", as.character(x$snapshot_date), " was ", latest_version, ", which has ", total_deps, " unique dependencies (", total_terminal_nodes, " with no dependencies.)\n", sep = "")
 }
 
-as.edgelist <- function(x, ...) {
+#' @export
+convert_edgelist <- function(x) {
     output <- data.frame(x = x$pkg, y = .keep_queryable_dependencies(x$original))
     for (dep in x$dep) {
         if (!.is_terminal_node(dep)) {

### 2023-01-15 R/resolve.R
use consistent naming `unresolved_deps` #7

@@ -86,7 +86,7 @@ resolve <- function(pkg, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, 
     output$snapshot_date <- snapshot_date
     output[['original']] <- pkg_dep_df
     output$deps <- list()
-    output$unresolveddep <- character(0)
+    output$unresolved_deps <- character(0)
     q <- dequer::deque()
     for (dep in .keep_queryable_dependencies(pkg_dep_df, no_enhances, no_suggests)) {
         dequer::pushback(q, dep)
@@ -107,7 +107,7 @@ resolve <- function(pkg, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, 
             }
         }, error = function(e) {
             ## can't query
-            output$unresolveddep <- c(output$unresolveddep, current_pkg)
+            output$unresolved_deps <- c(output$unresolved_deps, current_pkg)
         })
     }
     attr(output, "class") <- "gran"

### 2023-01-15 README.Rmd
make `resolve` default to resolve multiple pkgs fix #7
* implement granlist S3

@@
 
-The goal of gran (gran: R anachronism nullifier) [^gesis] is to obtain the dependency graph of an R package at a specific time point.
+The goal of gran (gran: R anachronism nullifier) [^gesis] is to obtain the dependency graph of R packages at a specific time point.
 
@@
 library(gran)
-rtoot_graph <- resolve(pkg = "rtoot", snapshot_date = "2022-12-31")
-rtoot_graph
+graph <- resolve(pkgs = c("rtoot", "rtweet"), snapshot_date = "2022-12-31")
+graph
 ```

### 2023-01-15 R/resolve.R

@@ -73,11 +73,43 @@ NULL
 ## snapshot_date <- "2022-12-10"
 
 #' @export
-resolve <- function(pkg, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, verbose = FALSE) {
+resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, verbose = FALSE) {
+    if (missing(snapshot_date)) {
+        if (isTRUE(verbose)) {
+            cat("No `snapshot_date`: Assuming `snapshot_date` to be a week ago.\n")
+        }
+        snapshot_date <- Sys.Date() - 1
+    }
     snapshot_date <- anytime::anytime(snapshot_date, tz = "UTC", asUTC = TRUE)
     if (snapshot_date >= anytime::anytime(Sys.Date())) {
         stop("We don't know the future.", call. = FALSE)
     }
+    output <- list()
+    output$call <- match.call()
+    output$grans <- list()
+    output$snapshot_date <- snapshot_date
+    output$no_enhances <- no_enhances
+    output$no_suggests <- no_suggests
+    output$unresolved_pkgs <- character(0)
+    output$deps_sysreqs <- list() ##TBI
+    output$r_version <- character(0) ## TBI
+    for (pkg in pkgs) {
+        tryCatch({
+            res <- .resolve_pkg(pkg = pkg, snapshot_date = snapshot_date, no_enhances = no_enhances,
+                                no_suggests = no_suggests, verbose = verbose)
+            output$grans[[pkg]] <- res
+        }, error = function(e) {
+            if (isTRUE(verbose)) {
+                cat("Query failed: ", pkg, "\n")
+                output$unresolved_pkgs <- c(output$unresolved_pkgs, pkg)
+            }
+        })
+    }
+    attr(output, "class") <- "granlist"    
+    return(output)
+}
+
+.resolve_pkg <- function(pkg, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, verbose = FALSE) {
     pkg_dep_df <- .get_snapshot_dependencies(pkg = pkg, snapshot_date = snapshot_date)
     output <- list()
     output$pkg <- pkg
@@ -124,6 +156,18 @@ print.gran <- function(x, ...) {
     }
     latest_version <- unique(x$original$x_version)
[…]

### 2023-01-16 R/resolve.R
use remotes instead

@@ -6,18 +6,18 @@ NULL
 
 .search <- memoise::memoise(pkgsearch::cran_package_history, cache = cachem::cache_mem(max_age = 60 * 60))
 
-.raw_sysreqs <- function(pkg) {
-    suppressWarnings(jsonlite::fromJSON(readLines(paste0("https://sysreqs.r-hub.io/pkg/", pkg)), simplifyVector = FALSE))
-}
+## .raw_sysreqs <- function(pkg) {
+##     suppressWarnings(jsonlite::fromJSON(readLines(paste0("https://sysreqs.r-hub.io/pkg/", pkg)), simplifyVector = FALSE))
+## }
 
-.msysreps <- memoise::memoise(.raw_sysreqs, cache = cachem::cache_mem(max_age = 60 * 60))
+## .msysreps <- memoise::memoise(.raw_sysreqs, cache = cachem::cache_mem(max_age = 60 * 60))
 
-.sysreps <- function(pkg, verbose = FALSE) {
-    if (isTRUE(verbose)) {
-        cat("Querying SystemRequirements of", pkg, "\n")
-    }
-    .msysreps(pkg)
-}
+## .sysreps <- function(pkg, verbose = FALSE) {
+##     if (isTRUE(verbose)) {
+##         cat("Querying SystemRequirements of", pkg, "\n")
+##     }
+##     .msysreps(pkg)
+## }
 
 ## get the latest version as of date
 ## let's call this output dep_df; basically is a rough version of edgelist
@@ -86,7 +86,7 @@ NULL
 ## snapshot_date <- "2022-12-10"
 
 #' @export
-resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, verbose = FALSE) {
+resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, get_sysreqs = TRUE, verbose = FALSE) {
     if (missing(snapshot_date)) {
         if (isTRUE(verbose)) {
             cat("No `snapshot_date`: Assuming `snapshot_date` to be a week ago.\n")
@@ -104,7 +104,7 @@ resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE,
     output$no_enhances <- no_enhances
     output$no_suggests <- no_suggests
     output$unresolved_pkgs <- character(0)
-    output$deps_sysreqs <- list() ##TBI
+    output$deps_sysreqs <- list()
     output$r_version <- character(0) ## TBI
[…]

### 2023-01-26 R/resolve.R
query R version fix #6

@@ -6,9 +6,11 @@ NULL
 
 .search <- memoise::memoise(pkgsearch::cran_package_history, cache = cachem::cache_mem(max_age = 60 * 60))
 
-## .raw_sysreqs <- function(pkg) {
-##     suppressWarnings(jsonlite::fromJSON(readLines(paste0("https://sysreqs.r-hub.io/pkg/", pkg)), simplifyVector = FALSE))
-## }
+.get_rver <- function(snapshot_date) {
+    allvers <- suppressWarnings(jsonlite::fromJSON(readLines("https://api.r-hub.io/rversions/r-versions"), simplifyVector = TRUE))
+    allvers$date <- anytime::anytime(allvers$date, tz = "UTC", asUTC = TRUE)
+    tail(allvers[allvers$date < snapshot_date,], 1)$version
+}
 
 ## .msysreps <- memoise::memoise(.raw_sysreqs, cache = cachem::cache_mem(max_age = 60 * 60))
 
@@ -105,7 +107,7 @@ resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE,
     output$no_suggests <- no_suggests
     output$unresolved_pkgs <- character(0)
     output$deps_sysreqs <- list()
-    output$r_version <- character(0) ## TBI
+    output$r_version <- .get_rver(snapshot_date)
     for (pkg in pkgs) {
         tryCatch({
             res <- .resolve_pkg(pkg = pkg, snapshot_date = snapshot_date, no_enhances = no_enhances,

### 2023-01-26 README.Rmd
use remotes instead
Fix #2

@@
 
-```{r example}
+```r
 library(gran)
-graph <- resolve(pkgs = c("rtoot", "rtweet"), snapshot_date = "2022-12-31")
+graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"), snapshot_date = "2020-01-16")
 graph
+```
+
+```{r, include = FALSE}
+graph <- readRDS("tests/testdata/graph.RDS")
+```
+
+```{r example2}
+graph$deps_sysreqs
+```
+
+```{r example 3}
+graph$r_version
+```
+
+## Recreate the computing environment
+
+The object can be used to recreate the computing environment via [Rocker](https://github.com/rocker-org/rocker).
+
+```r
+make_docker(graph, "~/rocker_test")
+```
+
+Now, you can build and run the Docker container.
+
+```bash
+cd rocker_test
+sudo docker build -t gran .
+sudo docker run --rm --name "grantest" -ti gran
+```
+
+Using the above example, `sessionInfo()` outputs the following. You have successfully gone back to the pre-pandemic time.
+
+```
+R version 3.6.2 (2019-12-12)
+Platform: x86_64-pc-linux-gnu (64-bit)
+Running under: Debian GNU/Linux 10 (buster)
+
+Matrix products: default
+BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.3.5.so
+
+locale:
+ [1] LC_CTYPE=en_US.UTF-8          LC_NUMERIC=C                 
+ [3] LC_TIME=en_US.UTF-8           LC_COLLATE=en_US.UTF-8       
+ [5] LC_MONETARY=en_US.UTF-8       LC_MESSAGES=C                
+ [7] LC_PAPER=en_US.UTF-8          LC_NAME=en_US.UTF-8          
+ [9] LC_ADDRESS=en_US.UTF-8        LC_TELEPHONE=en_US.UTF-8     
+[11] LC_MEASUREMENT=en_US.UTF-8    LC_IDENTIFICATION=en_US.UTF-8
+
+attached base packages:
+[1] stats     graphics  grDevices utils     datasets  methods   base
+
+other attached packages:
+[1] topicmodels_0.2-9 LDAvis_0.3.2      openNLP_0.2-7     quanteda_1.5.2
+
+loaded via a namespace (and not attached):
+ [1] NLP_0.2-0           Rcpp_1.0.3          pillar_1.4.3       
+ [4] compiler_3.6.2      tools_3.6.2         stopwords_1.0      
+ [7] lubridate_1.7.4     lifecycle_0.1.0     tibble_2.1.3       
+[10] gtable_0.3.0        lattice_0.20-38     pkgconfig_2.0.3    
+[13] rlang_0.4.2         Matrix_1.2-18       fastmatch_1.1-0    
+[16] parallel_3.6.2      openNLPdata_1.5.3-4 rJava_0.9-11       
+[19] xml2_1.2.2          stringr_1.4.0       stats4_3.6.2       
+[22] grid_3.6.2          data.table_1.12.8   R6_2.4.1           
+[25] ggplot2_3.2.1       spacyr_1.2          magrittr_1.5       
+[28] scales_1.1.0        modeltools_0.2-22   colorspace_1.4-1   
+[31] stringi_1.4.5       RcppParallel_4.4.4  lazyeval_0.2.2     
+[34] munsell_0.5.0       tm_0.7-7            slam_0.1-47        
+[37] crayon_1.3.4    
 ```

### 2023-01-26 R/installation.R

@@ -67,6 +67,29 @@
     if ("default-jdk" %in% debs) {
         cmd <- paste(cmd, "liblzma-dev libpcre3-dev libbz2-dev && R CMD javareconf")
     }
-    cmd
+    paste("apt-get update -qq &&", cmd)
 }
 
+#' @export 
+make_docker <- function(granlist, output_dir) {
+    if (missing(output_dir)) {
+        stop("You must provide `output_dir`.")
+    }
+    install_order <- .determine_installation_order(granlist)
+    sysreps_cmd <- .consolidate_sysreqs(granlist)
+    if (!dir.exists(output_dir)) {
+        dir.create(output_dir)
+    }
+    file.create(file.path(output_dir, "gran.R"))
+    con <- file(file.path(output_dir, "gran.R"), open="w")
+    writeLines(readLines(system.file("header.R", package = "gran")), con = con)
+    cat("install_order <- ", file = con)
+    dput(install_order, file = con)
+    writeLines(readLines(system.file("footer.R", package = "gran")), con = con)
+    close(con)
+    basic_docker <- c("", "", "COPY gran.R ./gran.R", "RUN Rscript gran.R", "CMD [\"R\"]")
+    basic_docker[1] <- paste0("FROM rocker/r-ver:", granlist$r_version)
+    basic_docker[2] <- paste("RUN", sysreps_cmd)
+    writeLines(basic_docker, file.path(output_dir, "Dockerfile"))
+    invisible()
+}

### 2023-01-26 inst/footer.R

@@ -1,0 +1,4 @@
+
+for (i in seq_along(install_order)) {
+    .install_from_cran(install_order[i])
+}

### 2023-01-26 inst/header.R

@@ -1,0 +1,25 @@
+## generated by gran, do not edit by hand
+
+.install_from_cran <- function(x, lib, path = tempdir()) {
+    url <- paste0("https://cran.r-project.org/src/contrib/Archive/", names(x), "/", names(x), "_", x, ".tar.gz")
+    tarball_path <- file.path(path, paste0(names(x), "_", x, ".tar.gz"))
+    tryCatch({
+        suppressWarnings(download.file(url, destfile = tarball_path))
+    }, error = function(e) {
+        ## is the current latest
+        url <- paste0("https://cran.r-project.org/src/contrib/", names(x), "_", x, ".tar.gz")
+        download.file(url, destfile = tarball_path)        
+    })
+    install.packages(pkg = tarball_path, lib = lib, repos = NULL)
+    ## check and error
+    if (!missing(lib)) {
+        installed_packages <- installed.packages(lib.loc = lib)
+    } else {
+        installed_packages <- installed.packages()
+    }
+    if (!names(x) %in% dimnames(installed_packages)[[1]]) {
+        stop("Fail to install ", names(x), "\n")
+    }
+    invisible()
+}
+

### 2023-01-26 README.Rmd
Point the Github repo to a less controversial site
Add bacl the missing tilde
Display the output of `graph` in README

@@
 ``` r
-remotes::install_github("GESIS-Methods-Hub/gran")
+remotes::install_github("chainsawriot/gran")
 ```
@@
 graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"), snapshot_date = "2020-01-16")
-graph
 ```
@@
 ```{r, include = FALSE}
+library(gran)
 graph <- readRDS("tests/testdata/graph.RDS")
+```
+
+```{r example1}
+graph
 ```
@@
 ```bash
-cd rocker_test
+cd ~/rocker_test
 sudo docker build -t gran .

### 2023-01-26 R/installation.R
Change the function name to `dockerize` before it's too late

@@ -71,7 +71,7 @@
 }
 
 #' @export 
-make_docker <- function(granlist, output_dir) {
+dockerize <- function(granlist, output_dir) {
     if (missing(output_dir)) {
         stop("You must provide `output_dir`.")
     }

### 2023-01-26 R/installation.R
Add man

@@ -70,6 +70,18 @@
     paste("apt-get update -qq &&", cmd)
 }
 
+#' Docker The Resolved Result
+#'
+#' This function exports the result from [resolve()] to a Docker file.
+#' @param granlist output from [resolve()]
+#' @param output_dir where to put the Docker file
+#' @return output_dir, invisibly
+#' @examples
+#' \donttest{
+#' graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
+#'                 snapshot_date = "2020-01-16")
+#" dockerize(graph, tempdir())
+#' }
 #' @export 
 dockerize <- function(granlist, output_dir) {
     if (missing(output_dir)) {
@@ -91,5 +103,5 @@ dockerize <- function(granlist, output_dir) {
     basic_docker[1] <- paste0("FROM rocker/r-ver:", granlist$r_version)
     basic_docker[2] <- paste("RUN", sysreps_cmd)
     writeLines(basic_docker, file.path(output_dir, "Dockerfile"))
-    invisible()
+    invisible(output_dir)
 }

### 2023-01-26 R/resolve.R

@@ -87,13 +87,32 @@ NULL
 ## pkg <- "rtoot"
 ## snapshot_date <- "2022-12-10"
 
+#' Resolve Dependencies Of R Packages
+#'
+#' This function recursively queries dependencies of R packages at a specific snapshot time. The dependency graph can then be used to recreate the computing environment. The data on dependencies are provided by R-hub.
+#' 
+#' @param pkgs character vector of R packages to resolve
+#' @param snapshot_date Snapshot date, if not specified, assume to be a month ago
+#' @param no_enhances logical, whether to ignore packages in the "Enhances" field
+#' @param no_suggests logical, whether to ignore packages in the "Suggests" field
+#' @param get_sysreps logical, whether to query for System Requirements
+#' @param os character, which OS to query for system requirements
+#' @param verbose logical, whether to display messages
+#' @return S3 object `granlist`
 #' @export
+#' @seealso [dockerize()]
+#' @examples
+#' \donttest{
+#' graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
+#'                 snapshot_date = "2020-01-16")
+#' graph
+#' }
 resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, get_sysreqs = TRUE, os = "ubuntu-20.04", verbose = FALSE) {
     if (missing(snapshot_date)) {
         if (isTRUE(verbose)) {
-            cat("No `snapshot_date`: Assuming `snapshot_date` to be a week ago.\n")
+            cat("No `snapshot_date`: Assuming `snapshot_date` to be a month ago.\n")
         }
-        snapshot_date <- Sys.Date() - 1
+        snapshot_date <- Sys.Date() - 30
     }
     snapshot_date <- anytime::anytime(snapshot_date, tz = "UTC", asUTC = TRUE)
     if (snapshot_date >= anytime::anytime(Sys.Date())) {

### 2023-01-26 R/installation.R
Fix man

@@ -80,7 +80,7 @@
 #' \donttest{
 #' graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
 #'                 snapshot_date = "2020-01-16")
-#" dockerize(graph, tempdir())
+#' dockerize(graph, ".")
 #' }
 #' @export 
 dockerize <- function(granlist, output_dir) {

### 2023-01-29 README.Rmd
make Readme display better
Change the function name to `dockerize` before it's too late
add logo
Fix #12

@@
 
-# gran
+# gran <img src="man/figures/gran_logo.svg" align="right" height="200" />
 
@@
 library(gran)
-graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"), snapshot_date = "2020-01-16")
+graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
+                 snapshot_date = "2020-01-16")
 ```
@@
 ```r
-make_docker(graph, "~/rocker_test")
+dockerize(graph, "~/rocker_test")
 ```
@@
 
+## Using alternative Rocker images
+
+One can also select other Rocker versioned images: `rstudio`, `tidyverse`, `verse`, `geospatial`.
+
+```r
+dockerize(graph, "~/rocker_test", image = "rstudio")
+```
+
+`tidyverse`, `verse`, and `geospatial` are similar to the default (`r-ver`). For `rstudio`, one needs to build and launch it with:
+
+```bash
+cd ~/rocker_test
+sudo docker build -t gran .
+sudo docker run -p 8787:8787 -e PASSWORD=abc123 --rm --name "grantest" -ti gran
+```
+
+With any browser, go to: `local:8787`. The default username is `rstudio`, password is as specified.
+
+## Acknowledgment
+
+The logo of gran is a remix of [this](https://commons.wikimedia.org/wiki/File:Flag_of_the_Canary_Islands.svg) public domain image.
+
 ---

### 2023-01-29 R/installation.R

@@ -74,6 +74,7 @@
 #' This function exports the result from [resolve()] to a Docker file.
 #' @param granlist output from [resolve()]
 #' @param output_dir where to put the Docker file
+#' @param image character, which versioned Rocker image to use. Can only be "r-ver", "rstudio", "tidyverse", "verse", "geospatial"
 #' @return output_dir, invisibly
 #' @examples
 #' \donttest{
@@ -84,10 +85,11 @@
 #' }
 #' }
 #' @export 
-dockerize <- function(granlist, output_dir) {
+dockerize <- function(granlist, output_dir, image = c("r-ver", "rstudio", "tidyverse", "verse", "geospatial")) {
     if (missing(output_dir)) {
         stop("You must provide `output_dir`.")
     }
+    image <- match.arg(image)
     install_order <- .determine_installation_order(granlist)
     sysreps_cmd <- .consolidate_sysreqs(granlist)
     if (!dir.exists(output_dir)) {
@@ -101,8 +103,12 @@ dockerize <- function(granlist, output_dir) {
     writeLines(readLines(system.file("footer.R", package = "gran")), con = con)
     close(con)
     basic_docker <- c("", "", "COPY gran.R ./gran.R", "RUN Rscript gran.R", "CMD [\"R\"]")
-    basic_docker[1] <- paste0("FROM rocker/r-ver:", granlist$r_version)
+    basic_docker[1] <- paste0("FROM rocker/", image, ":", granlist$r_version)
     basic_docker[2] <- paste("RUN", sysreps_cmd)
+    if (image == "rstudio") {
+        basic_docker[5] <- "EXPOSE 8787"
+        basic_docker[6] <- "CMD [\"/init\"]"
+    }
     writeLines(basic_docker, file.path(output_dir, "Dockerfile"))
     invisible(output_dir)
 }

### 2023-01-31 R/installation.R
Add defensive programming fix #10

@@ -89,6 +89,9 @@ dockerize <- function(granlist, output_dir, image = c("r-ver", "rstudio", "tidyv
     if (missing(output_dir)) {
         stop("You must provide `output_dir`.")
     }
+    if (!grepl("^ubuntu", granlist$os)) {
+        stop("System dependencies of ", granlist$os, " can't be dockerized.")
+    }
     image <- match.arg(image)
     install_order <- .determine_installation_order(granlist)
     sysreps_cmd <- .consolidate_sysreqs(granlist)

### 2023-01-31 R/resolve.R

@@ -110,6 +110,9 @@ NULL
 #' }
 #' }
 resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE, get_sysreqs = TRUE, os = "ubuntu-20.04", verbose = FALSE) {
+    if (!os %in% supported_os) {
+        stop("Don't know how to resolve ", os, ". Supported OSes are: ", paste(supported_os, collapse = ", "))
+    }
     if (missing(snapshot_date)) {
         if (isTRUE(verbose)) {
             cat("No `snapshot_date`: Assuming `snapshot_date` to be a month ago.\n")
@@ -233,3 +236,8 @@ convert_edgelist <- function(x) {
     targets <- .granlist_extract_all_deps(granlist)
     remotes::system_requirements(package = targets, os = os) ## don't gp me
 }
+
+## os <- names(remotes:::supported_os_versions())
+## supported_os <- unlist(mapply(function(x, y) paste(x,"-", y, sep = ""), os, remotes:::supported_os_versions()))
+## names(supported_os) <- NULL
+## usethis::use_data(supported_os, internal = TRUE)

### 2023-01-31 paper/paper.Rmd
Add paper placeholder

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'Gala: A Python package for galactic dynamics'
tags:
  - Python
  - astronomy
  - dynamics
  - galactic dynamics
  - milky way
authors:
  - name: Adrian M. Price-Whelan
    orcid: 0000-0003-0872-7098
    affiliation: "1, 2" # (Multiple affiliations must be quoted)
  - name: Author 2
    orcid: 0000-0000-0000-0000
    affiliation: 2
affiliations:
 - name: Lyman Spitzer, Jr. Fellow, Princeton University
   index: 1
 - name: Institution 2
   index: 2
citation_author: Price-Whelan et. al.
date: 13 August 2017
year: 2017
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

The forces on stars, galaxies, and dark matter under external gravitational
fields lead to the dynamical evolution of structures in the universe. The orbits
of these bodies are therefore key to understanding the formation, history, and
future state of galaxies. The field of "galactic dynamics," which aims to model
the gravitating components of galaxies to study their structure and evolution,
is now well-established, commonly taught, and frequently used in astronomy.
Aside from toy problems and demonstrations, the majority of problems require
efficient numerical tools, many of which require the same base code (e.g., for
performing numerical orbit integration).

``Gala`` is an Astropy-affiliated Python package for galactic dynamics. Python
enables wrapping low-level languages (e.g., C) for speed without losing
flexibility or ease-of-use in the user-interface. The API for ``Gala`` was
designed to provide a class-based and user-friendly interface to fast (C or
Cython-optimized) implementations of common operations such as gravitational
potential and force evaluation, orbit integration, dynamical transformations,
and chaos indicators for nonlinear dynamics. ``Gala`` also relies heavily on and
interfaces well with the implementations of physical units and astronomical
coordinate systems in the ``Astropy`` package [@astropy] (``astropy.units`` and
``astropy.coordinates``).

``Gala`` was designed to be used by both astronomical researchers and by
students in courses on gravitational dynamics or astronomy. It has already been
used in a number of scientific publications [@Pearson:2017] and has also been
used in graduate courses on Galactic dynamics to, e.g., provide interactive
visualizations of textbook material [@Binney:2008]. The combination of speed,
design, and support for Astropy functionality in ``Gala`` will enable exciting
scientific explorations of forthcoming data releases from the *Gaia* mission
[@gaia] by students and experts alike.

# Mathematics

Single dollars ($) are required for inline mathematics e.g. $f(x) = e^{\pi/x}$

Double dollars make self-standing equations:

$$\Theta(x) = \left\{\begin{array}{l}
0\textrm{ if } x < 0\cr
1\textrm{ else}
\end{array}\right.$$

# Citations

Citations to entries in paper.bib should be in
[rMarkdown](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)
format.

For a quick reference, the following citation commands can be used:
- `@author:2001`  ->  "Author et al. (2001)"
- `[@author:2001]` -> "(Author et al., 2001)"
- `[@author1:2001; @author2:2001]` -> "(Author1 et al., 2001; Author2 et al., 2002)"

# Rendered R Figures

Figures can be plotted like so:

```{r}
plot(1:10)
```

# Acknowledgements

We acknowledge contributions from Brigitta Sipocz, Syrtis Major, and Semyeong
Oh, and support from Kathryn Johnston during the genesis of this project.

# References

### 2023-02-02 R/resolve.R
Add tests for S3 printing and fix resolve tests

@@ -208,7 +208,7 @@ print.gran <- function(x, ...) {
 #' @export
 print.granlist <- function(x, all_pkgs = FALSE, ...) {
     n_grans <- length(x$grans)
-    cat("resolved:", n_grans, "package(s). Unresolved package(s): ", length(x$unresolved_pkgs), "\n")
+    cat("resolved:", n_grans, "package(s). Unresolved package(s):", length(x$unresolved_pkgs), "\n")
     if (n_grans > 0) {
         if (n_grans <= 5 || isTRUE(all_pkgs)) {
             print(x$grans)

### 2023-02-02 tests/testthat/test_resolve.R

@@ -8,9 +8,9 @@ test_that("normal", {
     ## normal case
     skip_if_offline()
     skip_on_cran()
-    expect_silent(x2 <- resolve("LDAvis", snapshot_date = "2023-01-01"))
-    expect_equal(x2$unresolved_pkgs, character(0))
-    expect_equal(x2$deps_sysreqs, character(0))
+    expect_silent(x <- resolve("LDAvis", snapshot_date = "2023-01-01"))
+    expect_equal(x$unresolved_pkgs, character(0))
+    expect_equal(x$deps_sysreqs, character(0))
 })
 
 test_that("unresolved", {
@@ -18,7 +18,7 @@ test_that("unresolved", {
     skip_if_offline()
     skip_on_cran()
     expect_error(res <- .get_snapshot_dependencies("LDAvis", snapshot_date = "2001-10-01"))
-    warns <- capture_warnings(res <- resolve("LDAvis", snapshot_date = "2000-10-01"))
+    warns <- capture_warnings(x <- resolve("LDAvis", snapshot_date = "2000-10-01"))
     expect_true(length(warns) == 2)
     expect_true(any(grepl("^Some package", warns)))
     expect_true(any(grepl("^No packages to query", warns)))
@@ -26,11 +26,11 @@ test_that("unresolved", {
     expect_equal(x$grans, list())
     expect_true(is.na(x$deps_sysreqs))
     ## a mixture; sna is old enough
-    warns2 <- capture_warnings(x3 <- resolve(c("sna", "LDAvis"), snapshot_date = "2001-10-01"))
+    warns2 <- capture_warnings(x2 <- resolve(c("sna", "LDAvis"), snapshot_date = "2001-10-01"))
     expect_true(length(warns2) == 1)
     expect_true(any(grepl("^Some package", warns2)))
     expect_false(any(grepl("^No packages to query", warns2)))
-    expect_equal(x3$unresolved_pkgs, "LDAvis")
-    expect_false(length(x3$grans) == 0)
-    expect_equal(x3$deps_sysreqs, character(0))
+    expect_equal(x2$unresolved_pkgs, "LDAvis")
+    expect_false(length(x2$grans) == 0)
+    expect_equal(x2$deps_sysreqs, character(0))
 })

### 2023-02-02 tests/testthat/test_s3.R

@@ -1,0 +1,27 @@
+test_that("printing with and without resolved packages", {
+    gran_mixture <- readRDS("../testdata/gran_mixture.RDS")
+    gran_ok <- readRDS("../testdata/gran_ok.RDS")
+    gran_unresolved <- readRDS("../testdata/gran_unresolved.RDS")
+    gran_output <- capture_output(print(gran_ok$grans[[1]]))
+    expect_equal(gran_output, "The latest version of `LDAvis` at 2023-01-01 was 0.3.2, which has 2 unique dependencies (2 with no dependencies.)")
+    gran_ok_output <- capture_output(print(gran_ok))
+    expect_true(grepl("^resolved: 1 package\\(s\\). Unresolved package\\(s\\): 0", gran_ok_output))
+    gran_mixture_output <- capture_output(print(gran_mixture))
+    expect_true(grepl("^resolved: 1 package\\(s\\). Unresolved package\\(s\\): 1", gran_mixture_output))
+    gran_unresolved_output <- capture_output(print(gran_unresolved))
+    expect_true(grepl("^resolved: 0 package\\(s\\). Unresolved package\\(s\\): 1", gran_unresolved_output))
+    expect_false(grepl("The latest version", gran_unresolved_output))
+})
+
+test_that("all_pkgs", {
+    gran_mixture <- readRDS("../testdata/gran_mixture.RDS")
+    output1 <- capture_output(print(gran_mixture, all_pkgs = FALSE))
+    output2 <- capture_output(print(gran_mixture, all_pkgs = TRUE))
+    expect_true(output1 == output2)
+    gran_6 <- readRDS("../testdata/gran_6.RDS")
+    output1 <- capture_output(print(gran_6, all_pkgs = FALSE))
+    output2 <- capture_output(print(gran_6, all_pkgs = TRUE))
+    expect_false(output1 == output2)
+    expect_true(grepl("First 5 packages", output1))
+    expect_false(grepl("First 5 packages", output2))
+})

### 2023-02-02 README.Rmd
Update README
Add GHA

@@
 <!-- badges: start -->
+[![R-CMD-check](https://github.com/chainsawriot/gran/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/chainsawriot/gran/actions/workflows/R-CMD-check.yaml)
 <!-- badges: end -->
@@
 ```{r, include = FALSE}
-library(gran)
+devtools::load_all()
 graph <- readRDS("tests/testdata/graph.RDS")

### 2023-02-02 paper/paper.rmd
Change basic content [paper] [no ci]

---
title: 'gran: Reconstructing reproducible R computing environments with ease'
tags:
  - R
  - reproducibility
  - docker
authors:
  - name: "Chung-hong Chan"
    orcid: 0000-0002-6232-7530
    affiliation: GESIS Leibniz-Institut für Sozialwissenschaften
abstract: |
  Enter the text of your abstract here.
keywords:
  - blah
  - blee
  - bloo
  - these are optional and can be removed
date: 02 February 2023
year: 2023
bibliography: paper.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

# Statement of need

The ability to reconstruct the computational environment is one of the most important aspects of reproducibility. In the realm of R, the important aspects are what and which version of installed R packages as well as the exact R version.

`renv` [@renvrpkg] (and its derivatives such as `jetpack` and its predecessor `packrat`) takes a similar approach to Python's `virtualenv` and Ruby's `Gem` to pin down the exact version of R packages using a "lock file". `containerit` [@nuest:2019] takes the current state of the computational environment and "containerize" it as a Dockerfile. But `containerit` does not pin down the exact version of R packages. Other solutions such as `groundhog` [@groundhogrpkg] and `checkpoint` [@checkpointrpkg] depend on the availability of The Microsoft R Application Network (MRAN), which will be shut down on July 1st, 2023.

These solutions are better for prospective usage, i.e. using them now to ensure the reproducibility of the current research for future researchers. `gran` targets mostly on retrospective usage, i.e. using `gran` to reconstruct historical R computational environments which have not been completely declared. One can think of `gran` as an archaeological tool. In the following examples, `gran` is used to enable the reproducibility of published literature.

# Basic usage

There are two important functions of `gran`: `resolve()` and `dockerize()`.

`resolve()` queries various web services from the r-hub project of the R Consortium for information about R packages at a specific time point that is necessary for reconstructing a computational environment, e.g. recursive dependencies, system requirements, and R version. For instance, if there was a computational environment constructed on 2020-01-16 (called "snapshot date") with the several natural language processing R packages, `resolve()` can be used to resolve all the dependencies of these R packages.

```r
library(gran)
graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
                 snapshot_date = "2020-01-16")
graph
```

The resolved result is an S3 object called `granlist`. The information contained in a `granlist` can then be used to construct a computational environment in the similar manner as `containerit`, but with the packages and R versions pinned on the snapshot date. Then, `dockerize()` is used to generate the `Dockerfile` and other scripts in the `output_dir`. For R >= 3.1, the images from the Rocker project are used [@boettiger:2017:IR]. For R < 3.1 but >= 2.1, a custom image based on Debian Woody is used. As of writing, `gran` does not support R < 2.1, i.e. snapshot date earlier than 2005-04-19. The development to support R < 2.1 is in progress.

```r
dockerize(granlist graph, output_dir = "docker")
```

There are two features of `dockerize()` that are important for future reproducibility.

1. By default, the container building process downloads source packages from CRAN and then compiles them from source. This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continous availability since 1997-04-23) [^1]. However, it is possible also to cache (or archive) the source packages now. The archived R packages can then be used instead during the building process.

[^1]: https://stat.ethz.ch/pipermail/r-announce/1997/000001.html

```r
dockerize(granlist graph, output_dir = "docker", cache = TRUE)
```

2. It is also possible to install R packages in a separate library during the building process to isolate all these R packages from the main library.

```r
dockerize(granlist graph, output_dir = "docker", cache = TRUE,
          lib = "anotherlibrary")
```

For the sake of completeness, the instructions for building and running the Docker container for Unix-like systems are included here.

```bash
cd docker
## might need to sudo
docker build -t gran .
docker run --rm --name "grantest" -ti gran
```

# Case Studies

## quanteda JOSS paper

The software paper of the text analysis R package `quanteda` was published on 2018-10-06 [@benoit:2018]. In the paper, the following R code snippet is included.

```r
library("quanteda")
# construct the feature co-occurrence matrix
examplefcm <-
tokens(data_corpus_irishbudget2010, remove_punct = TRUE) %>%
tokens_tolower() %>%
tokens_remove(stopwords("english"), padding = FALSE) %>%
fcm(context = "window", window = 5, tri = FALSE)
# choose 30 most frequency features
topfeats <- names(topfeatures(examplefcm, 30))
# select the top 30 features only, plot the network
set.seed(100)
textplot_network(fcm_select(examplefcm, topfeats), min_freq = 0.8)
```

On 2023-02-08, this code snippet is not executable with the current version of `quanteda` (3.2.4). It is possible to install the "period appropriate" version of `quanteda` (1.3.4) using `remotes` on the current version of R (4.2.2). And indeed, the above code snippet can still be executed.

```r
remotes::install_version("quanteda", version = "1.3.4")
```

The issue is that installing `quanteda` 1.3.4 this way installs the latest dependencies from CRAN. `quanteda` 1.3.4 uses a deprecated (but not yet removed) function of `Matrix` (`as(<dgTMatrix>, "dgCMatrix")`). If this function were removed in the future, the above code snippet would not be executable anymore.

Using `gran`, one can query the version of `quanteda` on 2018-10-06 and create a Docker container with all the "period appropriate" dependencies. Here, the `rstudio` Rocker image is selected.

```r
library(gran)
graph <- resolve(pkgs = "quanteda",
                 snapshot_date = "2018-10-06",
                 os = "ubuntu-18.04")
dockerize(graph, output_dir = "quanteda_docker",
          image = "rstudio")
```

The above code snippet can be executed with the generated container without any problem.

```{r, fig.cap = "The code snippet running in a R 3.5.1 container created with gran", echo = FALSE}
knitr::include_graphics("quanteda_rstudio.png", dpi = 300)
```

# References

### 2023-02-05 R/installation.R
fix #18

@@ -69,7 +69,8 @@
     paste("apt-get update -qq &&", cmd)
 }
 
-.write_granlist_as_comment <- function(granlist, con, path) {
+.write_granlist_as_comment <- function(granlist, con, path, verbose, lib,
+                                       cran_mirror, check_cran_mirror) {
     cat("## ## To reconstruct this file, please install version",
         as.character(utils::packageVersion("gran")), "of `gran` and run:\n", file = con)
     cat("## granlist <- \n", file = con)
@@ -77,7 +78,39 @@
     dput(granlist, file = tempgrancontent)
     granlist_src <- readLines(tempgrancontent)
     writeLines(paste("##", granlist_src), con = con)
-    writeLines(paste("## gran::export_granlist(granlist = granlist, path =\"", path, "\")"), con = con)
+    if (is.na(lib)) {
+        lib_as_character <- "NA"
+    } else {
+        lib_as_character <- paste0("\"", lib, "\"")
+    }
+    writeLines(paste0("## gran::export_granlist(granlist = granlist, path = \"", path, "\", verbose = ",
+                     as.character(verbose), ", lib = ", lib_as_character,
+                     ", cran_mirror = \"", cran_mirror, "\", check_cran_mirror = ",
+                     as.character(check_cran_mirror), ")"), con = con)
+}
+
+.check_mirror <- function(mirror) {
+    if (mirror == "https://cran.r-project.org/") {
+        return(TRUE)
+    }
+    all_mirrors <- utils::getCRANmirrors()$URL
+    mirror %in% all_mirrors
+}
+
+.normalize_url <- function(mirror) {
+    if (grepl("^http://", mirror)) {
+        mirror <- gsub("^http://", "https://", mirror)
+    }
+    if (!grepl("^https://", mirror)) {
+        mirror <- paste0("https://", mirror)
+    }
+    if (!grepl("/$", mirror)) {
+        mirror <- paste0(mirror, "/")
+    }
+    if (grepl("/+$", mirror)) {
+        mirror <- gsub("/+$", "/", mirror)
+    }
+    return(mirror)
 }
 
 #' Export The Resolved Result As Installation Script
@@ -90,6 +123,8 @@
 #' @param verbose logical, pass to [install.packages()], the negated value is also passed as `quiet` to both [install.packages()]
 #' and [download.file()].
 #' @param lib character, pass to [install.packages()]. By default, it is NA (to install the packages to the default location)
[…]

### 2023-02-05 tests/testthat/test_dockerize.R

@@ -34,5 +34,23 @@ test_that("integration of #16 in dockerize()", {
     x <- readLines(file.path(temp_dir, "gran.R"))
     expect_true(any(grepl("^lib <- \"abc\"", x)))
     Dockerfile <- readLines(file.path(temp_dir, "Dockerfile"))
-    expect_true(any(grepl("^RUN mkdir", Dockerfile)))    
+    expect_true(any(grepl("^RUN mkdir", Dockerfile)))
+})
+
+test_that("integration of #18 in dockerize()", {
+    gran_ok <- readRDS("../testdata/gran_ok.RDS")
+    temp_dir <- file.path(tempdir(), sample(1:10000, size = 1))
+    dockerize(granlist = gran_ok, output_dir = temp_dir) ## cran_mirror = "https://cran.r-project.org/"
+    x <- readLines(file.path(temp_dir, "gran.R"))
+    expect_true(any(grepl("^cran_mirror <- \"https://cran\\.r\\-project\\.org/\"", x)))
+    dockerize(granlist = gran_ok, output_dir = temp_dir, cran_mirror = "cran.r-project.org")
+    x <- readLines(file.path(temp_dir, "gran.R"))
+    expect_true(any(grepl("^cran_mirror <- \"https://cran\\.r\\-project\\.org/\"", x)))
+    dockerize(granlist = gran_ok, output_dir = temp_dir, cran_mirror = "https://cloud.r-project.org/")
+    x <- readLines(file.path(temp_dir, "gran.R"))
+    expect_true(any(grepl("^cran_mirror <- \"https://cloud\\.r\\-project\\.org/\"", x)))
+    expect_error(dockerize(granlist = gran_ok, output_dir = temp_dir, cran_mirror = "https://www.chainsawriot.com/"))
+    expect_error(dockerize(granlist = gran_ok, output_dir = temp_dir, cran_mirror = "https://www.chainsawriot.com/", check_cran_mirror = FALSE), NA)
+    x <- readLines(file.path(temp_dir, "gran.R"))
+    expect_true(any(grepl("^cran_mirror <- \"https://www\\.chainsawriot\\.com/\"", x)))
 })

### 2023-02-05 R/resolve.R
Change all `computing` to `computational` [no ci]

@@ -89,7 +89,7 @@ NULL
 
 #' Resolve Dependencies Of R Packages
 #'
-#' This function recursively queries dependencies of R packages at a specific snapshot time. The dependency graph can then be used to recreate the computing environment. The data on dependencies are provided by R-hub.
+#' This function recursively queries dependencies of R packages at a specific snapshot time. The dependency graph can then be used to recreate the computational environment. The data on dependencies are provided by R-hub.
 #' 
 #' @param pkgs character vector of R packages to resolve
 #' @param snapshot_date Snapshot date, if not specified, assume to be a month ago

### 2023-02-05 README.Rmd
Change all `computing` to `computational` [no ci]
Update readme to reflect #1 [no ci]

@@
 The goal of gran (gran: R anachronism nullifier) [^gesis] is to obtain the dependency graph of R packages at a specific time point.
+
+Although this package can also be used to ensure the current R computational environment can be reconstructed by future researchers, this package gears towards reconstructing historical R computational environments which have not been completely declared.
+For the former purpose, packages such as [renv](https://github.com/rstudio/renv/), [groundhog](https://github.com/CredibilityLab/groundhog), [miniCRAN](https://github.com/andrie/miniCRAN), and [Require](https://github.com/PredictiveEcology/Require) should be used.
+One can think of gran as an archaeological tool.
+
+To reconstruct a historical R computational environment, this package assumes only the availability of source packages on CRAN.
+The reconstruction procedures have been tested in several vintage versions of R.
 
@@
 
-## Recreate the computing environment
+## Recreate the computational environment via Rocker
 
-The object can be used to recreate the computing environment via [Rocker](https://github.com/rocker-org/rocker).
+The object can be used to recreate the computational environment via [Rocker](https://github.com/rocker-org/rocker). Please note that the oldest R version one can get from Rocker is R 3.1.0.
 
@@
 The logo of gran is a remix of [this](https://commons.wikimedia.org/wiki/File:Flag_of_the_Canary_Islands.svg) public domain image.
+The two dogs should be *Presa Canario*, the native dog breed on the islands of Gran Canaria and Tenerife.
 

### 2023-02-05 README.Rmd
Update `readme` to introduce `export_granlist` [no ci]

@@
 
+The resolved result is an S3 object called `granlist` and can be exported as an installation script. The installation script can be execute on a vanilla R installation.
+
+```r
+export_granlist(graph, "gran.R")
+```
+
+However, the execution of the installation script often fails (now) due to missing system dependencies and incompatible R versions. Therefore, the approach outlined below should be used.
+
 ## Recreate the computational environment via Rocker
 
-The object can be used to recreate the computational environment via [Rocker](https://github.com/rocker-org/rocker). Please note that the oldest R version one can get from Rocker is R 3.1.0.
+A `granlist` object can be used to recreate the computational environment via [Rocker](https://github.com/rocker-org/rocker). Please note that the oldest R version one can get from Rocker is R 3.1.0.
 

### 2023-02-07 README.Rmd
Update README to reflect #17 and #14 [no ci]

@@
 
-## Using alternative Rocker images
+### Caching R packages
+
+One can also cache (or archive) the R packages from CRAN at the time `dockerize` is executed.
+The cached R packages will then transfer to the container.
+Please note that system requirements (i.e.
+`deb` packages) are not cached.
+
+```r
+dockerize(graph, "~/rocker_test", cache = TRUE)
+```
+
+### Using alternative Rocker images
 
@@
 
+## Recreate the computational environment for R < 3.1.0
+
+`gran` can still be used to recreate computational environments for R < 3.1.0.
+The Dockerfile generated is based on Debian Woody (3.0) and the requested version of R is compiled from source.
+As of writing, this method works for R < 3.1.0 but not R < 2.1.0.
+The `image` parameter is ignored in this case.
+
+```r
+gran_rio <- resolve("rio", snapshot_date = "2013-08-28") ## R 3.0.1
+dockerize(gran_rio, output_dir = "~/old_renviron")
+```
+
 ## Acknowledgment

### 2023-02-08 paper/paper.rmd
Add David as an author [no ci]

@@
     orcid: 0000-0002-6232-7530
+    affiliation: GESIS Leibniz-Institut für Sozialwissenschaften
+  - name: "David Schoch"
+    orcid: 0000-0003-2952-4812
     affiliation: GESIS Leibniz-Institut für Sozialwissenschaften    

### 2023-02-12 R/installation.R
rename the functional part of the package to rang
ref #35

@@ -1,5 +1,3 @@
-##granlist <- readRDS("tests/testdata/graph.RDS")
-
 .safe_get_uid <- function(pkgref, uid) {
     if (is.null(uid$get(pkgref))) {
         return(NA_character_)
@@ -8,28 +6,28 @@
     }
 }
 
-.determine_installation_order <- function(granlist) {
+.determine_installation_order <- function(rang) {
     dep <- fastmap::fastmap()
     version <- fastmap::fastmap()
     uid <- fastmap::fastmap()
     ## package name as per DESCRIPTION, aka. x
     ## can't use x here because it is too generic
     pkgname <- fastmap::fastmap()
-    for (gran in granlist$grans) {
-        current_pkgref <- unique(gran$original$x_pkgref)
-        current_ver <- unique(gran$original$x_version)
-        current_dep <- .keep_queryable_dependencies(dep_df = gran$original, no_enhances = gran$no_enhances, no_suggests = gran$no_suggests)
-        current_pkgname <- unique(gran$original$x)
+    for (ranglet in rang$ranglets) {
+        current_pkgref <- unique(ranglet$original$x_pkgref)
+        current_ver <- unique(ranglet$original$x_version)
+        current_dep <- .keep_queryable_dependencies(dep_df = ranglet$original, no_enhances = ranglet$no_enhances, no_suggests = ranglet$no_suggests)
+        current_pkgname <- unique(ranglet$original$x)
         dep$set(current_pkgref, current_dep)
         version$set(current_pkgref, current_ver)
         pkgname$set(current_pkgref, current_pkgname)
-        if ("x_uid" %in% colnames(gran$original)) {
-            uid$set(current_pkgref, unique(gran$original$x_uid))
+        if ("x_uid" %in% colnames(ranglet$original)) {
+            uid$set(current_pkgref, unique(ranglet$original$x_uid))
         }
-        for (dep_df in gran$deps) {
+        for (dep_df in ranglet$deps) {
             current_pkgref <- unique(dep_df$x_pkgref)
             current_ver <- unique(dep_df$x_version)
-            current_dep <- .keep_queryable_dependencies(dep_df = dep_df, no_enhances = gran$no_enhances, no_suggests = gran$no_suggests)
+            current_dep <- .keep_queryable_dependencies(dep_df = dep_df, no_enhances = ranglet$no_enhances, no_suggests = ranglet$no_suggests)
             current_pkgname <- unique(dep_df$x)
[…]

### 2023-02-12 R/resolve.R

@@ -182,7 +182,7 @@ NULL
 #' packages are assumed to have no system requirement.
 #' @param os character, which OS to query for system requirements
 #' @param verbose logical, whether to display messages
-#' @return S3 object `granlist`
+#' @return S3 object `rang`
 #' @export
 #' @seealso [dockerize()]
 #' @references
@@ -216,7 +216,7 @@ resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE,
     pkgrefs <- .normalize_pkgs(pkgs)
     output <- list()
     output$call <- match.call()
-    output$grans <- list()
+    output$ranglets <- list()
     output$snapshot_date <- snapshot_date
     output$no_enhances <- no_enhances
     output$no_suggests <- no_suggests
@@ -228,23 +228,23 @@ resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE,
         tryCatch({
             res <- .resolve_pkgref(pkgref = pkgref, snapshot_date = snapshot_date, no_enhances = no_enhances,
                                 no_suggests = no_suggests, verbose = verbose)
-            output$grans[[pkgref]] <- res
+            output$ranglets[[pkgref]] <- res
         }, error = function(err) {
             if (isTRUE(verbose)) {
                 cat("Query failed: ", pkgref, "\n")
             }
         })
     }
-    output$unresolved_pkgrefs <- setdiff(pkgrefs, names(output$grans))
+    output$unresolved_pkgrefs <- setdiff(pkgrefs, names(output$ranglets))
     if (length(output$unresolved_pkgrefs) > 0) {
         warning("Some package(s) can't be resolved: ", paste(output$unresolved_pkgrefs, collapse = ", "), call. = FALSE)
     }
     if (isTRUE(get_sysreqs)) {
-        res <- .granlist_query_sysreqs(output, os = os)
+        res <- .rang_query_sysreqs(output, os = os)
         output$deps_sysreqs <- res
         .has_ppa_in_sysreqs(output)
     }
-    attr(output, "class") <- "granlist"
+    attr(output, "class") <- "rang"
     return(output)
 }
 
@@ -281,12 +281,12 @@ resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE,
[…]

### 2023-02-12 tests/testthat/test_dockerize.R

@@ -8,105 +8,105 @@ test_that("defensive programming", {
 })
 
 test_that("integration of #13 in dockerize()", {
-    gran_ok <- readRDS("../testdata/gran_ok.RDS")
+    rang_ok <- readRDS("../testdata/rang_ok.RDS")
     temp_dir <- .gen_temp_dir()
-    dockerize(granlist = gran_ok, output_dir = temp_dir) ## granlist_as_comment = TRUE
-    x <- readLines(file.path(temp_dir, "gran.R"))
+    dockerize(rang = rang_ok, output_dir = temp_dir) ## rang_as_comment = TRUE
+    x <- readLines(file.path(temp_dir, "rang.R"))
     expect_true(any(grepl("^## ## To reconstruct this file", x)))
-    dockerize(granlist = gran_ok, output_dir = temp_dir, granlist_as_comment = FALSE)
-    x <- readLines(file.path(temp_dir, "gran.R"))
+    dockerize(rang = rang_ok, output_dir = temp_dir, rang_as_comment = FALSE)
+    x <- readLines(file.path(temp_dir, "rang.R"))
     expect_false(any(grepl("^## ## To reconstruct this file", x)))
 })
 
 test_that("integration of #16 in dockerize()", {
     ## verbose
-    gran_ok <- readRDS("../testdata/gran_ok.RDS")
+    rang_ok <- readRDS("../testdata/rang_ok.RDS")
     temp_dir <- .gen_temp_dir()
-    dockerize(granlist = gran_ok, output_dir = temp_dir) ## verbose = TRUE
-    x <- readLines(file.path(temp_dir, "gran.R"))
+    dockerize(rang = rang_ok, output_dir = temp_dir) ## verbose = TRUE
+    x <- readLines(file.path(temp_dir, "rang.R"))
     expect_true(any(grepl("^verbose <- TRUE", x)))
-    dockerize(granlist = gran_ok, output_dir = temp_dir, verbose = FALSE)
-    x <- readLines(file.path(temp_dir, "gran.R"))
+    dockerize(rang = rang_ok, output_dir = temp_dir, verbose = FALSE)
+    x <- readLines(file.path(temp_dir, "rang.R"))
     expect_true(any(grepl("^verbose <- FALSE", x)))
     ## lib
-    dockerize(granlist = gran_ok, output_dir = temp_dir) ## lib = NA
-    x <- readLines(file.path(temp_dir, "gran.R"))
[…]

### 2023-02-12 tests/testthat/test_resolve.R

@@ -27,7 +27,7 @@ test_that("unresolved", {
     expect_true(any(grepl("^Some package", warns)))
     expect_true(any(grepl("^No packages to query", warns)))
     expect_equal(x$unresolved_pkgrefs, "cran::LDAvis")
-    expect_equal(x$grans, list())
+    expect_equal(x$ranglets, list())
     expect_true(is.na(x$deps_sysreqs))
     ## a mixture; sna is old enough
     warns2 <- capture_warnings(x2 <- resolve(c("sna", "LDAvis"), snapshot_date = "2001-10-01"))
@@ -35,7 +35,7 @@ test_that("unresolved", {
     expect_true(any(grepl("^Some package", warns2)))
     expect_false(any(grepl("^No packages to query", warns2)))
     expect_equal(x2$unresolved_pkgrefs, "cran::LDAvis")
-    expect_false(length(x2$grans) == 0)
+    expect_false(length(x2$ranglets) == 0)
     expect_equal(x2$deps_sysreqs, character(0))
 })
 
@@ -51,32 +51,32 @@ test_that("issue #19", {
 test_that("cache #17", {
     skip_if_offline()
     skip_on_cran()
-    gran_ok <- readRDS("../testdata/gran_ok.RDS")
+    rang_ok <- readRDS("../testdata/rang_ok.RDS")
     temp_dir <- .gen_temp_dir()
-    dockerize(gran_ok, output_dir = temp_dir) ## cache = FALSE
+    dockerize(rang_ok, output_dir = temp_dir) ## cache = FALSE
     x <- readLines(file.path(temp_dir, "Dockerfile"))
     expect_false(any(grepl("^COPY cache", x)))
     expect_false(dir.exists(file.path(temp_dir, "cache")))
     expect_false(file.exists(file.path(temp_dir, "cache", "LDAvis_0.3.2.tar.gz")))
     expect_false(file.exists(file.path(temp_dir, "cache", "proxy_0.4-27.tar.gz")))
     expect_false(file.exists(file.path(temp_dir, "cache", "RJSONIO_1.3-1.6.tar.gz")))
-    expect_silent(dockerize(gran_ok, output_dir = temp_dir, cache = TRUE, verbose = FALSE))
+    expect_silent(dockerize(rang_ok, output_dir = temp_dir, cache = TRUE, verbose = FALSE))
     x <- readLines(file.path(temp_dir, "Dockerfile"))
[…]

### 2023-02-12 tests/testthat/test_s3.R

@@ -1,26 +1,26 @@
 test_that("printing with and without resolved packages", {
-    gran_mixture <- readRDS("../testdata/gran_mixture.RDS")
-    gran_ok <- readRDS("../testdata/gran_ok.RDS")
-    gran_unresolved <- readRDS("../testdata/gran_unresolved.RDS")
-    gran_output <- capture_output(print(gran_ok$grans[[1]]))
-    expect_equal(gran_output, "The latest version of `LDAvis` [cran] at 2023-01-01 was 0.3.2, which has 2 unique dependencies (2 with no dependencies.)")
-    gran_ok_output <- capture_output(print(gran_ok))
-    expect_true(grepl("^resolved: 1 package\\(s\\). Unresolved package\\(s\\): 0", gran_ok_output))
-    gran_mixture_output <- capture_output(print(gran_mixture))
-    expect_true(grepl("^resolved: 1 package\\(s\\). Unresolved package\\(s\\): 1", gran_mixture_output))
-    gran_unresolved_output <- capture_output(print(gran_unresolved))
-    expect_true(grepl("^resolved: 0 package\\(s\\). Unresolved package\\(s\\): 1", gran_unresolved_output))
-    expect_false(grepl("The latest version", gran_unresolved_output))
+    rang_mixture <- readRDS("../testdata/rang_mixture.RDS")
+    rang_ok <- readRDS("../testdata/rang_ok.RDS")
+    rang_unresolved <- readRDS("../testdata/rang_unresolved.RDS")
+    rang_output <- capture_output(print(rang_ok$ranglets[[1]]))
+    expect_equal(rang_output, "The latest version of `LDAvis` [cran] at 2023-01-01 was 0.3.2, which has 2 unique dependencies (2 with no dependencies.)")
+    rang_ok_output <- capture_output(print(rang_ok))
+    expect_true(grepl("^resolved: 1 package\\(s\\). Unresolved package\\(s\\): 0", rang_ok_output))
+    rang_mixture_output <- capture_output(print(rang_mixture))
+    expect_true(grepl("^resolved: 1 package\\(s\\). Unresolved package\\(s\\): 1", rang_mixture_output))
+    rang_unresolved_output <- capture_output(print(rang_unresolved))
+    expect_true(grepl("^resolved: 0 package\\(s\\). Unresolved package\\(s\\): 1", rang_unresolved_output))
[…]

### 2023-02-12 README.Rmd
Add warning in README [no ci]
Rename gran in README and logo
ref #35

@@
 
-# gran <img src="man/figures/gran_logo.svg" align="right" height="200" />
+# rang <img src="man/figures/rang_logo.svg" align="right" height="200" />
 
 <!-- badges: start -->
-[![R-CMD-check](https://github.com/chainsawriot/gran/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/chainsawriot/gran/actions/workflows/R-CMD-check.yaml)
+[![R-CMD-check](https://github.com/chainsawriot/rang/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/chainsawriot/rang/actions/workflows/R-CMD-check.yaml)
 <!-- badges: end -->
 
-The goal of gran (gran: R anachronism nullifier) [^gesis] is to obtain the dependency graph of R packages at a specific time point.
+The goal of rang (Reconstructing Ancient Number-crunching Gears) [^gesis] is to obtain the dependency graph of R packages at a specific time point.
 
@@
 For the former purpose, packages such as [renv](https://github.com/rstudio/renv/), [groundhog](https://github.com/CredibilityLab/groundhog), [miniCRAN](https://github.com/andrie/miniCRAN), and [Require](https://github.com/PredictiveEcology/Require) should be used.
-One can think of gran as an archaeological tool.
+One can think of rang as an archaeological tool.
 
@@
 
+**Important:** This package is in a very early stage. API might still change.
+
 ## Installation
 
-You can install the development version of gran like so:
+You can install the development version of rang like so:
 
 ``` r
-remotes::install_github("chainsawriot/gran")
+remotes::install_github("chainsawriot/rang")
 ```
@@
 ```r
-library(gran)
+library(rang)
 graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
@@
 
-The resolved result is an S3 object called `granlist` and can be exported as an installation script. The installation script can be execute on a vanilla R installation.
+The resolved result is an S3 object called `rang` and can be exported as an installation script. The installation script can be execute on a vanilla R installation.
 
 ```r
-export_granlist(graph, "gran.R")
+export_rang(graph, "rang.R")
 ```
@@
 
-A `granlist` object can be used to recreate the computational environment via [Rocker](https://github.com/rocker-org/rocker). Please note that the oldest R version one can get from Rocker is R 3.1.0.
+A `rang` object can be used to recreate the computational environment via [Rocker](https://github.com/rocker-org/rocker). Please note that the oldest R version one can get from Rocker is R 3.1.0.
 
@@
 cd ~/rocker_test
-sudo docker build -t gran .
-sudo docker run --rm --name "grantest" -ti gran
+sudo docker build -t rang .
+sudo docker run --rm --name "rangtest" -ti rang
 ```
@@
 cd ~/rocker_test
-sudo docker build -t gran .
-sudo docker run -p 8787:8787 -e PASSWORD=abc123 --rm --name "grantest" -ti gran
+sudo docker build -t rang .
+sudo docker run -p 8787:8787 -e PASSWORD=abc123 --rm --name "rangtest" -ti rang
 ```
@@
 
-`gran` can still be used to recreate computational environments for R < 3.1.0.
+`rang` can still be used to recreate computational environments for R < 3.1.0.
 The Dockerfile generated is based on Debian Woody (3.0) and the requested version of R is compiled from source.
@@
 ```r
-gran_rio <- resolve("rio", snapshot_date = "2013-08-28") ## R 3.0.1
-dockerize(gran_rio, output_dir = "~/old_renviron")
+rang_rio <- resolve("rio", snapshot_date = "2013-08-28") ## R 3.0.1
+dockerize(rang_rio, output_dir = "~/old_renviron")
 ```
@@
 
-The logo of gran is a remix of [this](https://commons.wikimedia.org/wiki/File:Flag_of_the_Canary_Islands.svg) public domain image.
+The logo of rang is a remix of [this](https://commons.wikimedia.org/wiki/File:Flag_of_the_Canary_Islands.svg) public domain image.
 The two dogs should be *Presa Canario*, the native dog breed on the islands of Gran Canaria and Tenerife.
@@
 
-[^gesis]: Actually, it stands for "GESIS R Archive Network".
+[^gesis]: It stands for "R Archiving Nerds at GESIS".
+The package was previously named `gran`, but we decided to rename it to `rang` because there is another package named [gRAN](https://cran.r-project.org/web/packages/GRANBase/vignettes/GRAN_Overview.html).

### 2023-02-12 paper/paper.rmd
Rewrite the paper to reflect the name change

--- 2023-02-02 paper/paper.rmd
@@
 ---
-title: 'gran: Reconstructing reproducible R computing environments with ease'
+title: 'rang: Reconstructing reproducible R computing environments with ease'
 tags:
@@
 using them now to ensure the reproducibility of the current research for future researchers.
-`gran` targets mostly on retrospective usage, i.e.
-using `gran` to reconstruct historical R computational environments which have not been completely declared.
-One can think of `gran` as an archaeological tool.
-In the following examples, `gran` is used to enable the reproducibility of published literature.
+`rang` targets mostly on retrospective usage, i.e.
+using `rang` to reconstruct historical R computational environments which have not been completely declared.
+One can think of `rang` as an archaeological tool.
+In the following examples, `rang` is used to enable the reproducibility of published literature.
 
@@
 
-There are two important functions of `gran`: `resolve()` and `dockerize()`.
+There are two important functions of `rang`: `resolve()` and `dockerize()`.
 
@@
 
-The resolved result is an S3 object called `granlist`.
-The information contained in a `granlist` can then be used to construct a computational environment in the similar manner as `containerit`, but with the packages and R versions pinned on the snapshot date.
+The resolved result is an S3 object called `rang`.
+The information contained in a `rang` can then be used to construct a computational environment in the similar manner as `containerit`, but with the packages and R versions pinned on the snapshot date.
 Then, `dockerize()` is used to generate the `Dockerfile` and other scripts in the `output_dir`.
@@
 For R < 3.1 but >= 2.1, a custom image based on Debian Woody is used.
-As of writing, `gran` does not support R < 2.1, i.e.
+As of writing, `rang` does not support R < 2.1, i.e.
 snapshot date earlier than 2005-04-19.
@@
 ```r
-dockerize(granlist graph, output_dir = "docker")
+dockerize(graph, output_dir = "docker")
 ```
@@
 ```r
-dockerize(granlist graph, output_dir = "docker", cache = TRUE)
+dockerize(graph, output_dir = "docker", cache = TRUE)
 ```
@@
 ```r
-dockerize(granlist graph, output_dir = "docker", cache = TRUE,
+dockerize(graph, output_dir = "docker", cache = TRUE,
           lib = "anotherlibrary")
@@
 ## might need to sudo
-docker build -t gran .
-docker run --rm --name "grantest" -ti gran
+docker build -t rang .
+docker run --rm --name "rangtest" -ti rang
 ```
@@
 
-Using `gran`, one can query the version of `quanteda` on 2018-10-06 and create a Docker container with all the "period appropriate" dependencies. Here, the `rstudio` Rocker image is selected.
+Using `rang`, one can query the version of `quanteda` on 2018-10-06 and create a Docker container with all the "period appropriate" dependencies. Here, the `rstudio` Rocker image is selected.
 
 ```r
-library(gran)
+library(rang)
 graph <- resolve(pkgs = "quanteda",
@@
 
-```{r, fig.cap = "The code snippet running in a R 3.5.1 container created with gran", echo = FALSE}
+```{r, fig.cap = "The code snippet running in a R 3.5.1 container created with rang", echo = FALSE}
 knitr::include_graphics("quanteda_rstudio.png", dpi = 300)

### 2023-02-13 R/resolve.R
Make sure `.system_requirements_github` work as expected
Also expand `gh` to `github` for consistency

@@ -18,8 +18,6 @@ NULL
     utils::tail(allvers[allvers$date < snapshot_date,], 1)$version
 }
 
-
-
 ## .msysreps <- memoise::memoise(.raw_sysreqs, cache = cachem::cache_mem(max_age = 60 * 60))
 
 ## .sysreps <- function(pkg, verbose = FALSE) {
@@ -38,7 +36,7 @@ NULL
                return(.get_snapshot_dependencies_cran(handle = .parse_pkgref(pkgref) ,snapshot_date = snapshot_date))
            },
            "github" = {
-               return(.get_snapshot_dependencies_gh(handle = .parse_pkgref(pkgref), snapshot_date = snapshot_date))
+               return(.get_snapshot_dependencies_github(handle = .parse_pkgref(pkgref), snapshot_date = snapshot_date))
            })
 }
 
@@ -60,7 +58,7 @@ NULL
     }
 }
 
-.get_snapshot_dependencies_gh <- function(handle = "schochastics/rtoot", snapshot_date = "2022-12-10") {
+.get_snapshot_dependencies_github <- function(handle = "schochastics/rtoot", snapshot_date = "2022-12-10") {
     snapshot_date <- anytime::anytime(snapshot_date, tz = "UTC", asUTC = TRUE)
     sha <- .get_sha(handle, snapshot_date)
     repo_descr <- gh::gh(paste0("GET /repos/", handle,"/contents/DESCRIPTION"), ref = sha$sha)
@@ -358,7 +356,7 @@ convert_edgelist <- function(x) {
                    query_fun <- .system_requirements_cran
                },
                "github" = {
-                   query_fun <- .system_requirements_gh
+                   query_fun <- .system_requirements_github
                }
                )
         tryCatch({
@@ -377,41 +375,40 @@ convert_edgelist <- function(x) {
 }
 
 ## get system requirements for github packages
-.system_requirements_gh <- function(handle, os){
-  DEFAULT_RSPM <- "https://packagemanager.rstudio.com"
-  DEFAULT_RSPM_REPO_ID <- "1"
-
-  curl = Sys.which("curl")
+.system_requirements_github <- function(handle, os) {
+    os_info <- strsplit(os, "-")[[1]]
+    DEFAULT_RSPM <- "https://packagemanager.rstudio.com"
+    DEFAULT_RSPM_REPO_ID <- "1"
+    curl <- Sys.which("curl")
[…]

### 2023-02-13 tests/testthat/test_resolve.R

@@ -88,15 +88,32 @@ test_that("cache for R < 3.1 and R >= 2.1", {
     expect_true(file.exists(file.path(temp_dir, "cache", "testthat_0.7.1.tar.gz")))
 })
 
+test_that(".system_requirements_github", {
+    skip_if_offline()
+    skip_on_cran()
+    res <- .system_requirements_github("schochastics/rtoot", os = "ubuntu-20.04")
+    expect_true(all(grepl("^apt-get", res)))
+    res <- .system_requirements_github("schochastics/rtoot", "opensuse-42.3")
+    expect_true(all(grepl("^zypper", res)))
+    res <- .system_requirements_github("schochastics/rtoot", "centos-8")
+    expect_true(all(grepl("^dnf", res)))
+})
+
 test_that("github correct querying", {
+    skip_if_offline()
+    skip_on_cran()
     x <- resolve(c("cran/sna"), snapshot_date = "2020-05-01", get_sysreqs = FALSE)
     expect_equal(length(x$ranglets[[1]]$deps), 19)
     x <- resolve("schochastics/netUtils", snapshot_date = "2020-06-01", get_sysreqs = FALSE)
     expect_equal(x$ranglets[[1]]$pkgref, "github::schochastics/netUtils")
     expect_equal(unique(x$ranglets[[1]]$original$x), "igraphUtils")
+    x <- resolve("tidyverse/stringr", snapshot_date = "2022-12-31")
+    expect_true(all(grepl("^apt-get", x$deps_sysreqs)))
 })
 
 test_that("Non-cran must enforce caching ref #22", {
+    skip_if_offline()
+    skip_on_cran()
     temp_dir <- .gen_temp_dir()
     graph <- readRDS("../testdata/ancientsna.RDS")
     expect_equal(graph$ranglets[[1]]$pkgref, "github::cran/sna")

### 2023-02-13 README.Rmd
Update doc and bump version [no ci]

@@
 
-To obtain the dependency graph of an R package, use `resolve`
+To obtain the dependency graph of R packages, use `resolve`. Currently, this package supports both CRAN and Github packages.
 
@@
 library(rang)
+resolve(pkgs = c("sna", "schochastics/rtoot"), snapshot_date = "2022-11-30")
+```
+
+```r
 graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),

### 2023-02-13 R/installation.R

@@ -262,11 +262,11 @@
       basic_docker[rang_line:length(basic_docker)])
 }
 
-.insert_materials_dir <- function(basic_docker){
-  rang_line <- which(basic_docker == "COPY rang.R ./rang.R")
-  c(basic_docker[1:rang_line], 
-    "COPY materials/ ./materials/",
-    basic_docker[(rang_line + 1):length(basic_docker)])
+.insert_materials_dir <- function(basic_docker) {
+    rang_line <- which(basic_docker == "COPY rang.R ./rang.R")
+    c(basic_docker[1:rang_line], 
+      "COPY materials/ ./materials/",
+      basic_docker[(rang_line + 1):length(basic_docker)])
 }
 
 .generate_pre310_docker <- function(r_version, lib, sysreqs_cmd, cache, debian_version = "lenny") {
@@ -361,7 +361,7 @@ export_rang <- function(rang, path, rang_as_comment = TRUE, verbose = TRUE, lib 
 #' @param materials_dir character, path to the directiry containing dditional resources (e.g. analysis scripts) to be copied into `output_dir` and in turn into the Docker container
 #' @param image character, which versioned Rocker image to use. Can only be "r-ver", "rstudio", "tidyverse", "verse", "geospatial"
 #' This applies only to R version <= 3.1
-#' @param cache logical, whether to cache the packages now. Please note that the system requirements are not cached. For query with non-CRAN packages, this option is strongly recommended.
+#' @param cache logical, whether to cache the packages now. Please note that the system requirements are not cached. For query with non-CRAN packages, this option is strongly recommended. For R version < 3.1, this must be TRUE if there is any non-CRAN packages.
 #' @param ... arguments to be passed to `dockerize`
 #' @return `output_dir`, invisibly
 #' @inheritParams export_rang

### 2023-02-13 R/resolve.R

@@ -118,7 +118,7 @@ NULL
             type = unlist(type),
             y_raw_version = unlist(version)
         ))
-    } else{
+    } else {
         return(data.frame(
             snapshot_date = as.Date(snapshot_date),
             x = descr_df[["Package"]],

### 2023-02-14 R/resolve.R
Fix #45

@@ -444,7 +444,7 @@ query_sysreqs <- function(rang, os = "ubuntu-20.04") {
             "--silent",
             "--data-binary",
             shQuote(paste0("@", desc_file)),
-            shQuote(sprintf("%s/sysreqs?distribution=%s&release=%s&suggests=true",
+            shQuote(sprintf("%s/sysreqs?distribution=%s&release=%s&suggests=false",
                             rspm_repo_url,
                             os_info[1],
                             os_info[2])

### 2023-02-14 tests/testthat/test_resolve.R

@@ -95,6 +95,7 @@ test_that(".system_requirements_github", {
     skip_on_cran()
     res <- .system_requirements_github("schochastics/rtoot", os = "ubuntu-20.04")
     expect_true(all(grepl("^apt-get", res)))
+    expect_true(length(res) == 2) ## issue #45
     res <- .system_requirements_github("schochastics/rtoot", "opensuse-42.3")
     expect_true(all(grepl("^zypper", res)))
     res <- .system_requirements_github("schochastics/rtoot", "centos-8")

### 2023-02-14 paper/paper.rmd

@@
 By default, the container building process downloads source packages from CRAN and then compiles them from source.
-This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continous availability since 1997-04-23) [^1].
-However, it is possible also to cache (or archive) the source packages now.
+This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) [^1].
+However, it is also possible to cache (or archive) the source packages now.
 The archived R packages can then be used instead during the building process.

### 2023-02-14 R/installation.R
Clean all verbs #44

@@ -1,4 +1,5 @@
-.safe_get_uid <- function(pkgref, uid) {
+## here, uid is a fastmap; see below
+.extract_uid_safe <- function(pkgref, uid) {
     if (is.null(uid$get(pkgref))) {
         return(NA_character_)
     } else {
@@ -6,7 +7,7 @@
     }
 }
 
-.determine_installation_order <- function(rang) {
+.generate_installation_order <- function(rang) {
     dep <- fastmap::fastmap()
     version <- fastmap::fastmap()
     uid <- fastmap::fastmap()
@@ -16,7 +17,7 @@
     for (ranglet in rang$ranglets) {
         current_pkgref <- unique(ranglet$original$x_pkgref)
         current_ver <- unique(ranglet$original$x_version)
-        current_dep <- .keep_queryable_dependencies(dep_df = ranglet$original, no_enhances = ranglet$no_enhances, no_suggests = ranglet$no_suggests)
+        current_dep <- .extract_queryable_dependencies(dep_df = ranglet$original, no_enhances = ranglet$no_enhances, no_suggests = ranglet$no_suggests)
         current_pkgname <- unique(ranglet$original$x)
         dep$set(current_pkgref, current_dep)
         version$set(current_pkgref, current_ver)
@@ -27,7 +28,7 @@
         for (dep_df in ranglet$deps) {
             current_pkgref <- unique(dep_df$x_pkgref)
             current_ver <- unique(dep_df$x_version)
-            current_dep <- .keep_queryable_dependencies(dep_df = dep_df, no_enhances = ranglet$no_enhances, no_suggests = ranglet$no_suggests)
+            current_dep <- .extract_queryable_dependencies(dep_df = dep_df, no_enhances = ranglet$no_enhances, no_suggests = ranglet$no_suggests)
             current_pkgname <- unique(dep_df$x)
             dep$set(current_pkgref, current_dep)
             version$set(current_pkgref, current_ver)
@@ -43,7 +44,7 @@
     needed_pkgrefs <- dep$keys()
     ## install all terminal nodes
     for (pkgref in needed_pkgrefs) {
-        if(.is_github(pkgref)) {
+        if (.is_github(pkgref)) {
           noncran_pkgrefs <- c(noncran_pkgrefs, pkgref)
           next()
         }
@@ -68,32 +69,13 @@
     ordered_version <- vapply(ordered_pkgrefs, function(x) version$get(x), character(1), USE.NAMES = FALSE)
[…]

### 2023-02-14 R/memo_misc.R

@@ -1,0 +1,32 @@
+#' @importFrom memoise memoise
+#' @importFrom pkgsearch cran_package_history
+NULL
+
+## one hr
+
+.memo_search<- memoise::memoise(pkgsearch::cran_package_history, cache = cachem::cache_mem(max_age = 60 * 60))
+
+.rver <- function() {
+    suppressWarnings(jsonlite::fromJSON(readLines("https://api.r-hub.io/rversions/r-versions"), simplifyVector = TRUE))
+}
+
+.memo_rver <- memoise::memoise(.rver, cache = cachem::cache_mem(max_age = 120 * 60))
+
+## internal data generation
+## ---
+## os <- names(remotes:::supported_os_versions())
+## supported_os <- unlist(mapply(function(x, y) paste(x,"-", y, sep = ""), os, remotes:::supported_os_versions()))
+## names(supported_os) <- NULL
+## cached_rver <- .rver()
+## attr(cached_rver, "newest_date") <- anytime::anytime(tail(cached_rver, n = 1)$date, tz = "UTC", asUTC = TRUE)
+## usethis::use_data(supported_os, cached_rver, internal = TRUE, overwrite = TRUE)
+
+## test data upgrade
+## ---
+## devtools::load_all()
+## all_rds_files <- list.files(here::here("tests/testdata"), ".RDS", full.names = TRUE)
+## for (rds in all_rds_files) {
+##     x <- readRDS(rds)
+##     y <- eval(x$call)
+##     saveRDS(y, rds)
+## }

### 2023-02-14 R/resolve.R

@@ -1,18 +1,4 @@
-#' @importFrom memoise memoise
-#' @importFrom pkgsearch cran_package_history
-NULL
-
-## one hr
-
-.search <- memoise::memoise(pkgsearch::cran_package_history, cache = cachem::cache_mem(max_age = 60 * 60))
-
-.rver <- function() {
-    suppressWarnings(jsonlite::fromJSON(readLines("https://api.r-hub.io/rversions/r-versions"), simplifyVector = TRUE))
-}
-
-.memo_rver <- memoise::memoise(.rver, cache = cachem::cache_mem(max_age = 120 * 60))
-
-.get_rver <- function(snapshot_date) {
+.query_rver <- function(snapshot_date) {
     if (snapshot_date < attr(cached_rver, "newest_date")) {
         allvers <- cached_rver
     } else {
@@ -33,20 +19,20 @@ NULL
 
 ## get the latest version as of date
 ## let's call this output dep_df; basically is a rough version of edgelist
-.get_snapshot_dependencies <- function(pkgref = "cran::rtoot", snapshot_date = "2022-12-10") {
+.query_snapshot_dependencies <- function(pkgref = "cran::rtoot", snapshot_date = "2022-12-10") {
     source <- .parse_pkgref(pkgref, return_handle = FALSE)
     switch(source,
            "cran" = {
-               return(.get_snapshot_dependencies_cran(handle = .parse_pkgref(pkgref) ,snapshot_date = snapshot_date))
+               return(.query_snapshot_dependencies_cran(handle = .parse_pkgref(pkgref) ,snapshot_date = snapshot_date))
            },
            "github" = {
-               return(.get_snapshot_dependencies_github(handle = .parse_pkgref(pkgref), snapshot_date = snapshot_date))
+               return(.query_snapshot_dependencies_github(handle = .parse_pkgref(pkgref), snapshot_date = snapshot_date))
            })
 }
 
-.get_snapshot_dependencies_cran <- function(handle = "rtoot", snapshot_date = "2022-12-10") {
+.query_snapshot_dependencies_cran <- function(handle = "rtoot", snapshot_date = "2022-12-10") {
     snapshot_date <- anytime::anytime(snapshot_date, tz = "UTC", asUTC = TRUE)
-    search_res <- .search(handle)
+    search_res <- .memo_search(handle)
[…]

### 2023-02-14 tests/testthat/test_dockerize.R

@@ -1,4 +1,4 @@
-.gen_temp_dir <- function() {
+.generate_temp_dir <- function() {
     file.path(tempdir(), paste(sample(c(LETTERS, letters), 20, replace = TRUE), collapse = ""))
 }
 
@@ -9,7 +9,7 @@ test_that("defensive programming", {
 
 test_that("integration of #13 in dockerize()", {
     rang_ok <- readRDS("../testdata/rang_ok.RDS")
-    temp_dir <- .gen_temp_dir()
+    temp_dir <- .generate_temp_dir()
     dockerize(rang = rang_ok, output_dir = temp_dir) ## rang_as_comment = TRUE
     x <- readLines(file.path(temp_dir, "rang.R"))
     expect_true(any(grepl("^## ## To reconstruct this file", x)))
@@ -21,7 +21,7 @@ test_that("integration of #13 in dockerize()", {
 test_that("integration of #16 in dockerize()", {
     ## verbose
     rang_ok <- readRDS("../testdata/rang_ok.RDS")
-    temp_dir <- .gen_temp_dir()
+    temp_dir <- .generate_temp_dir()
     dockerize(rang = rang_ok, output_dir = temp_dir) ## verbose = TRUE
     x <- readLines(file.path(temp_dir, "rang.R"))
     expect_true(any(grepl("^verbose <- TRUE", x)))
@@ -43,7 +43,7 @@ test_that("integration of #16 in dockerize()", {
 
 test_that("integration of #18 in dockerize()", {
     rang_ok <- readRDS("../testdata/rang_ok.RDS")
-    temp_dir <- .gen_temp_dir()
+    temp_dir <- .generate_temp_dir()
     dockerize(rang = rang_ok, output_dir = temp_dir) ## cran_mirror = "https://cran.r-project.org/"
     x <- readLines(file.path(temp_dir, "rang.R"))
     expect_true(any(grepl("^cran_mirror <- \"https://cran\\.r\\-project\\.org/\"", x)))
@@ -62,7 +62,7 @@ test_that("integration of #18 in dockerize()", {
 test_that("integration of #20 to dockerize()", {
     rang_ok <- readRDS("../testdata/rang_ok.RDS")
     expect_equal(rang_ok$r_version, "4.2.2")
-    temp_dir <- .gen_temp_dir()
+    temp_dir <- .generate_temp_dir()
     dockerize(rang_ok, output_dir = temp_dir) ## cran_mirror = "https://cran.r-project.org/"
[…]

### 2023-02-14 tests/testthat/test_expost_rang.R

@@ -53,10 +53,10 @@ test_that(".normalize_url https issue #20", {
     expect_equal(x, "http://cran.r-project.org/")
 })
 
-test_that(".check_mirror", {
-    expect_true(.check_mirror("https://cran.r-project.org/"))
-    expect_true(.check_mirror("https://cloud.r-project.org/"))
-    expect_false(.check_mirror("https://www.chainsawriot.com/"))
+test_that(".query_mirror_validity", {
+    expect_true(.query_mirror_validity("https://cran.r-project.org/"))
+    expect_true(.query_mirror_validity("https://cloud.r-project.org/"))
+    expect_false(.query_mirror_validity("https://www.chainsawriot.com/"))
 })
 
 test_that("integration of mirror selection to `export_rang` #18", {

### 2023-02-14 tests/testthat/test_resolve.R

@@ -1,4 +1,4 @@
-.gen_temp_dir <- function() {
+.generate_temp_dir <- function() {
     file.path(tempdir(), paste(sample(c(LETTERS, letters), 20, replace = TRUE), collapse = ""))
 }
 
@@ -14,7 +14,7 @@ test_that("normal", {
     skip_on_cran()
     expect_silent(x <- resolve("LDAvis", snapshot_date = "2023-01-01"))
     expect_equal(x$unresolved_pkgrefs, character(0))
-    expect_equal(x$deps_sysreqs, character(0))
+    expect_equal(x$sysreqs, character(0))
     expect_equal(x$r_version, "4.2.2")
 })
 
@@ -22,7 +22,7 @@ test_that("unresolved", {
     ## too early for LDAvis
     skip_if_offline()
     skip_on_cran()
-    expect_error(res <- .get_snapshot_dependencies("cran::LDAvis", snapshot_date = "2001-10-01"))
+    expect_error(res <- .query_snapshot_dependencies("cran::LDAvis", snapshot_date = "2001-10-01"))
     warns <- capture_warnings(x <- resolve("LDAvis", snapshot_date = "2000-10-01"))
     expect_equal(x$r_version, "1.1.1")    
     expect_true(length(warns) == 2)
@@ -30,7 +30,7 @@ test_that("unresolved", {
     expect_true(any(grepl("^No packages to query", warns)))
     expect_equal(x$unresolved_pkgrefs, "cran::LDAvis")
     expect_equal(x$ranglets, list())
-    expect_true(is.na(x$deps_sysreqs))
+    expect_true(is.na(x$sysreqs))
     ## a mixture; sna is old enough
     warns2 <- capture_warnings(x2 <- resolve(c("sna", "LDAvis"), snapshot_date = "2001-10-01"))
     expect_true(length(warns2) == 1)
@@ -38,7 +38,7 @@ test_that("unresolved", {
     expect_false(any(grepl("^No packages to query", warns2)))
     expect_equal(x2$unresolved_pkgrefs, "cran::LDAvis")
     expect_false(length(x2$ranglets) == 0)
-    expect_equal(x2$deps_sysreqs, character(0))
+    expect_equal(x2$sysreqs, character(0))
 })
 
 test_that("issue #19", {
@@ -47,14 +47,14 @@ test_that("issue #19", {
[…]

### 2023-02-15 R/resolve.R
Update doc to give more info about the S3 object [no ci]

@@ -170,7 +170,16 @@
 #' packages are assumed to have no system requirement.
 #' @param os character, which OS to query for system requirements
 #' @param verbose logical, whether to display messages
-#' @return S3 object `rang`
+#' @return a `rang` S3 object with the following items
+#' \item{call}{original function call}
+#' \item{ranglets}{List of dependency graphs of all packages in `pkgs`}
+#' \item{snapshot_date}{`snapshot_date`}
+#' \item{no_enhances}{`no_enhances`}
+#' \item{no_suggests}{`no_suggests`}
+#' \item{unresolved_pkgsrefs}{Packages that can't be resolved}
+#' \item{sysreqs}{System requirements as Linux commands}
+#' \item{r_version}{The latest R version as of `snapshot_date`}
+#' \item{os}{`os`}
 #' @export
 #' @seealso [dockerize()]
 #' @references

### 2023-02-15 R/installation.R
adding readme to the output folder (#50)

@@ -249,6 +249,18 @@
     return(dockerfile_content)
 }
 
+.generate_docker_readme <- function(output_dir,image){
+  file.create(file.path(output_dir,"README"))
+  con <- file(file.path(output_dir,"README"), open="w")
+  readme <- readLines(system.file("readme_template.txt", package = "rang"))
+  readme <- gsub("__DATE__",Sys.Date(),readme)
+  readme <- gsub("__OUTPUT__",output_dir,readme)
+  readme <- gsub("__IMAGE__",image,readme)
+  writeLines(readme,file.path(output_dir,"README"))
+  close(con)
+  invisible(readme)
+}
+
 #' Export The Resolved Result As Installation Script
 #'
 #' This function exports the results from [resolve()] to an installation script that can be run in a fresh R environment.
@@ -404,6 +416,9 @@ dockerize <- function(rang, output_dir, materials_dir = NULL, image = c("r-ver",
         dockerfile_content <- .insert_materials_dir(dockerfile_content)
     }
     writeLines(dockerfile_content, file.path(output_dir, "Dockerfile"))
+    
+    .generate_docker_readme(output_dir = output_dir,image = image)
+    
     invisible(output_dir)
 }
 

### 2023-02-15 README.Rmd
Update README and bump version ref #44
Fix issues from WinBuilder and r-hub [no CI]

@@
 
-**Important:** This package is in a very early stage. API might still change.
-
 ## Installation
@@
 library(rang)
-resolve(pkgs = c("sna", "schochastics/rtoot"), snapshot_date = "2022-11-30")
+x <- resolve(pkgs = c("sna", "schochastics/rtoot"), snapshot_date = "2022-11-30")
 ```
@@
 ```{r example2}
-graph$deps_sysreqs
+graph$sysreqs
 ```
@@
 [^gesis]: It stands for "R Archiving Nerds at GESIS".
-The package was previously named `gran`, but we decided to rename it to `rang` because there is another package named [gRAN](https://cran.r-project.org/web/packages/GRANBase/vignettes/GRAN_Overview.html).
+The package was previously named `gran`, but we decided to rename it to `rang` because there is another package named [gRAN](https://CRAN.R-project.org/package=GRANBase).

### 2023-02-15 R/memo_misc.R

@@ -1,5 +1,6 @@
 #' @importFrom memoise memoise
 #' @importFrom pkgsearch cran_package_history
+#' @importFrom remotes system_requirements
 NULL
 
 ## one hr

### 2023-02-15 tests/testthat/test_dockerize.R
Add tests for #50

@@ -190,3 +190,13 @@ test_that("material_dir, existing, with 1 subdir, #23", {
     expect_equal(content[1], "831721")
     expect_equal(content[2], "GESIS")
 })
+
+test_that("readme issue #50", {
+    rang_ok <- readRDS("../testdata/rang_ok.RDS")
+    expect_equal(rang_ok$r_version, "4.2.2")
+    temp_dir <- .generate_temp_dir()
+    dockerize(rang_ok, output_dir = temp_dir)
+    expect_true(file.exists(file.path(temp_dir, "README")))
+    content <- readLines(file.path(temp_dir, "README"))
+    expect_true(any(grepl(temp_dir, content)))
+})

### 2023-02-15 tests/testthat/test_dockerize.R
Disable a test case that might fail on Windows

@@ -192,6 +192,8 @@ test_that("material_dir, existing, with 1 subdir, #23", {
 })
 
 test_that("readme issue #50", {
+    skip_on_os("windows", arch = NULL)
+    skip_on_cran()
     rang_ok <- readRDS("../testdata/rang_ok.RDS")
     expect_equal(rang_ok$r_version, "4.2.2")
     temp_dir <- .generate_temp_dir()

### 2023-02-15 R/installation.R
Fix spelling [no ci]

@@ -331,7 +331,7 @@ export_rang <- function(rang, path, rang_as_comment = TRUE, verbose = TRUE, lib 
 #' This function exports the result from [resolve()] to a Docker file. For R version >= 3.1.0, the Dockerfile is based on the versioned Rocker image.
 #' For R version < 3.1.0, the Dockerfile is based on Debian and it compiles R from source.
 #' @param output_dir character, where to put the Docker file and associated content
-#' @param materials_dir character, path to the directiry containing dditional resources (e.g. analysis scripts) to be copied into `output_dir` and in turn into the Docker container
+#' @param materials_dir character, path to the directory containing additional resources (e.g. analysis scripts) to be copied into `output_dir` and in turn into the Docker container
 #' @param image character, which versioned Rocker image to use. Can only be "r-ver", "rstudio", "tidyverse", "verse", "geospatial"
 #' This applies only to R version <= 3.1
 #' @param cache logical, whether to cache the packages now. Please note that the system requirements are not cached. For query with non-CRAN packages, this option is strongly recommended. For R version < 3.1, this must be TRUE if there is any non-CRAN packages.

### 2023-02-15 README.Rmd
Clean README a little [no ci]

@@
 cd ~/rocker_test
-sudo docker build -t rang .
-sudo docker run --rm --name "rangtest" -ti rang
+docker build -t rang .
+docker run --rm --name "rangtest" -ti rang
 ```
@@
 
-One can also cache (or archive) the R packages from CRAN at the time `dockerize` is executed.
+One can also cache (or archive) the R packages from CRAN and Github at the time `dockerize` is executed.
 The cached R packages will then transfer to the container.
@@
 cd ~/rocker_test
-sudo docker build -t rang .
-sudo docker run -p 8787:8787 -e PASSWORD=abc123 --rm --name "rangtest" -ti rang
+docker build -t rang .
+docker run -p 8787:8787 -e PASSWORD=abc123 --rm --name "rangtest" -ti rang
 ```

### 2023-02-15 paper/paper.rmd
Add Hilgard

--- 2023-02-12 paper/paper.rmd
@@
 ---
-title: 'rang: Reconstructing reproducible R computing environments with ease'
+title: 'rang: Reconstructing reproducible R computational environments with ease'
 tags:
@@
 abstract: |
-  Enter the text of your abstract here.
+  A complete description of the computational environment is often missing when researchers share their materials.
+Without such description, software obsolescence and missing system components can jeopardize computational reproducibility in the future, even when data and computer code are available.
+The R package rang is a complete solution for generating instructions for other researchers to completely reconstruct the computational environment at a specific time point.
+The reconstruction process, based on Docker, has been tested for R code as old as 2005 and does not depend on the long term availability of any commericial service.
+The description generated by rang satisfies the definition of reproducibile compendium and can be shared as such.
 keywords:
-  - blah
-  - blee
-  - bloo
-  - these are optional and can be removed
+  - R
+  - reproducibility
+  - docker
 date: 02 February 2023
@@
 For instance, if there was a computational environment constructed on 2020-01-16 (called "snapshot date") with the several natural language processing R packages, `resolve()` can be used to resolve all the dependencies of these R packages.
+Currently, `rang` supports both CRAN and Github packages.
 
@@
 
+## Psychological Science
+
+@cruewell:2023:WB evaluate the computational reproducibility of 14 articles published in **Psyhocological Science**.
+Among these articles, the paper by @hilgard:2019:NEG has been rated as having "package dependency issues".
+
+All data and computer code are available from Github with the last commit on 2019-01-17.
+The R code contains a list of R packages used in the project, included an R package on Github that is written by the main author of that paper.
+
+Based on the shared, one can run `rang` to obtain the dependecy graph of all R packages on 2019-01-17.
+
+```r
+graph <- resolve(c("readxl", "rms", "ordinal", "tidyverse", "lubridate", "MBESS", "censReg",
+                   "BayesFactor", "psych", "Joe-Hilgard/hilgard", "lsmeans"),
+                 snapshot_date = "2019-01-17")
+```
+
 # References

### 2023-02-16 paper/paper.rmd
Add hilgard

@@
 
-@cruewell:2023:WB evaluate the computational reproducibility of 14 articles published in **Psyhocological Science**.
+ @cruewell:2023:WB evaluate the computational reproducibility of 14 articles published in *Psyhocological Science*.
 Among these articles, the paper by @hilgard:2019:NEG has been rated as having "package dependency issues".
 
-All data and computer code are available from Github with the last commit on 2019-01-17.
-The R code contains a list of R packages used in the project, included an R package on Github that is written by the main author of that paper.
+All data and computer code are available from Github with the last commit on 2019-01-17 [^hilgard].
+The R code contains a list of R packages used in the project as `library()` statements, included an R package on Github that is written by the main author of that paper.
+However, we identified one package (`compute.es`) that was not written in those `library()` statements but used with the namespace operator, i.e.
+`compute.es::tes()`.
+This undocumented package can be detected by `renv::dependencies()`.
 
-Based on the shared, one can run `rang` to obtain the dependecy graph of all R packages on 2019-01-17.
+[^hilgard]: https://github.com/Joe-Hilgard/vvg-2d4d
+
+Based on the above information, one can run `resolve()` to obtain the dependency graph of all R packages on 2019-01-17.
 
 ```r
-graph <- resolve(c("readxl", "rms", "ordinal", "tidyverse", "lubridate", "MBESS", "censReg",
-                   "BayesFactor", "psych", "Joe-Hilgard/hilgard", "lsmeans"),
+graph <- resolve(c("readxl", "rms", "ordinal", "tidyverse", "lubridate", "MBESS",
+                   "censReg", "BayesFactor", "psych", "Joe-Hilgard/hilgard", "lsmeans",
+                   "compute.es"),
                  snapshot_date = "2019-01-17")
@@
 
+When running `dockerize()`, one can take advantage of the `materials_dir` parameter to transfer the shared materials from @hilgard:2019:NEG into the Docker image.
+
+```r
+dockerize(graph, "hilgard", materials_dir = "vvg-2d4d", cache = TRUE)
+```
+
+We then built the Docker and launch a Docker container. For this container, we changed the entry point from R to bash so that the container goes to the Linux command shell instead.
+
+```r
+cd hilgard
+docker build -t hilgard .
+docker run --rm --name "hilgardcontainer" --entrypoint bash -ti hilgard
+```
+
+Inside the container, the materials is located in the `materials` directory. We used the following shell script to test the reproducibility of all R scripts.
+
+```sh
+rfiles=(0_data_aggregation.R 1_data_cleaning.R 2_analysis.R 3_plotting.R)
+for i in ${rfiles[@]}
+do
+	Rscript $i
+	code=$?
+	if [ $code != 0 ]
+	then
+		exit 1
+	fi
+done   
+```
+
+All R scripts ran fine inside the container and the figures generated are the same as the ones in @hilgard:2019:NEG.
+
 # References

### 2023-02-16 paper/paper.rmd

@@
 The reconstruction process, based on Docker, has been tested for R code as old as 2005 and does not depend on the long term availability of any commericial service.
-The description generated by rang satisfies the definition of reproducibile compendium and can be shared as such.
+The description generated by rang satisfies the definition of a reproducibile compendium and can be shared as such.
 keywords:
@@
 using them now to ensure the reproducibility of the current research for future researchers.
-`rang` targets mostly on retrospective usage, i.e.
+`rang` mostly targets retrospective usage, i.e.
 using `rang` to reconstruct historical R computational environments which have not been completely declared.
@@
 For instance, if there was a computational environment constructed on 2020-01-16 (called "snapshot date") with the several natural language processing R packages, `resolve()` can be used to resolve all the dependencies of these R packages.
-Currently, `rang` supports both CRAN and Github packages.
+Currently, `rang` supports both CRAN and GitHub packages.
 
@@
 The resolved result is an S3 object called `rang`.
-The information contained in a `rang` can then be used to construct a computational environment in the similar manner as `containerit`, but with the packages and R versions pinned on the snapshot date.
+The information contained in a `rang` can then be used to construct a computational environment in a similar manner as `containerit`, but with the packages and R versions pinned on the snapshot date.
 Then, `dockerize()` is used to generate the `Dockerfile` and other scripts in the `output_dir`.
@@
 
-All data and computer code are available from Github with the last commit on 2019-01-17 [^hilgard].
-The R code contains a list of R packages used in the project as `library()` statements, included an R package on Github that is written by the main author of that paper.
+All data and computer code are available from GitHub with the last commit on 2019-01-17 [^hilgard].
+The R code contains a list of R packages used in the project as `library()` statements, including an R package on GitHub that is written by the main author of that paper.
 However, we identified one package (`compute.es`) that was not written in those `library()` statements but used with the namespace operator, i.e.
@@
 
-Inside the container, the materials is located in the `materials` directory. We used the following shell script to test the reproducibility of all R scripts.
+Inside the container, the materials are located in the `materials` directory. We used the following shell script to test the reproducibility of all R scripts.
 

### 2023-02-17 paper/paper.rmd
Add political analysis and readme

@@
 The reconstruction process, based on Docker, has been tested for R code as old as 2005 and does not depend on the long term availability of any commericial service.
-The description generated by rang satisfies the definition of a reproducibile compendium and can be shared as such.
+The description generated by rang satisfies the definition of a reproducible research compendium and can be shared as such.
 keywords:
@@
 ```sh
+cd materials
 rfiles=(0_data_aggregation.R 1_data_cleaning.R 2_analysis.R 3_plotting.R)
@@
 
+## Political Analysis
+
+The study by @trisovic:2022 evaluates the reproducibility of R scripts shared on Dataverse.
+They found that 75\% of R scripts cannot be successfully executed.
+Among these failed R scripts is an R script shared by @beck:2019:EGD.
+
+This R script has been "rescued" by the author of the R package `groundhog` [@groundhogrpkg], as demonstrated in a blog post [^groundhog].
+We were wondering if `rang` can also be used to rescue the concerned R script.
+
+The date of the R script, as indicated on Dataverse, is 2018-12-12. This date is used as the snapshot date.
+
+```r
+graph <- resolve(c("foreign", "bife"), snapshot_date = "2018-12-12")
+dockerize(graph, output_dir = "nat", materials_dir = "nathaniel")
+```
+
+```r
+cd nat
+docker build -t nat .
+docker run --rm --name "natcontainer" --entrypoint bash -ti nat
+```
+
+Inside the container
+
+```r
+cd materials
+Rscript fn_5.R
+```
+
+The same file can also be "rescued" by `rang`.
+
+[^groundhog]: http://datacolada.org/100
+
+## Resuscitate a discontinued reproducible format
+
+cacher
+
 # References

### 2023-02-20 tests/testthat/test_dockerize.R
test dockerize with bioc

@@ -202,3 +202,11 @@ test_that("readme issue #50", {
     content <- readLines(file.path(temp_dir, "README"))
     expect_true(any(grepl(temp_dir, content)))
 })
+
+test_that("dockerize with bioc #58", {
+  rang_bioc <- readRDS("../testdata/rang_bioc.RDS")
+  temp_dir <- .generate_temp_dir()
+  dockerize(rang = rang_bioc, output_dir = temp_dir) ## verbose = TRUE
+  x <- readLines(file.path(temp_dir, "rang.R"))
+  expect_true(any(grepl("bioc_mirror",x)))
+})

### 2023-02-20 paper/paper.rmd
Add `maxent` example

@@
 
-## Resuscitate a discontinued reproducible format
+## Recover a removed R package
 
-cacher
+The R package `maxent` introduces a machine learning algorithm with a small memory footprint and was available on CRAN until 2019.
+A software paper was published by the original authors in 2012 [@jurka:2012].
+The R package was also used in some subsequent automated content analytic papers [e.g.
+@loercher:2017:D].
+
+Despite the covert editing of the package by a staffer of CRAN [^evidence], the package was removed from CRAN in 2019 [^checkerror].
+We attempted to install the second last (the original submitted version) and last (with covert editing) versions of `maxent` on R 4.2.2.
+Both of them didn't work.
+
+Using `rang`, we are able to reconstruct a computational environment with R 2.15.0 (2012-03-30) to run all code snippets published in @jurka:2012 [^speed].
+For removed CRAN packages, we strongly recommend querying the Github read-only mirror of CRAN instead (https://github.com/cran).
+It is because in this way, the resolved system requirements have a higher chance of being correct.
+
+```r
+maxent <- resolve("cran/maxent", "2012-06-10")
+dockerize(maxent, "maxentdir", cache = TRUE)
+```
+
+[^checkerror]: https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html
+
+[^evidence]: https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9
+
+[^speed]: On an interesting historical note: The original paper reported, based on a benchmark, that "the algorithm is very fast; `maxent` uses only 135.4 megabytes of RAM and finishes in 53.3 seconds." On a modest computer in 2023 with a dockerized R 2.15.0, the benchmark finishes in 4 seconds.
 

### 2023-02-24 paper/paper.qmd
Switch over to quarto

--- 2023-02-20 paper/paper.rmd
@@
   - docker
-date: 02 February 2023
 year: 2023
 bibliography: paper.bib
-biblio-style: unsrt
-output: rticles::arxiv_article
+format:
+  pdf: default
+  arxiv-pdf:
+    keep-tex: true    
 ---
@@
 
-```{r, fig.cap = "The code snippet running in a R 3.5.1 container created with rang", echo = FALSE}
+```{r}
+#| fig.cap: The code snippet running in a R 3.5.1 container created with rang
+#| echo: FALSE
 knitr::include_graphics("quanteda_rstudio.png", dpi = 300)

### 2023-02-24 R/as_pkgrefs.R
added as_pkgrefs.renv_lockfile (#55)

@@ -25,6 +25,9 @@ as_pkgrefs.default <- function(x, ...) {
     if (is.numeric(x) || is.logical(x) || is.integer(x)) {
         stop("Don't know how to convert this to package references.", call. = FALSE)
     }
+    if(.detect_renv_lockfile(x)){
+        return(as_pkgrefs(.parse_renv_lockfile(x)))
+    }
     return(.normalize_pkgs(x))
 }
 
@@ -32,6 +35,27 @@ as_pkgrefs.default <- function(x, ...) {
 #' @export
 as_pkgrefs.sessionInfo <- function(x, ...) {
     vapply(X = x$otherPkgs, FUN = .extract_pkgref_packageDescription, FUN.VALUE = character(1), USE.NAMES = FALSE)
+}
+
+#' @rdname as_pkgrefs
+#' @export
+as_pkgrefs.renv_lockfile <- function(x, ...){
+  sources <- vapply(lockfile[["Packages"]],`[[`,character(1),"Source",USE.NAMES = FALSE)
+  pkgs <- c()
+  if("Repository"%in%sources){
+    pkgs <- c(pkgs, paste0("cran::",vapply(lockfile[["Packages"]][sources=="Repository"],`[[`,character(1),"Package",USE.NAMES = FALSE)))
+  }
+  if("Bioconductor"%in%sources){
+    pkgs <- c(pkgs,paste0("bioc::",vapply(lockfile[["Packages"]][sources=="Bioconductor"],`[[`,character(1),"Package",USE.NAMES = FALSE)))
+  }
+  if("GitHub"%in%sources){
+    pkgs <- c(pkgs,
+              paste0("github::",
+                     vapply(lockfile[["Packages"]][sources=="GitHub"],`[[`,character(1), "RemoteUsername", USE.NAMES = FALSE),"/",
+                     vapply(lockfile[["Packages"]][sources=="GitHub"],`[[`,character(1), "Package", USE.NAMES = FALSE))
+    )
+  }
+  pkgs
 }
 
 .extract_pkgref_packageDescription <- function(packageDescription) {
@@ -51,4 +75,25 @@ as_pkgrefs.sessionInfo <- function(x, ...) {
     ## if (basename(attr(packageDescription, "file")) == "package.rds") {
     return(paste0("cran::", handle))
     ## }
+}
+
+.detect_renv_lockfile <- function(path){
[…]

### 2023-02-24 R/resolve.R

@@ -199,7 +199,7 @@
 #'
 #' This function recursively queries dependencies of R packages at a specific snapshot time. The dependency graph can then be used to recreate the computational environment. The data on dependencies are provided by R-hub.
 #'
-#' @param pkgs `pkgs` can be 1) a character vector of R packages to resolve, or 2) a data structure that [as_pkgrefs()] can convert to a character vector of package references. For 1) `pkgs` can be either in shorthands, e.g. "rtoot", "ropensci/readODS", or in package references, e.g. "cran::rtoot", "github::ropensci/readODS". Please refer to the [Package References documentation](https://r-lib.github.io/pkgdepends/reference/pkg_refs.html) of `pak` for details. Currently, this package supports only cran and github packages. For 2) [as_pkgrefs()] support the output of [sessionInfo()].
+#' @param pkgs `pkgs` can be 1) a character vector of R packages to resolve, 2) a path to a [`renv` lockfile](https://rstudio.github.io/renv/articles/lockfile.html), or 3) a data structure that [as_pkgrefs()] can convert to a character vector of package references. For 1) `pkgs` can be either in shorthands, e.g. "rtoot", "ropensci/readODS", or in package references, e.g. "cran::rtoot", "github::ropensci/readODS". Please refer to the [Package References documentation](https://r-lib.github.io/pkgdepends/reference/pkg_refs.html) of `pak` for details. Currently, this package supports only cran and github packages. For 2) [as_pkgrefs()] support the output of [sessionInfo()].
 #' @param snapshot_date Snapshot date, if not specified, assume to be a month ago
 #' @param no_enhances logical, whether to ignore packages in the "Enhances" field
 #' @param no_suggests logical, whether to ignore packages in the "Suggests" field
@@ -248,6 +248,7 @@ resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE,
     if (snapshot_date >= anytime::anytime(Sys.Date())) {
         stop("We don't know the future.", call. = FALSE)
     }
+    
     if (class(pkgs) %in% c("sessionInfo")) {
         pkgrefs <- as_pkgrefs(pkgs)
         if(any(grepl("^bioc::",pkgrefs))){
@@ -256,6 +257,9 @@ resolve <- function(pkgs, snapshot_date, no_enhances = TRUE, no_suggests = TRUE,
           bioc_version <- NULL
         }
     } else {
[…]

### 2023-02-28 tests/testthat/test_resolve.R
Add tests for `.query_snapshot_dependecies` for local packages

@@ -283,3 +283,33 @@ test_that(".query_sysreqs_local", {
     expect_true("apt-get install -y libxml2-dev" %in% sysreqs)
     expect_true("apt-get install -y libbz2-dev" %in% sysreqs)
 })
+
+test_that(".query_snapshot_dependencies for local packages", {
+    skip_if_offline()
+    skip_on_cran()
+    skip_on_os("windows") ## don't want to be slabbed in the back by Windows' paths
+    expect_error(dep_df <- .query_snapshot_dependencies("local::../testdata/fakeRhtslib",
+                                                   snapshot_date = "2023-01-01", bioc_version = "3.3"), NA)
+    expect_true("y" %in% colnames(dep_df))
+    expect_true("bioc::zlibbioc" %in% dep_df$y_pkgref)
+    expect_true("cran::knitr" %in% dep_df$y_pkgref)
+    expect_equal("Rhtslib", unique(dep_df$x))
+    expect_true(grepl("^/", unique(dep_df$x_uid))) ## path expanded to abs. path
+    ## do the same thing but with tar.gz
+    expect_error(dep_df <- .query_snapshot_dependencies("local::../testdata/fakeRhtslib.tar.gz",
+                                                   snapshot_date = "2023-01-01", bioc_version = "3.3"), NA)
+    expect_true("y" %in% colnames(dep_df))
+    expect_equal(unique(dep_df$x_pubdate), parsedate::parse_date("2023-01-01"))
+    expect_true("bioc::zlibbioc" %in% dep_df$y_pkgref)
+    expect_true("cran::knitr" %in% dep_df$y_pkgref)
+    expect_equal("Rhtslib", unique(dep_df$x))
+    expect_true(grepl("^/", unique(dep_df$x_uid))) ## path expanded to abs. path
+    ## No y
+    expect_error(dep_df <- .query_snapshot_dependencies("local::../testdata/fakezlibbioc",
+                                                   snapshot_date = "2023-01-01", bioc_version = "3.3"), NA)
+    expect_false("y" %in% colnames(dep_df))
+    ## real data
+    expect_error(dep_df <- .query_snapshot_dependencies("local::../testdata/askpass_1.1.tar.gz",
+                                                   snapshot_date = "2023-01-01", bioc_version = "3.3"), NA)
+    expect_true("cran::sys" %in% dep_df$y_pkgref)
+})

### 2023-02-28 paper/paper.qmd
Update the abstract about Bioconductor

@@
 abstract: |
-  A complete description of the computational environment is often missing when researchers share their materials.
+  A complete declarative description of the computational environment is often missing when researchers share their materials.
 Without such description, software obsolescence and missing system components can jeopardize computational reproducibility in the future, even when data and computer code are available.
-The R package rang is a complete solution for generating instructions for other researchers to completely reconstruct the computational environment at a specific time point.
-The reconstruction process, based on Docker, has been tested for R code as old as 2005 and does not depend on the long term availability of any commericial service.
+The R package rang is a complete solution for generating the declarative description for other researchers to automatically reconstruct the computational environment at a specific time point.
+The reconstruction process, based on Docker, has been tested for R code as old as 2005 and does not depend on the long term availability of any commercial service.
 The description generated by rang satisfies the definition of a reproducible research compendium and can be shared as such.
+In this contribution, we show how rang can be used to make otherwise unexecutable code, spanning from fields such as computational social science and bioinformatics, executable again.
+We also provide instructions on how to use rang to construct reproducible and shareable research compendium of current research.
 keywords:
@@
 
-# Statement of need
+# Introduction
 
@@
 In the realm of R, the important aspects are what and which version of installed R packages as well as the exact R version.
+
+## Existing solutions
 
@@
 For instance, if there was a computational environment constructed on 2020-01-16 (called "snapshot date") with the several natural language processing R packages, `resolve()` can be used to resolve all the dependencies of these R packages.
-Currently, `rang` supports both CRAN and GitHub packages.
+Currently, `rang` supports CRAN, Bioconductor, and GitHub packages.
 
@@
 
-For the sake of completeness, the instructions for building and running the Docker container for Unix-like systems are included here.
+For the sake of completeness, the instructions for building and running the Docker container on Unix-like systems are included here.
 
@@
 
+## Recover a removed Bioconductor package
+
+# Generating a research compendium
+
+# Conclusion
+
 # References

### 2023-02-28 paper/paper.qmd
Further revise the abstract

@@
 ---
-title: 'rang: Reconstructing reproducible R computational environments with ease'
+title: 'rang: Reconstructing reproducible R computational environments'
 tags:
@@
 We also provide instructions on how to use rang to construct reproducible and shareable research compendium of current research.
+The package is currently available from CRAN (https://cran.r-project.org/web/packages/rang/index.html) and GitHub (https://github.com/chainsawriot/rang).
 keywords:
@@
 `resolve()` queries various web services from the r-hub project of the R Consortium for information about R packages at a specific time point that is necessary for reconstructing a computational environment, e.g.
-recursive dependencies, system requirements, and R version.
+(deep) dependencies, system requirements, and R version.
 For instance, if there was a computational environment constructed on 2020-01-16 (called "snapshot date") with the several natural language processing R packages, `resolve()` can be used to resolve all the dependencies of these R packages.
@@
 The resolved result is an S3 object called `rang`.
-The information contained in a `rang` can then be used to construct a computational environment in a similar manner as `containerit`, but with the packages and R versions pinned on the snapshot date.
-Then, `dockerize()` is used to generate the `Dockerfile` and other scripts in the `output_dir`.
+The information contained in a `rang` object can then be used to construct a computational environment in a similar manner as `containerit`, but with the packages and R versions pinned on the snapshot date.
+Then, the function `dockerize()` is used to generate the `Dockerfile` and other scripts in the `output_dir`.
+
+```r
+dockerize(graph, output_dir = "docker")
+```
+
 For R >= 3.1, the images from the Rocker project are used [@boettiger:2017:IR].
@@
 The development to support R < 2.1 is in progress.
-
-```r
-dockerize(graph, output_dir = "docker")
-```
-
 There are two features of `dockerize()` that are important for future reproducibility.
@@
 # Case Studies
+
+The following are some examples of how `rang` can be used to make otherwise shared, but unexecutable, R code runnable again.
+The examples were drawn from various fields spanning from political science, psychological science, and bioinformatics.
 

### 2023-02-28 tests/testthat/test_resolve.R
Add test for .query_sysreqs_smart dispatch with local packages

@@ -282,6 +282,9 @@ test_that(".query_sysreqs_local", {
     expect_error(sysreqs <- .query_sysreqs_local(c("../testdata/fakexml2", "../testdata/askpass_1.1.tar.gz", "../testdata/fakeRhtslib.tar.gz"), "ubuntu-20.04"), NA)
     expect_true("apt-get install -y libxml2-dev" %in% sysreqs)
     expect_true("apt-get install -y libbz2-dev" %in% sysreqs)
+    ## dispatch in .query_sysreqs_smart
+    expect_error(sysreqs2 <- .query_sysreqs_smart(c("local::../testdata/fakexml2", "local::../testdata/askpass_1.1.tar.gz", "local::../testdata/fakeRhtslib.tar.gz"), "ubuntu-20.04"), NA)
+    expect_equal(sysreqs, sysreqs2)
 })
 
 test_that(".query_snapshot_dependencies for local packages", {

### 2023-03-01 R/as_pkgrefs.R
Rearrange test_pkgref and add tests for renv.lock with local packages

@@ -55,17 +55,20 @@ as_pkgrefs.sessionInfo <- function(x, ...) {
     sources <- vapply(lockfile[["Packages"]],`[[`,character(1),"Source",USE.NAMES = FALSE)
     pkgs <- c()
     if("Repository" %in% sources) {
-      pkgs <- c(pkgs, paste0("cran::",vapply(lockfile[["Packages"]][sources=="Repository"],`[[`,character(1),"Package",USE.NAMES = FALSE)))
+        pkgs <- c(pkgs, paste0("cran::",vapply(lockfile[["Packages"]][sources=="Repository"],`[[`,character(1),"Package",USE.NAMES = FALSE)))
     }
     if("Bioconductor" %in% sources) {
-      pkgs <- c(pkgs,paste0("bioc::",vapply(lockfile[["Packages"]][sources=="Bioconductor"],`[[`,character(1),"Package",USE.NAMES = FALSE)))
+        pkgs <- c(pkgs,paste0("bioc::",vapply(lockfile[["Packages"]][sources=="Bioconductor"],`[[`,character(1),"Package",USE.NAMES = FALSE)))
     }
     if("GitHub" %in% sources) {
-      pkgs <- c(pkgs,
-                paste0("github::",
-                       vapply(lockfile[["Packages"]][sources=="GitHub"],`[[`,character(1), "RemoteUsername", USE.NAMES = FALSE),"/",
-                       vapply(lockfile[["Packages"]][sources=="GitHub"],`[[`,character(1), "Package", USE.NAMES = FALSE))
-      )
+        pkgs <- c(pkgs,
+                  paste0("github::",
+                         vapply(lockfile[["Packages"]][sources=="GitHub"],`[[`,character(1), "RemoteUsername", USE.NAMES = FALSE),"/",
+                         vapply(lockfile[["Packages"]][sources=="GitHub"],`[[`,character(1), "Package", USE.NAMES = FALSE))
+                  )
+    }
+    if ("Local" %in% sources) {
+        pkgs <- c(pkgs, paste0("local::", vapply(lockfile[["Packages"]][sources=="Local"],`[[`,character(1),"RemoteUrl",USE.NAMES = FALSE)))
     }
     return(pkgs)
 }

### 2023-03-01 tests/testthat/test_pkgref.R

@@ -1,5 +1,7 @@
 ## using all the cases in https://r-lib.github.io/pkgdepends/reference/pkg_refs.html
 
+## .normalize_pkgs aka as_pkgrefs
+
 test_that(".normalize_pkgs, defensive programming", {
     expect_error(.normalize_pkgs(c("forecast", "r-lib/crayon")), NA)
     expect_error(.normalize_pkgs(NULL), NA)
@@ -52,6 +54,27 @@ test_that(".normalize_pkgs: github remote string", {
     expect_equal(.normalize_pkgs("github::git@github.com:r-lib/pak.git"), "github::r-lib/pak")
 })
 
+test_that(".normalize_pkgs: local", {
+    expect_equal(.normalize_pkgs("local::/foo/bar/package_1.0.0.tar.gz"), "local::/foo/bar/package_1.0.0.tar.gz")
+    expect_equal(.normalize_pkgs("local::/foo/bar/pkg"), "local::/foo/bar/pkg")
+    expect_equal(.normalize_pkgs("local::."), "local::.")
+    expect_equal(.normalize_pkgs("/absolute/path/package_1.0.0.tar.gz"), "local::/absolute/path/package_1.0.0.tar.gz")
+    expect_equal(.normalize_pkgs("~/path/from/home"), "local::~/path/from/home")
+    expect_equal(.normalize_pkgs("./relative/path"), "local::./relative/path")
+    expect_equal(.normalize_pkgs("."), "local::.")
+})
+
+test_that("as_pkgrefs dispatch", {
+    expect_error(as_pkgrefs(TRUE))
+    expect_error(as_pkgrefs(7.21))
+    expect_error(as_pkgrefs(1L))
+    expect_equal(as_pkgrefs("rtoot"), "cran::rtoot")
+    expect_equal(as_pkgrefs(c("rtoot", "sna")), c("cran::rtoot", "cran::sna"))
+    expect_equal(as_pkgrefs(c("rtoot", "S4Vectors")), c("cran::rtoot", "cran::S4Vectors")) ## the bioc version is in test_resolve
+})
+
+## .parse_pkgref
+
 test_that(".parse_pkgref", {
     expect_error(.parse_pkgref("withr"))
     expect_error(.parse_pkgref("r-lib/withr"))
@@ -67,21 +90,14 @@ test_that(".parse_pkgref", {
     expect_equal(.parse_pkgref("local::./relative/path", FALSE), "local")
 })
 
+## as_pkgrefs.sessionInfo()
+
[…]

### 2023-03-01 R/installation.R
add warning in comments in rang.R when local packages are found

@@ -129,7 +129,11 @@
 }
 
 .write_rang_as_comment <- function(rang, con, path, verbose, lib,
-                                       cran_mirror, check_cran_mirror, bioc_mirror) {
+                                   cran_mirror, check_cran_mirror, bioc_mirror) {
+    if (isTRUE(any(grepl("^local::", .extract_pkgrefs(rang))))) {
+        cat("## ## WARNING:", file = con)
+        cat("## ## Local packages found. The following instructions are not reproducible.", file = con)
+    }
     cat("## ## To reconstruct this file, please install version",
         as.character(utils::packageVersion("rang")), "of `rang` and run:\n", file = con)
     cat("## rang <- \n", file = con)

### 2023-03-01 tests/testthat/test_dockerize.R

@@ -20,6 +20,7 @@ test_that("integration of #13 in dockerize()", {
     dockerize(rang = rang_ok, output_dir = temp_dir) ## rang_as_comment = TRUE
     x <- readLines(file.path(temp_dir, "rang.R"))
     expect_true(any(grepl("^## ## To reconstruct this file", x)))
+    expect_false(any(grepl("^## ## WARNING", x)))
     dockerize(rang = rang_ok, output_dir = temp_dir, rang_as_comment = FALSE)
     x <- readLines(file.path(temp_dir, "rang.R"))
     expect_false(any(grepl("^## ## To reconstruct this file", x)))

### 2023-03-01 tests/testthat/test_resolve.R

@@ -339,4 +339,6 @@ test_that("dockerize local package as tarball", {
     expect_error(dockerize(graph, output_dir = temp_dir, cache = TRUE, verbose = FALSE), NA) ## cache = FALSE
     expect_true(file.exists(file.path(temp_dir, "cache", "sys_3.4.1.tar.gz")))
     expect_true(dir.exists(file.path(temp_dir, "cache", "dir_askpass_1.1")))
+    x <- readLines(file.path(temp_dir, "rang.R"))
+    expect_true(any(grepl("^## ## WARNING", x)))
 })

### 2023-03-01 tests/testthat/test_dockerize.R
Add unittests for `.check_tarball_path`

@@ -238,3 +238,10 @@ test_that("no_rocker #67", {
     expect_error(dockerize(rang = rang_ok, output_dir = temp_dir, no_rocker = TRUE,
               debian_version = "3.11"))
 })
+
+test_that(".check_tarball_path", {
+    expect_error(.check_tarball_path("../testdata/gesis_2.0.tar.gz", "gesis")) ##dir = FALSE
+    expect_error(.check_tarball_path("../testdata/askpass_1.1.tar.gz", "askpass"), NA)
+    expect_error(.check_tarball_path("../testdata/gesis", "gesis", dir = TRUE))
+    expect_error(.check_tarball_path("../testdata/askpass", "askpass", dir = TRUE), NA)
+})

### 2023-03-01 R/installation.R
Update doc and vignette

@@ -419,7 +419,7 @@ export_rang <- function(rang, path, rang_as_comment = TRUE, verbose = TRUE, lib 
 #' @param output_dir character, where to put the Docker file and associated content
 #' @param materials_dir character, path to the directory containing additional resources (e.g. analysis scripts) to be copied into `output_dir` and in turn into the Docker container
 #' @param image character, which versioned Rocker image to use. Can only be "r-ver", "rstudio", "tidyverse", "verse", "geospatial"
-#' This applies only to R version <= 3.1
+#' This applies only to R version >= 3.1
 #' @param cache logical, whether to cache the packages now. Please note that the system requirements are not cached. For query with non-CRAN packages, this option is strongly recommended. For query with local packages, this must be TRUE regardless of R version. For R version < 3.1, this must be also TRUE if there is any non-CRAN packages.
 #' @param no_rocker logical, whether to skip using Rocker images even when an appropriate version is available. Please keep this as `TRUE` unless you know what you are doing
 #' @param debian_version, when Rocker images are not used, which EOL version of Debian to use. Can only be "lenny", "etch", "squeeze", "wheezy", "jessie", "stretch". Please keep this as default "lenny" unless you know what you are doing

### 2023-03-02 paper/paper.qmd
Update the paper

@@
 
+Similar to CRAN, Bioconductor is a series of repositories of R package for bioinformatics and computational biology. Also similar to CRAN, packages can get removed over time.
+
+The Bioconductor package `Sushi` has been deprecated by the original authors and is removed from Bioconductor version 3.16 (2022-11-02).
+It is a data visualization tool for genomic data and was used in various online tutorials and papers, including the original paper announcing the package by the original authors [@phanstiel:2014:S].
+
 # Generating a research compendium

### 2023-03-02 R/pkgref.R
Fix #102

@@ -52,7 +52,16 @@
     return(source)
 }
 
+.is_local <- function(pkg) {
+    ## according to the standard, it must be started by ".", "~", "/"
+    grepl("^[\\.~/]", pkg)
+}
+
 .is_github <- function(pkg) {
+    ## make .is_local precedes .is_github
+    if (isTRUE(.is_local(pkg))) {
+        return(FALSE)
+    }
     if (grepl("github\\.com", pkg)) {
         return(TRUE)
     }
@@ -67,12 +76,6 @@
     }
     bioc_pkgs <- .memo_search_bioc(bioc_version)
     pkg %in% bioc_pkgs$Package
-}
-
-
-.is_local <- function(pkg) {
-    ## according to the standard, it must be started by ".", "~", "/"
-    grepl("^[\\.~/]", pkg)
 }
 
 ## TBI: .is_valid_pkgref
@@ -98,10 +101,8 @@
     if (pkg == "" || is.na(pkg)) {
         stop("Invalid `pkg`.", call. = FALSE)
     }
-    if (isTRUE(.is_github(pkg))) {
-        if (isTRUE(grepl("github\\.com", pkg))) {
-          pkg <- .extract_github_handle(pkg)
-        }
+    if (isTRUE(grepl("github\\.com", pkg))) {
+        pkg <- .extract_github_handle(pkg)
     }
     if (isTRUE(.is_pkgref(pkg))) {
         return(.clean_suffixes(pkg))

### 2023-03-02 tests/testthat/test_pkgref.R

@@ -169,3 +169,8 @@ test_that(".is_local", {
     expect_true(.is_local("/hello/world/"))
     expect_true(.is_local("../testdata/fakexml2"))
 })
+
+test_that(".is_local precedes .is_github", {
+    expect_false(.is_github("~/helloworld"))
+    expect_false(.is_github("./helloworld"))
+})

### 2023-03-02 tests/testthat/test_resolve.R

@@ -342,3 +342,11 @@ test_that("dockerize local package as tarball", {
     x <- readLines(file.path(temp_dir, "rang.R"))
     expect_true(any(grepl("^## ## WARNING", x)))
 })
+
+test_that("issue 102 confusion between github and local pkgref", {
+    original_wd <- getwd()
+    setwd(normalizePath(file.path("../testdata")))
+    expect_error(suppressWarnings(graph <- resolve("local::./askpass", snapshot_date = "2023-01-01")), NA)
+    expect_equal(.parse_pkgref(graph$ranglets[[1]]$original$x_pkgref, FALSE), "local")
+    setwd(original_wd)
+})

### 2023-03-06 R/installation.R
Introduce skip_r17

@@ -447,7 +447,8 @@ export_renv <- function(rang, path = ".") {
 #' This applies only to R version >= 3.1
 #' @param cache logical, whether to cache the packages now. Please note that the system requirements are not cached. For query with non-CRAN packages, this option is strongly recommended. For query with local packages, this must be TRUE regardless of R version. For R version < 3.1, this must be also TRUE if there is any non-CRAN packages.
 #' @param no_rocker logical, whether to skip using Rocker images even when an appropriate version is available. Please keep this as `TRUE` unless you know what you are doing
-#' @param debian_version, when Rocker images are not used, which EOL version of Debian to use. Can only be "lenny", "etch", "squeeze", "wheezy", "jessie", "stretch". Please keep this as default "lenny" unless you know what you are doing
+#' @param debian_version when Rocker images are not used, which EOL version of Debian to use. Can only be "lenny", "etch", "squeeze", "wheezy", "jessie", "stretch". Please keep this as default "lenny" unless you know what you are doing
+#' @param skip_r17 logical, whether to skip R 1.7.x. Currently, it is not possible to compile R 1.7.x (R 1.7.0 and R 1.7.1) with the method provided by  `rang`. It affects `snapshot_date` from 2003-04-16 to 2003-10-07. When `skip_r17` is TRUE and `snapshot_date` is within the aforementioned range, R 1.8.0 is used instead.
 #' @param ... arguments to be passed to `dockerize`
 #' @return `output_dir`, invisibly
 #' @inheritParams export_rang
@@ -470,7 +471,8 @@ dockerize <- function(rang, output_dir, materials_dir = NULL, image = c("r-ver",
                       cran_mirror = "https://cran.r-project.org/", check_cran_mirror = TRUE,
                       bioc_mirror = "https://bioconductor.org/packages/",
                       no_rocker = FALSE,
-                      debian_version = c("lenny", "squeeze", "wheezy", "jessie", "stretch")) {
+                      debian_version = c("lenny", "squeeze", "wheezy", "jessie", "stretch"),
+                      skip_r17 = TRUE) {
     if (length(rang$ranglets) == 0) {
         warning("Nothing to dockerize.")
         return(invisible(NULL))
@@ -509,16 +511,20 @@ dockerize <- function(rang, output_dir, materials_dir = NULL, image = c("r-ver",
     if (isTRUE(cache)) {
[…]

### 2023-03-06 tests/testthat/test_resolve.R

@@ -318,6 +318,21 @@ test_that("dockerize R1.3.1 local", {
     expect_true(any(grepl("^## DEBUG INFO: CMD", x)))
 })
 
+test_that("skip_r17", {
+    skip_if_offline()
+    skip_on_cran()
+    expect_error(suppressWarnings(graph <- resolve("../testdata/sna_0.3.tar.gz", snapshot_date = "2003-04-17")), NA)
+    expect_equal(graph$r_version, "1.7.0")
+    temp_dir <- .generate_temp_dir()
+    expect_error(dockerize(graph, output_dir = temp_dir, cache = TRUE, verbose = FALSE), NA) ## skip_r17 = TRUE
+    x <- readLines(file.path(temp_dir, "Dockerfile"))
+    expect_true(any(grepl("^RUN bash compile_r\\.sh 1\\.8\\.0", x)))
+    expect_error(dockerize(graph, output_dir = temp_dir, cache = TRUE, verbose = FALSE, skip_r17 = FALSE), NA)
+    x <- readLines(file.path(temp_dir, "Dockerfile"))
+    expect_false(any(grepl("^RUN bash compile_r\\.sh 1\\.8\\.0", x)))
+    expect_true(any(grepl("^RUN bash compile_r\\.sh 1\\.7\\.0", x)))
+})
+
 ## always keep this at the very last
 
 test_that("issue 102 confusion between github and local pkgref", {

### 2023-03-06 README.Rmd
Add CRAN post-acceptance things [no CI]
Fix README about Woody Ref #67 [no ci]
Update doc regarding R 1.3.1 support [no ci]

@@
 remotes::install_github("chainsawriot/rang")
+```
+
+Or the stable CRAN version
+
+```r
+install.packages("rang")
 ```
@@
 `rang` can still be used to recreate computational environments for R < 3.1.0.
-The Dockerfile generated is based on Debian Woody (3.0) and the requested version of R is compiled from source.
-As of writing, this method works for R < 3.1.0 but not R < 2.1.0.
+The Dockerfile generated is based on Debian Lenny (5.0) and the requested version of R is compiled from source.
+As of writing, this method works for R < 3.1.0 but not R < 1.3.1.
 The `image` parameter is ignored in this case.

### 2023-03-06 paper/paper.qmd
in

@@
 One can think of `rang` as an archaeological tool.
-In the following examples, `rang` is used to enable the reproducibility of published literature.
+In Section 3, `rang` is used to enable the reproducibility of published literature.
 
@@
 
-## Recover a removed R package
+## Recover a removed R package: maxent
 
@@
 
+## Recover a removed R package: ptproc
+
+The R package `ptproc` was published as a paper in 2003 [@peng:2003:MDP].
+But the package has been removed from CRAN for over a decade (at least).
+The only release on CRAN was on 2002-10-10.
+The package is still listed in the "Handling and Analyzing Spatio-Temporal Data" CRAN Task View.
+As of writing, the package is still downloadable from the original author's website [^peng].
+
+Despite this over-a-decade removal, there is evidence that the package is still being sought for. As late as 2017, there are blog posts on how to install the long obsolete package on modern R [^blog].
+
+[^peng]: https://www.biostat.jhsph.edu/~rpeng/software/
+
+[^blog]: https://blog.mathandpencil.com/installing-ptproc-on-osx and https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/
+
 ## Recover a removed Bioconductor package
@@
 The Bioconductor package `Sushi` has been deprecated by the original authors and is removed from Bioconductor version 3.16 (2022-11-02).
-It is a data visualization tool for genomic data and was used in various online tutorials and papers, including the original paper announcing the package by the original authors [@phanstiel:2014:S].
+It is a data visualization tool for genomic data and was used in many online tutorials and papers, including the original paper announcing the package by the original authors [@phanstiel:2014:S].
 
-# Generating a research compendium
+# Producing a research compendium
 

### 2023-03-07 paper/paper.qmd
Add peng example

@@
 This R script has been "rescued" by the author of the R package `groundhog` [@groundhogrpkg], as demonstrated in a blog post [^groundhog].
-We were wondering if `rang` can also be used to rescue the concerned R script.
-
-The date of the R script, as indicated on Dataverse, is 2018-12-12. This date is used as the snapshot date.
+We were wondering if `rang` can also be used to "rescue" the concerned R script.
+The date of the R script, as indicated on Dataverse, is 2018-12-12.
+This date is used as the snapshot date.
 
@@
 @loercher:2017:D].
-
 Despite the covert editing of the package by a staffer of CRAN [^evidence], the package was removed from CRAN in 2019 [^checkerror].
@@
 
-[^speed]: On an interesting historical note: The original paper reported, based on a benchmark, that "the algorithm is very fast; `maxent` uses only 135.4 megabytes of RAM and finishes in 53.3 seconds." On a modest computer in 2023 with a dockerized R 2.15.0, the benchmark finishes in 4 seconds.
+[^speed]: On an interesting historical side note: The original paper reported ---based on a benchmark--- that "the algorithm is very fast; `maxent` uses only 135.4 megabytes of RAM and finishes in 53.3 seconds." On a modest computer in 2023 with a dockerized R 2.15.0, the benchmark finishes in 4 seconds.
 
@@
 
-The R package `ptproc` was published as a paper in 2003 [@peng:2003:MDP].
+The software paper of the R package `ptproc` was published in 2003 and introduced multidimensional point process models [@peng:2003:MDP].
 But the package has been removed from CRAN for over a decade (at least).
 The only release on CRAN was on 2002-10-10.
-The package is still listed in the "Handling and Analyzing Spatio-Temporal Data" CRAN Task View.
-As of writing, the package is still downloadable from the original author's website [^peng].
+The package is still listed in the "Handling and Analyzing Spatio-Temporal Data" CRAN Task View despite being uninstallable without modification on any modern R system (see below).
+As of writing, the package, as a tarball file (tar.gz), is still downloadable from the original author's website [^peng].
 
-Despite this over-a-decade removal, there is evidence that the package is still being sought for. As late as 2017, there are blog posts on how to install the long obsolete package on modern R [^blog].
+Even with this over-a-decade removal and new packages with similar functionalities have been created, there is evidence that `ptproc` is still being sought for.
+As late as 2017, there are blog posts on how to install the long obsolete package on modern R [^blog].
+The package is extremely challenging to install on a modern R system because the package was written before the introduction of name space management in R 1.7.0 [@RN-2003-001].
+In other words, the available tarball file from the original author's website does not contain a `NAMESPACE` file as all other modern R packages do.
+
+The oldest version of R that `rang` can support, as of writing, is R 1.3.1.
+`rang` is probably the only solution available that can support the 1.x series of R (i.e.
+before 2004-10-04).
+Similar to the case of `maxent` above, a Docker image with the available tarball file of `ptproc` can be generated with two lines of code.
+
+```r
+graph <- resolve("ptproc", snapshot_date = "2004-07-01")
+dockerize(graph, "~/dev/misc/ptproc", cache = TRUE)
+```
+
+Suppose we have an R script called "peng.R" like this (extracted from @peng:2003:MDP)
+
+```r
+require(ptproc)
+
+set.seed(1000)
+x <- cbind(runif(100), runif(100), runif(100))
+hPois.cond.int <- function(params, eval.pts, pts = NULL, data = NULL, TT = NULL) {
+    mu <- params[1]
+    if(is.null(TT))
+        rep(mu, nrow(eval.pts))
+    else {
+        vol <- prod(apply(TT, 2, diff))
+        mu * vol
+    }
+}
+ppm <- ptproc(pts = x, cond.int = hPois.cond.int, params = 50,
+              ranges = cbind(c(0,1), c(0,1), c(0,1)))
+fit <- ptproc.fit(ppm, optim.control = list(trace = 2), method = "BFGS")
+summary(fit)
+```
+
+One can integrate `rang` into a BASH script to completely automate the batch execution of the above R script.
+
+```bash
+Rscript -e "require(rang); dockerize(resolve('ptproc', '2004-07-01'),
+'pengdocker', cache = TRUE)"
+docker build -t pengimg ./pengdocker
+docker run -d --rm --name "pengcontainer" -ti pengimg
+docker cp peng.R pengcontainer:/peng.R
+docker exec pengcontainer R CMD BATCH peng.R
+docker exec pengcontainer cat peng.Rout
+docker cp pengcontainer:/peng.Rout peng.Rout
+docker stop pengcontainer
+```
+
+The file `peng.Rout` contains the execution results of the script from inside the Docker container.
 
@@
 The Bioconductor package `Sushi` has been deprecated by the original authors and is removed from Bioconductor version 3.16 (2022-11-02).
-It is a data visualization tool for genomic data and was used in many online tutorials and papers, including the original paper announcing the package by the original authors [@phanstiel:2014:S].
+It is a data visualization tool for genomic data and was used in many online tutorials and scientific papers, including the original paper announcing the package by the original authors [@phanstiel:2014:S].
 
-# Producing a research compendium
+# Preparing research compendia with long-term computational reproducibility
 

### 2023-03-07 paper/peng.sh

@@ -1,0 +1,8 @@
+Rscript -e "devtools::load_all(); dockerize(resolve('ptproc', '2004-07-01'), 'pengdocker', cache = TRUE)"
+docker build -t pengimg ./pengdocker
+docker run -d --rm --name "pengcontainer" -ti pengimg
+docker cp peng.R pengcontainer:/peng.R
+docker exec pengcontainer R CMD BATCH peng.R
+docker exec pengcontainer cat peng.Rout
+docker cp pengcontainer:/peng.Rout peng.Rout
+docker stop pengcontainer

### 2023-03-07 paper/paper.qmd
Add all examples and compendium

@@
 
-The above code snippet can be executed with the generated container without any problem.
+The above code snippet can be executed with the generated container without any problem (@fig-figure1).
 
@@
 #| echo: FALSE
+#| label: fig-figure1
 knitr::include_graphics("quanteda_rstudio.png", dpi = 300)
@@
 The only release on CRAN was on 2002-10-10.
-The package is still listed in the "Handling and Analyzing Spatio-Temporal Data" CRAN Task View despite being uninstallable without modification on any modern R system (see below).
+The package is still listed in the "Handling and Analyzing Spatio-Temporal Data" CRAN Task View [^taskview] despite being uninstallable without modification on any modern R system (see below).
 As of writing, the package, as a tarball file (tar.gz), is still downloadable from the original author's website [^peng].
+
+[^taskview]: https://cran.r-project.org/web/views/SpatioTemporal.html
 
@@
 before 2004-10-04).
-Similar to the case of `maxent` above, a Docker image with the available tarball file of `ptproc` can be generated with two lines of code.
+Similar to the case of `maxent` above, a Dockerfile to generate a Docker image with `ptproc` installed can be generated with two lines of code.
 
@@
 
-Suppose we have an R script called "peng.R" like this (extracted from @peng:2003:MDP)
+Suppose we have an R script, extracted from @peng:2003:MDP, called "peng.R" like this:
 
@@
 docker build -t pengimg ./pengdocker
+## launching a container in daemon mode -d
 docker run -d --rm --name "pengcontainer" -ti pengimg
@@
 The file `peng.Rout` contains the execution results of the script from inside the Docker container.
+As the random seed was preserved by the original author [@peng:2003:MDP], the above BASH script can perfectly reproduce the analysis [^random].
 
@@
 [^blog]: https://blog.mathandpencil.com/installing-ptproc-on-osx and https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/
+
+[^random]: It is also important to note that the random number generator (RNG) of R has been changed several times over the course of the development.
+In this case, we are using the same generation of RNG as @peng:2003:MDP.
 
@@
 The Bioconductor package `Sushi` has been deprecated by the original authors and is removed from Bioconductor version 3.16 (2022-11-02).
-It is a data visualization tool for genomic data and was used in many online tutorials and scientific papers, including the original paper announcing the package by the original authors [@phanstiel:2014:S].
+`Sushi` is a data visualization tool for genomic data and was used in many online tutorials and scientific papers, including the original paper announcing the package by the original authors [@phanstiel:2014:S].
+
+`rang` has native support for Bioconductor packages since 0.2.
+We obtained the R script `"PaperFigure.R"` from the Github repo of `Sushi` [^sushi], which generates the figure in @phanstiel:2014:S.
+Similar to the above case of `ptproc`, we made a completely automated BASH script to run `"PaperFigure.R"` and get the generated figure out of the container (@fig-figure2).
+We made no modification to `"PaperFigure.R"`.
+
+```sh
+Rscript -e "require(rang); dockerize(resolve('Sushi', '2014-06-05'),
+'sushidocker', no_rocker = TRUE, cache = TRUE)"
+docker build -t sushiimg ./sushidocker
+docker run -d --rm --name "sushicontainer" -ti sushiimg
+docker cp PaperFigure.R sushicontainer:/PaperFigure.R
+docker exec sushicontainer mkdir vignettes
+docker exec sushicontainer R CMD BATCH PaperFigure.R
+docker cp sushicontainer:/vignettes/Figure_1.pdf sushi_figure1.pdf
+docker stop sushicontainer
+```
+
+```{r}
+#| fig.cap: The figure from the batch execution of `PaperFigure.R` inside a Docker container generated by `rang`
+#| echo: FALSE
+#| label: fig-figure2
+knitr::include_graphics("sushi_figure1.pdf", dpi = 300)
+```
+
+[^sushi]: https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R
 
 # Preparing research compendia with long-term computational reproducibility
+
+The above six examples show how useful it is for `rang` reconstructing tricky computational environments which have not been completely declared in the literature.
+Although we position `rang` mostly as an archaeological tool, we also think that `rang` can also be used to prepare research compendia of current research.
+We can't predict the future but research compendia generated by `rang` would probably have long-term computational reproducibility.
+
+To demonstrate this point, we took the recent paper by @oser:2022:HPE.
+This paper was selected because 1) the paper was published in *Political Communication*, a high impact journal that awards Open Science Badges; 2) shared data and R code are available; and most importantly, 3) the shared R code is well-written.
+In the repository of this paper, we based on the materials shared by @oser:2022:HPE and prepared a research compendium that should have long-term computational reproducibility.
+The research compendium is similar to the Executable Compendium suggested by the Turing way.
+
+The preparation of the research compendium is easy as `rang` can scan a materials directory for all R packages used [^dmetar].
+
+```r
+require(rang)
+## meta-analysis is the directory of all shared materials
+cran_pkgs <- as_pkgrefs("meta-analysis")
+
+## dmetar is an undeclared github package: MathiasHarrer/dmetar
+cran_pkgs[cran_pkgs == "cran::dmetar"] <- "MathiasHarrer/dmetar"
+x <- resolve(cran_pkgs, "2021-08-11", verbose = TRUE)
+print(x, all_pkgs = TRUE)
+dockerize(x, "oser_docker", materials_dir = "meta-analysis", cache = TRUE)
+```
+
+The above R script is saved as `oser.R`. The central piece of the executable compendium is the `Makefile`.
+
+```Makefile
+rmd = "rmarkdown::render('materials/README.Rmd', output_file = 'reproduced.html')"
+resolve:
+	Rscript oser.R
+build: oser_docker
+	docker build -t oserimg oser_docker
+render:
+	docker run -d --rm --name "osercontainer" -ti oserimg
+	docker exec osercontainer Rscript -e ${rmd}
+	docker cp osercontainer:/materials/reproduced.html oser_README.html
+	docker stop osercontainer
+export:
+	docker save oserimg | gzip > oserimg.tar.gz
+rebuild: oserimg.tar.gz
+	docker load < oserimg.tar.gz
+all: resolve build render
+	echo "finished"
+```
+
+With this `Makefile`, once can create the Dockerfile with `make resolve`, build the Docker image with `make build`, render the RMarkdown file inside the container with `make render`, export the built Docker image with `make export`, and rebuild the exported Docker image with `make rebuild`.
+
+The structure of the executable compendium looks like this:
+
+```bash
+Makefile
+oser.R
+meta-analysis/
+README.md
+oser_docker/
+oserimg.tar.gz
+```
+
+In this executable compendium, only the first four elements are essential.
+The directory `oser_docker` (116 MB) contains cached R packages, Dockerfile, and a further copy of the directory `meta-analysis/` to be transferred into the Docker image.
+That can be regenerated by running `make resolve`.
+However, having this directory preserved insures against the situations of some R packages used in the project are no longer available or any of the information providers used by `rang` for resolving the dependency relationships being not available.
+(Or in the rare circumstance of `rang` is no longer available.)
+
+`oserimg.tar.gz` (667 MB) is a backup copy of the Docker image.
+This can be regenerated by running `make export`.
+Preserving this file insurces against all the situations mentioned above, but also the situations of Docker Hub and the software repositories used by the dockerized operating system being not available.
+As a matter of fact, when `oserimg.tar.gz` is available it is possible to run `make rebuild` and `make render` even without internet access, provided that only Docker and make have been preinstalled.
+Of course, there is still an extremely rare situation where Docker (the program) itself is no longer available.
+However, it is possible to convert the image file for use on other containerization solutions such as Singularity [^singularity].
+
+Sharing of research artefacts less than 1G is not as challenging as it used to be.
+Zenodo, for example, allows the sharing of 50G of files.
+Therefore, sharing of the last two components of the executable compendium prepared with `rang` is at least possible on Zenodo.
+However, for data repositories with more restrictions on data size, sharing the executable compendium without the last two parts could be considered sufficient.
+For that, run `make all` will generate all the things needed for reproducing the analysis inside a container.
+
+[^singularity]: https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html
+
+[^dmetar]: We detected a minor issue in the code base that an undeclared Github package is used. But it can be easily solved, as in the Psychological Science example above.
 

### 2023-03-07 paper/PaperFigure.R

@@ -1,0 +1,471 @@
+library('Sushi')
+pdfname = "vignettes/Figure_1.pdf"
+Sushi_data = data(package = 'Sushi')
+data(list = Sushi_data$results[,3]) 
+makepdf = TRUE
+
+###
+### CODE
+###
+
+if (makepdf == TRUE)
+{
+  pdf(pdfname, height=10, width=12)
+}
+
+# make a layout for all of the plots
+layout(matrix(c(1,1,1,1,
+                1,1,1,1,
+                2,2,8,8,
+                2,2,9,9,
+                3,3,10,10,
+                3,3,10,10,
+                4,4,11,11,
+                4,4,11,11,
+                5,5,12,12,
+                5,5,12,12,
+                6,7,13,13,
+                6,7,14,14
+), 12, 4, byrow=TRUE))
+par(mgp=c(3,.3,0))
+
+
+###
+### (A) manhattan plot
+###
+
+# set the margins
+par(mar=c(3,4,3,2))
+
+# set the genomic regions
+chrom1            = "chr11"
+chromstart1       = 500000
+chromend1         = 5050000
+
+chrom2            = "chr15"
+chromstart2       = 73000000
+chromend2         = 89500000
+
+# make the manhattan plot
+plotManhattan(bedfile=Sushi_GWAS.bed, pvalues=Sushi_GWAS.bed[,5], genome=Sushi_hg18_genome, cex=0.75)
+
+# add zoom 1
+zoomsregion(region=c(chromstart1,chromend1), chrom=chrom1, genome=Sushi_hg18_genome, extend=c(0.07,0.2), wideextend=0.2, offsets=c(0,.535))
+
+# add zoom 2
+zoomsregion(region=c(chromstart2,chromend2), chrom=chrom2, genome=Sushi_hg18_genome, extend=c(0.07,0.2), wideextend=0.2, offsets=c(.535,0))
+
+# add labels
+labelgenome(genome=Sushi_hg18_genome, n=4, scale="Mb", edgeblankfraction=0.20)
+
+# add y-axis
+axis(side=2, las=2, tcl=.2)
+mtext("log10(P)", side=2, line=1.75, cex=.75, font=2)
+
+# Add plot label
+labelplot("A)", " GWAS", letteradj=-.025)
+
+
+###
+### (B) Hi-C
+###
+
+# set the margins
+par(mar=c(3,4,2,2))
+
+# set the genomic regions
+chrom            = "chr11"
[…]

### 2023-03-07 paper/oser/oser.R

@@ -1,0 +1,9 @@
+require(rang)
+
+## all packages used in the project
+cran_pkgs <- as_pkgrefs("meta-analysis") ## dmetar is an undeclared github package: MathiasHarrer/dmetar
+
+cran_pkgs[cran_pkgs == "cran::dmetar"] <- "MathiasHarrer/dmetar"
+x <- resolve(cran_pkgs, "2021-08-11", verbose = TRUE)
+print(x, all_pkgs = TRUE)
+dockerize(x, "oser_docker", materials_dir = "meta-analysis", cache = TRUE)

### 2023-03-07 paper/sushi.sh

@@ -1,0 +1,9 @@
+Rscript -e "devtools::load_all(); dockerize(resolve('Sushi', '2014-06-05'),
+'sushidocker', no_rocker = TRUE, cache = TRUE)"
+docker build -t sushiimg ./sushidocker
+docker run -d --rm --name "sushicontainer" -ti sushiimg
+docker cp PaperFigure.R sushicontainer:/PaperFigure.R
+docker exec sushicontainer mkdir vignettes
+docker exec sushicontainer R CMD BATCH PaperFigure.R
+docker cp sushicontainer:/vignettes/Figure_1.pdf sushi_figure1.pdf
+docker stop sushicontainer

### 2023-03-07 paper/paper.qmd
Clean up to reflect the upcoming v0.2

@@
 
-# Introduction
+# Background
 
-The ability to reconstruct the computational environment is one of the most important aspects of reproducibility.
-In the realm of R, the important aspects are what and which version of installed R packages as well as the exact R version.
+*"In some cases the polarization estimation will not work ...
+This is _NOT_ a problem in the method, it is entirely dependent on the numpy version (and even the OS's).
+If you have different versions of numpy or even the same version of numpy on a different OS configuration, different networks will fail randomly.
+...
+[F]or instance, the 109th Congress will fail, but will work entirely normally on a different numpy version, which will fail on a different Congress network."*
+
+> ~ excerpt of [this README file](https://www.michelecoscia.com/?page_id=2105)
+
+Other than bad programming practices [@trisovic:2022], the main computing barrier to computational reproducibility is the failure to reconstruct the computational environment like the one used by the original researchers.
+This task looks trivially simple.
+But as computer science research has shown, this task is incredibly complex [@abate:2015:MCR;@dolstra:2010:N].
+In the realm of R, that includes at least four aspects: a) operating system, b) system components such as `libxml2`, c) the exact R version, and d) what and which version of the installed R packages.
+We will call them component a, b, c, d in the following sections.
+Any change in these four components can possibly affect the execution of any shared computer code.
+For example, the lack of the system component `libxml2` can impact whether the R package `xml2` can be installed on a Linux system.
+If the shared computer code requires the R package `xml2`, then the whole execution fails.
+
+In reality, the impact of operating system is relative weak as mainstream, open source programming languages and their software libraries are usually cross platform.
+In modern computational research, Linux is the de-facto operating system in high performance computing environments.
+Instead, the impact of components b, c, and d is practically higher.
+Component d is probably the most volatile among them all.
+Software update with breaking changes might render existing shared code not executable anymore.
+Also, software obsolescence is a common place, especially when academic software is often not well maintained [@merow:2023:B].
+
+The DevOps (software development and IT operations) community has come up with a partial solution to this problem: containerization.
+In essence, to containerize is to deploy the software together with all the libraries and even the operating system in a virtualized environment.
+In this way, software dependency issues can be resolved within the tightly controlled virtualized environment and independent of what is installed on the host computer.
 
@@
 But `containerit` does not pin down the exact version of R packages.
-Other solutions such as `groundhog` [@groundhogrpkg] and `checkpoint` [@checkpointrpkg] depend on the availability of The Microsoft R Application Network (MRAN), which will be shut down on July 1st, 2023.
+`checkpoint` [@checkpointrpkg] depends on the availability of The Microsoft R Application Network (MRAN), which will be shut down on July 1st, 2023.
+`groundhog` [@groundhogrpkg] used to be depending on MRAN but with a plan to switch to their home-grown R package repository.
 
@@
 In Section 3, `rang` is used to enable the reproducibility of published literature.
+However, one can still use `rang` for prospective usage and arguably with a longer term computational reproducibility than other solutions.
+In Section 4, `rang` is used to create an executable research compendium.
 
@@
 For instance, if there was a computational environment constructed on 2020-01-16 (called "snapshot date") with the several natural language processing R packages, `resolve()` can be used to resolve all the dependencies of these R packages.
-Currently, `rang` supports CRAN, Bioconductor, and GitHub packages.
+Currently, `rang` supports CRAN, Bioconductor, GitHub, and local packages.
 
@@
 For R >= 3.1, the images from the Rocker project are used [@boettiger:2017:IR].
-For R < 3.1 but >= 2.1, a custom image based on Debian Woody is used.
-As of writing, `rang` does not support R < 2.1, i.e.
-snapshot date earlier than 2005-04-19.
-The development to support R < 2.1 is in progress.
+For R < 3.1 but >= 1.3.1, a custom image based on Debian is used.
+As of writing, `rang` does not support R < 1.3.1, i.e.
+snapshot date earlier than 2001-08-31.
 There are two features of `dockerize()` that are important for future reproducibility.
@@
 docker build -t rang .
+## interactive environment
 docker run --rm --name "rangtest" -ti rang
@@
 `compute.es::tes()`.
-This undocumented package can be detected by `renv::dependencies()`.
+This undocumented package can be detected by `renv::dependencies()`, which is the provider of the scanning function of `rang`.
 
@@
 ```r
-graph <- resolve(c("readxl", "rms", "ordinal", "tidyverse", "lubridate", "MBESS",
-                   "censReg", "BayesFactor", "psych", "Joe-Hilgard/hilgard", "lsmeans",
-                   "compute.es"),
-                 snapshot_date = "2019-01-17")
+## scan all packages
+r_pkgs <- as_pkgrefs("vvg-2d4d")
+## replace cran::hilgard with Github
+r_pkgs[r_pkgs == "cran::hilhard"] <- "Joe-Hilgard/hilgard"
+graph <- resolve(r_pkgs, snapshot_date = "2019-01-17")
 ```
@@
 ```r
-graph <- resolve(c("foreign", "bife"), snapshot_date = "2018-12-12")
+## as_pkgrefs is automatically run in this case
+graph <- resolve("nathaniel", snapshot_date = "2018-12-12")
 dockerize(graph, output_dir = "nat", materials_dir = "nathaniel")

### 2023-03-08 paper/paper.qmd
Add first complete draft of the paper

@@
 The R package rang is a complete solution for generating the declarative description for other researchers to automatically reconstruct the computational environment at a specific time point.
-The reconstruction process, based on Docker, has been tested for R code as old as 2005 and does not depend on the long term availability of any commercial service.
-The description generated by rang satisfies the definition of a reproducible research compendium and can be shared as such.
+The reconstruction process, based on Docker, has been tested for R code as old as 2001.
+The declarative description generated by rang satisfies the definition of a reproducible research compendium and can be shared as such.
 In this contribution, we show how rang can be used to make otherwise unexecutable code, spanning from fields such as computational social science and bioinformatics, executable again.
-We also provide instructions on how to use rang to construct reproducible and shareable research compendium of current research.
+We also provide instructions on how to use rang to construct reproducible and shareable research compendia of current research.
 The package is currently available from CRAN (https://cran.r-project.org/web/packages/rang/index.html) and GitHub (https://github.com/chainsawriot/rang).
@@
 This is _NOT_ a problem in the method, it is entirely dependent on the numpy version (and even the OS's).
-If you have different versions of numpy or even the same version of numpy on a different OS configuration, different networks will fail randomly.
-...
+If you have different versions of numpy or even the same version of numpy on a different OS configuration, different networks will fail randomly...
 [F]or instance, the 109th Congress will fail, but will work entirely normally on a different numpy version, which will fail on a different Congress network."*
@@
 But as computer science research has shown, this task is incredibly complex [@abate:2015:MCR;@dolstra:2010:N].
-In the realm of R, that includes at least four aspects: a) operating system, b) system components such as `libxml2`, c) the exact R version, and d) what and which version of the installed R packages.
-We will call them component a, b, c, d in the following sections.
+In the realm of a usual scripting language such as R [^r], that pertains four aspects: a) operating system, b) system components such as `libxml2`, c) the exact R version, and d) what and which version of the installed R packages.
+We will call them Component A, B, C, D in the following sections.
 Any change in these four components can possibly affect the execution of any shared computer code.
@@
 
-In reality, the impact of operating system is relative weak as mainstream, open source programming languages and their software libraries are usually cross platform.
-In modern computational research, Linux is the de-facto operating system in high performance computing environments.
-Instead, the impact of components b, c, and d is practically higher.
-Component d is probably the most volatile among them all.
-Software update with breaking changes might render existing shared code not executable anymore.
-Also, software obsolescence is a common place, especially when academic software is often not well maintained [@merow:2023:B].
+[^r]: In this paper, we will focus on R, a popular programming language used frequently in various computational fields (e.g. computational social science, bioinformatics).
 
-The DevOps (software development and IT operations) community has come up with a partial solution to this problem: containerization.
-In essence, to containerize is to deploy the software together with all the libraries and even the operating system in a virtualized environment.
-In this way, software dependency issues can be resolved within the tightly controlled virtualized environment and independent of what is installed on the host computer.
+In reality, the impact of Component A is relative weak as mainstream, open source programming languages and their software libraries are usually cross platform.
+In modern computational research, Linux is the de-facto operating system in high performance computing environments (e.g.
+Slurm).
+Instead, the impact of Components B, C, and D is far higher.
+Component D is the most volatile among them all as there are many possible combinations of R packages and versions.
+Software update with breaking changes (even in a dependency) might render existing shared code using those changed features not executable or not producing the same result anymore.
+Also, software obsolescence is a common place, especially when academic software is often not well maintained due to lack of incentives [@merow:2023:B].
+
+The DevOps (software development and IT operations) community is also confronted with this problem.
+The issue is usually half-jokingly referred to as "it works on my machine"-problem [@valstar:2020:UDS, a software works on someone's local machine but is not working anymore when deployed to the production system, indicates the software tacitly depends on the computational environment of the local machine].
+A partial solution to this problem from the DevOps community is called *containerization*.
+In essence, to containerize is to develop and deploy the software together with all the libraries and the operating system in an OS-level virtualization environment.
+In this way, software dependency issues can be resolved inside the isolated virtualized software environment and independent of what is installed on the local computer.
+Docker is a popular choice in the DevOps world for containerization.
+
+To build a container, one needs to write in plain text a declarative description of the required computational environment.
+Inside this declarative description, it should pin down all four Components mentioned above.
+For Docker, it is in the form of a plain text file called `Dockerfile`.
+This `Dockerfile` is then used as the recipe to build a Docker image, where the four Components are assembled.
+Then, one can launch a container with the built Docker image.
+
+There has been many papers written on how containerization solutions such as Docker can be helpful also to foster computational reproducibility of science [e.g.
+@nuest:2019;@peikert:2021:RDA;@boettiger:2017:IR].
+Although tutorials are available [e.g.
+@nuest:2019], providing a declarative description of the computational environment in the form of Dockerfile is far from the standard code sharing practice.
+This might probably be due to most scientists do not have the (DevOps) skill to craft a Dockerfile [@kim:2018:E].
+But there are many tools available to automate the process [e.g.
+@nuest:2019].
+The case in point described in this paper, `rang`, is among one of them.
+And we argue that `rang` is the only easy-to-use solution available that can pin down and restore all four components without the reliance on any commercial service such as MRAN.
 
@@
 `renv` [@renvrpkg] (and its derivatives such as `jetpack` and its predecessor `packrat`) takes a similar approach to Python's `virtualenv` and Ruby's `Gem` to pin down the exact version of R packages using a "lock file".
-`containerit` [@nuest:2019] takes the current state of the computational environment and "containerize" it as a Dockerfile.
+Other solutions such as `checkpoint` [@checkpointrpkg] depends on the availability of The Microsoft R Application Network (MRAN, a time-stamped daily backup of CRAN), which will be shut down on July 1st, 2023.
+`groundhog` [@groundhogrpkg] used to be depending on MRAN but with a plan to switch to their home-grown R package repository.
+These solution can effectively pin down Component C and D.
+But they can only restore component D.
+Also, for solution dependent on MRAN, there is a limit on how far back can this reproducibility go, because MRAN can only go as far back as September 17, 2014.
+Also, it covers only CRAN packages.
+
+`containerit` [@nuest:2019] takes the current state of the computational environment and document it as a Dockerfile.
+`containerit` makes an assumption that Component A has a weak influence on computational reproducibility and therefore defaults to Linux-based Rocker images.
+In this way, it fixes Component A.
 But `containerit` does not pin down the exact version of R packages.
-`checkpoint` [@checkpointrpkg] depends on the availability of The Microsoft R Application Network (MRAN), which will be shut down on July 1st, 2023.
-`groundhog` [@groundhogrpkg] used to be depending on MRAN but with a plan to switch to their home-grown R package repository.
+Therefore, it can pin down components A, B, C, but only a part of component D.
+`dockta` is another containerization solution that can potentially pin down all components due to the fact that MRAN is used.
+But it also suffers from the same limitations mentioned above.
+
+It is also worth mention that MRAN is not the only archival service.
+Posit also provides a free (*gratis*) time-stamped daily backup of CRAN and Bioconductor (a series of repositories of R package for bioinformatics and computational biology) called Posit Public Package Manager (https://packagemanager.rstudio.com/client/#/repos/2/packages/).
+It can goes as far back as October 10, 2017.
 
@@
 `rang` mostly targets retrospective usage, i.e.
-using `rang` to reconstruct historical R computational environments which have not been completely declared.
+using `rang` to reconstruct historical R computational environments which the declarative descriptions are not available.
 One can think of `rang` as an archaeological tool.
-In Section 3, `rang` is used to enable the reproducibility of published literature.
-However, one can still use `rang` for prospective usage and arguably with a longer term computational reproducibility than other solutions.
-In Section 4, `rang` is used to create an executable research compendium.
+In this realm, we could not find any existing solution targeting specifically for R but yet not depending on MRAN.
+
+## Structure of this paper
+
+In Section 2, we will explain how to use `rang`.
+In Section 3, `rang` is used to enable the reproducibility of published literature (with increasing sophistication).
+However, one can still use `rang` for prospective usage and arguably can ensure a longer term computational reproducibility than other solutions.
+In Section 4, `rang` is used to create an executable research compendium with Docker and `Make` [@baker:2020:UGM].
 
@@
 `resolve()` queries various web services from the r-hub project of the R Consortium for information about R packages at a specific time point that is necessary for reconstructing a computational environment, e.g.
-(deep) dependencies, system requirements, and R version.
+(deep) dependencies (Component D), R version (Component C), and system requirements (Component B).
 For instance, if there was a computational environment constructed on 2020-01-16 (called "snapshot date") with the several natural language processing R packages, `resolve()` can be used to resolve all the dependencies of these R packages.
@@
 As of writing, `rang` does not support R < 1.3.1, i.e.
-snapshot date earlier than 2001-08-31.
+snapshot date earlier than 2001-08-31 (which is 13 years earlier than all solutions depend on MRAN).
 There are two features of `dockerize()` that are important for future reproducibility.
@@
 1.
-By default, the container building process downloads source packages from CRAN and then compiles them from source.
-This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) [^1].
+By default, the container building process downloads source packages from their sources and then compiles them.
+This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) [^1]m Bioconductor, and Github.
 However, it is also possible to cache (or archive) the source packages now.
 The archived R packages can then be used instead during the building process.
+The significance of this step in terms of long-term computational reproducibility will be discussed in Section 4.
 
@@
 
+## Project scanning
+
+The first argument of `resolve()` is processed by a separate function called `as_pkgrefs()`.
+For interoperability, `rang` supports the "package references" standard [^pkgrefs] used also in other packages such as `renv` [@renvrpkg].
+It is mostly used for converting "shorthands" (e.g.
+`xml2` and `S4Vectors`) to package references (e.g.
+`cran::xml2` and `bioc::S4Vectors`).
+
+When `as_pkgrefs()` is applied to a single path of a directory, it scans all relevant files (`DESCRIPTION`, R scripts and R Markdown files) for all R packages used (based on `renv::dependencies()` <!-- [^gh] -->).
+How it works is demonstrated in three of the following examples below.
+But an important caveat is that it can only scan CRAN and Bioconductor packages.
+
+[^pkgrefs]: https://r-lib.github.io/pkgdepends/reference/pkg_refs.html
+
+<!-- [^gh]: This method can't scan R scripts written with `groundhog` unfortunately. In the example of `grounghog::groundhog.library("MASS")`, `MASS` is not considered as a used R package. -->
+
 # Case Studies
 
-The following are some examples of how `rang` can be used to make otherwise shared, but unexecutable, R code runnable again.
+The following are some examples of how `rang` can be used to make shared, but otherwise unexecutable, R code runnable again.
 The examples were drawn from various fields spanning from political science, psychological science, and bioinformatics.
@@
 before 2004-10-04).
-Similar to the case of `maxent` above, a Dockerfile to generate a Docker image with `ptproc` installed can be generated with two lines of code.
+Similar to the case of `maxent` above, a Dockerfile to assemble a Docker image with `ptproc` installed can be generated with two lines of code.
 
@@
 
-Similar to CRAN, Bioconductor is a series of repositories of R package for bioinformatics and computational biology. Also similar to CRAN, packages can get removed over time.
-
+Also similar to CRAN, packages can get removed over time on Bioconductor.
 The Bioconductor package `Sushi` has been deprecated by the original authors and is removed from Bioconductor version 3.16 (2022-11-02).
@@
 x <- resolve(cran_pkgs, "2021-08-11", verbose = TRUE)
-print(x, all_pkgs = TRUE)
-dockerize(x, "oser_docker", materials_dir = "meta-analysis", cache = TRUE)
+##print(x, all_pkgs = TRUE)
+dockerize(x, "oserdocker", materials_dir = "meta-analysis", cache = TRUE)
 ```
@@
 ```Makefile
-rmd = "rmarkdown::render('materials/README.Rmd', output_file = 'reproduced.html')"
-resolve:
-	Rscript oser.R
-build: oser_docker
-	docker build -t oserimg oser_docker
-render:
-	docker run -d --rm --name "osercontainer" -ti oserimg
-	docker exec osercontainer Rscript -e ${rmd}
-	docker cp osercontainer:/materials/reproduced.html oser_README.html
-	docker stop osercontainer
-export:
-	docker save oserimg | gzip > oserimg.tar.gz
-rebuild: oserimg.tar.gz
-	docker load < oserimg.tar.gz
+output_file=reproduced.html
+r_cmd = "rmarkdown::render('materials/README.Rmd', output_file = '${output_file}')"
+handle=oser
+local_file=${handle}_README.html
+
 all: resolve build render
 	echo "finished"
+resolve:
+	Rscript ${handle}.R
+build: ${handle}docker
+	docker build -t ${handle}img ${handle}docker
+render:
+	docker run -d --rm --name "${handle}container" -ti ${handle}img
+	docker exec ${handle}container Rscript -e ${r_cmd}
+	docker cp ${handle}container:/materials/${output_file} ${local_file}
+	docker stop ${handle}container
+export:
+	docker save ${handle}img | gzip > ${handle}img.tar.gz
+rebuild: ${handle}img.tar.gz
+	docker load < ${handle}img.tar.gz
 ```
 
-With this `Makefile`, once can create the Dockerfile with `make resolve`, build the Docker image with `make build`, render the RMarkdown file inside the container with `make render`, export the built Docker image with `make export`, and rebuild the exported Docker image with `make rebuild`.
+With this `Makefile`, one can create the Dockerfile with `make resolve`, build the Docker image with `make build`, render the RMarkdown file inside the container with `make render`, export the built Docker image with `make export`, and rebuild the exported Docker image with `make rebuild`.
 
-The structure of the executable compendium looks like this:
+The structure of the entire executable compendium looks like this:
 
@@
 README.md
-oser_docker/
+oserdocker/
 oserimg.tar.gz
@@
 In this executable compendium, only the first four elements are essential.
-The directory `oser_docker` (116 MB) contains cached R packages, Dockerfile, and a further copy of the directory `meta-analysis/` to be transferred into the Docker image.
+The directory `oserdocker` (116 MB) contains cached R packages, Dockerfile, and a verbatim copy of the directory `meta-analysis/` to be transferred into the Docker image.
 That can be regenerated by running `make resolve`.
-However, having this directory preserved insures against the situations of some R packages used in the project are no longer available or any of the information providers used by `rang` for resolving the dependency relationships being not available.
+However, having this directory preserved insures against the situations that some R packages used in the project were no longer available or any of the information providers used by `rang` for resolving the dependency relationships were not available.
 (Or in the rare circumstance of `rang` is no longer available.)
@@
 Preserving this file insurces against all the situations mentioned above, but also the situations of Docker Hub and the software repositories used by the dockerized operating system being not available.
-As a matter of fact, when `oserimg.tar.gz` is available it is possible to run `make rebuild` and `make render` even without internet access, provided that only Docker and make have been preinstalled.
-Of course, there is still an extremely rare situation where Docker (the program) itself is no longer available.
-However, it is possible to convert the image file for use on other containerization solutions such as Singularity [^singularity].
+When `oserimg.tar.gz` is available, it is possible to run `make rebuild` and `make render` even without internet access (provided that Docker and `make` have been preinstalled).
+Of course, there is still an extremely rare situation where Docker (the program) itself is no longer available [^make].
+However, it is possible to convert the image file for use on other containerization solutions such as Singularity [^singularity], if Docker is really not available anymore.
 
-Sharing of research artefacts less than 1G is not as challenging as it used to be.
+Sharing of research artifacts less than 1G is not as challenging as it used to be.
 Zenodo, for example, allows the sharing of 50G of files.
@@
 However, for data repositories with more restrictions on data size, sharing the executable compendium without the last two parts could be considered sufficient.
-For that, run `make all` will generate all the things needed for reproducing the analysis inside a container.
+For that, run `make` will make the default target `all` and generate all the things needed for reproducing the analysis inside a container.
+
+The above `Makefile` is general enough that one can reuse it by just modifying how the R scripts (the `r_cmd` variable) in the `materials` directory are executed.
+This can be a starting point of a standard executable compendium format.
+
+[^make]: We can't imagine a world without `Make`, a tool that has been available since 1976.
 
@@
 
-# Conclusion
+# Concluding remarks
+
+This paper presents `rang`, a solution to (re)construct R computational environments based on Docker.
+As the six examples in Section 3 shown, `rang` can be used archaeologically to rerun some old code, many of them not executable without the analytic and reconstruction processes facilitated by `rang`.
+These retrospectively use cases demonstrate how versatile `rang` is.
+`rang` is also helpful for prospective usage, as demonstrated in Section 4 whereby an executable compendium is created.
+
+There are still many features that we did not mention in this paper.
+`rang` is built with interoperability in mind.
+As of writing, `rang` is interoperable with existing R packages such as `renv` and R built-in `sessionInfo()`.
+Also, the `rang` object can be used for network analysis with R packages such as `igraph`.
+
+Computational reproducibility is a complex topic and as in all of these complex topic, there is no silver bullet [@canon:2019:CPR].
+All solutions have their trade-offs.
+The (re)construction process based on `rang` takes notably more time than other solutions because all packages are compiled from source.
+`rang` trades computational efficiency of this often one-off (re)constructing process for correctness, backward compatibility and independence from any commercial backups of software repositories such as MRAN.
+There are also other limitations.
+In the Vignette of `rang` (https://cran.r-project.org/web/packages/rang/vignettes/faq.html), we list out all of these limitations as well as possible mitigation.
 

### 2023-03-08 paper/oser/oser.R

@@ -5,5 +5,5 @@ cran_pkgs <- as_pkgrefs("meta-analysis") ## dmetar is an undeclared github packa
 
 cran_pkgs[cran_pkgs == "cran::dmetar"] <- "MathiasHarrer/dmetar"
 x <- resolve(cran_pkgs, "2021-08-11", verbose = TRUE)
-print(x, all_pkgs = TRUE)
-dockerize(x, "oser_docker", materials_dir = "meta-analysis", cache = TRUE)
+##print(x, all_pkgs = TRUE)
+dockerize(x, "oserdocker", materials_dir = "meta-analysis", cache = TRUE)

### 2023-03-08 paper/peng.sh
Remove `load_all()` in `peng.sh`

@@ -1,4 +1,4 @@
-Rscript -e "devtools::load_all(); dockerize(resolve('ptproc', '2004-07-01'), 'pengdocker', cache = TRUE)"
+Rscript -e "require(rang); dockerize(resolve('ptproc', '2004-07-01'), 'pengdocker', cache = TRUE)"
 docker build -t pengimg ./pengdocker
 docker run -d --rm --name "pengcontainer" -ti pengimg
 docker cp peng.R pengcontainer:/peng.R

### 2023-03-08 paper/sushi.sh
Correct the arxiv template the Sushi.sh

@@ -1,4 +1,4 @@
-Rscript -e "devtools::load_all(); dockerize(resolve('Sushi', '2014-06-05'),
+Rscript -e "require(rang); dockerize(resolve('Sushi', '2014-06-05'),
 'sushidocker', no_rocker = TRUE, cache = TRUE)"
 docker build -t sushiimg ./sushidocker
 docker run -d --rm --name "sushicontainer" -ti sushiimg

### 2023-03-08 paper/paper.qmd

@@
 
-In reality, the impact of Component A is relative weak as mainstream, open source programming languages and their software libraries are usually cross platform.
+In reality, the impact of Component A is relatively weak as mainstream, open source programming languages and their software libraries are usually cross platform.
 In modern computational research, Linux is the de-facto operating system in high performance computing environments (e.g.
 Slurm).
-Instead, the impact of Components B, C, and D is far higher.
+Instead, the impact of Components B, C, and D is much higher.
 Component D is the most volatile among them all as there are many possible combinations of R packages and versions.
-Software update with breaking changes (even in a dependency) might render existing shared code using those changed features not executable or not producing the same result anymore.
-Also, software obsolescence is a common place, especially when academic software is often not well maintained due to lack of incentives [@merow:2023:B].
+Software updates with breaking changes (even in a dependency) might render existing shared code using those changed features not executable or not producing the same result anymore.
+Also, software obsolescence is commonplace, especially since academic software is often not well maintained due to lack of incentives [@merow:2023:B].
 
@@
 
-To build a container, one needs to write in plain text a declarative description of the required computational environment.
+To build a container, one needs to write a plain text declarative description of the required computational environment.
 Inside this declarative description, it should pin down all four Components mentioned above.
@@
 @nuest:2019], providing a declarative description of the computational environment in the form of Dockerfile is far from the standard code sharing practice.
-This might probably be due to most scientists do not have the (DevOps) skill to craft a Dockerfile [@kim:2018:E].
+This might be due to a lack of (DevOps) skills of most scientists to create a Dockerfile [@kim:2018:E].
 But there are many tools available to automate the process [e.g.
 @nuest:2019].
-The case in point described in this paper, `rang`, is among one of them.
-And we argue that `rang` is the only easy-to-use solution available that can pin down and restore all four components without the reliance on any commercial service such as MRAN.
+The case in point described in this paper, `rang`, is one of them.
+We argue that `rang` is the only easy-to-use solution available that can pin down and restore all four components without the reliance on any commercial service such as MRAN.
 
@@
 `renv` [@renvrpkg] (and its derivatives such as `jetpack` and its predecessor `packrat`) takes a similar approach to Python's `virtualenv` and Ruby's `Gem` to pin down the exact version of R packages using a "lock file".
-Other solutions such as `checkpoint` [@checkpointrpkg] depends on the availability of The Microsoft R Application Network (MRAN, a time-stamped daily backup of CRAN), which will be shut down on July 1st, 2023.
-`groundhog` [@groundhogrpkg] used to be depending on MRAN but with a plan to switch to their home-grown R package repository.
+Other solutions such as `checkpoint` [@checkpointrpkg] depend on the availability of The Microsoft R Application Network (MRAN, a time-stamped daily backup of CRAN), which will be shut down on July 1st, 2023.
+`groundhog` [@groundhogrpkg] used to depend on MRAN but has a plan to switch to their home-grown R package repository.
 These solution can effectively pin down Component C and D.
 But they can only restore component D.
-Also, for solution dependent on MRAN, there is a limit on how far back can this reproducibility go, because MRAN can only go as far back as September 17, 2014.
-Also, it covers only CRAN packages.
+Also, for solutions depending on MRAN, there is a limit on how far back this reproducibility can go, since MRAN can only go back as far as September 17, 2014.
+Additionally, it only covers CRAN packages.
 
-`containerit` [@nuest:2019] takes the current state of the computational environment and document it as a Dockerfile.
-`containerit` makes an assumption that Component A has a weak influence on computational reproducibility and therefore defaults to Linux-based Rocker images.
+`containerit` [@nuest:2019] takes the current state of the computational environment and documents it as a Dockerfile.
+`containerit` makes the assumption that Component A has a weak influence on computational reproducibility and therefore defaults to Linux-based Rocker images.
 In this way, it fixes Component A.
@@
 
-It is also worth mention that MRAN is not the only archival service.
+It is also worth mentioning that MRAN is not the only archival service.
 Posit also provides a free (*gratis*) time-stamped daily backup of CRAN and Bioconductor (a series of repositories of R package for bioinformatics and computational biology) called Posit Public Package Manager (https://packagemanager.rstudio.com/client/#/repos/2/packages/).
@@
 `rang` mostly targets retrospective usage, i.e.
-using `rang` to reconstruct historical R computational environments which the declarative descriptions are not available.
+using `rang` to reconstruct historical R computational environments for which the declarative descriptions are not available.
 One can think of `rang` as an archaeological tool.
-In this realm, we could not find any existing solution targeting specifically for R but yet not depending on MRAN.
+In this realm, we could not find any existing solution targeting R specifically which does not currently depend on MRAN.
 
@@
 As of writing, `rang` does not support R < 1.3.1, i.e.
-snapshot date earlier than 2001-08-31 (which is 13 years earlier than all solutions depend on MRAN).
+snapshot date earlier than 2001-08-31 (which is 13 years earlier than all solutions depending on MRAN).
 There are two features of `dockerize()` that are important for future reproducibility.
@@
 By default, the container building process downloads source packages from their sources and then compiles them.
-This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) [^1]m Bioconductor, and Github.
+This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) [^1], Bioconductor, and Github.
 However, it is also possible to cache (or archive) the source packages now.
@@
 
-The same file can also be "rescued" by `rang`.
+The same file can thus also be "rescued" by `rang`.
 
@@
 Even with this over-a-decade removal and new packages with similar functionalities have been created, there is evidence that `ptproc` is still being sought for.
-As late as 2017, there are blog posts on how to install the long obsolete package on modern R [^blog].
+As late as 2017, there are blog posts on how to install the long obsolete package on modern versions of R [^blog].
 The package is extremely challenging to install on a modern R system because the package was written before the introduction of name space management in R 1.7.0 [@RN-2003-001].
@@
 
-Also similar to CRAN, packages can get removed over time on Bioconductor.
+Similar to CRAN, packages can also be removed over time from Bioconductor.
 The Bioconductor package `Sushi` has been deprecated by the original authors and is removed from Bioconductor version 3.16 (2022-11-02).
@@
 
-`rang` has native support for Bioconductor packages since 0.2.
-We obtained the R script `"PaperFigure.R"` from the Github repo of `Sushi` [^sushi], which generates the figure in @phanstiel:2014:S.
+`rang` has native support for Bioconductor packages since version 0.2.
+We obtained the R script `"PaperFigure.R"` from the Github repository of `Sushi` [^sushi], which generates the figure in @phanstiel:2014:S.
 Similar to the above case of `ptproc`, we made a completely automated BASH script to run `"PaperFigure.R"` and get the generated figure out of the container (@fig-figure2).
@@
 
-The above six examples show how useful it is for `rang` reconstructing tricky computational environments which have not been completely declared in the literature.
-Although we position `rang` mostly as an archaeological tool, we also think that `rang` can also be used to prepare research compendia of current research.
+The above six examples show how powerful `rang` is to reconstruct tricky computational environments which have not been completely declared in the literature.
+Although we position `rang` mostly as an archaeological tool, we think that `rang` can also be used to prepare research compendia of current research.
 We can't predict the future but research compendia generated by `rang` would probably have long-term computational reproducibility.
@@
 In this executable compendium, only the first four elements are essential.
-The directory `oserdocker` (116 MB) contains cached R packages, Dockerfile, and a verbatim copy of the directory `meta-analysis/` to be transferred into the Docker image.
+The directory `oserdocker` (116 MB) contains cached R packages, a Dockerfile, and a verbatim copy of the directory `meta-analysis/` to be transferred into the Docker image.
 That can be regenerated by running `make resolve`.
@@
 This can be regenerated by running `make export`.
-Preserving this file insurces against all the situations mentioned above, but also the situations of Docker Hub and the software repositories used by the dockerized operating system being not available.
-When `oserimg.tar.gz` is available, it is possible to run `make rebuild` and `make render` even without internet access (provided that Docker and `make` have been preinstalled).
+Preserving this file insures against all the situations mentioned above, but also the situations of Docker Hub and the software repositories used by the dockerized operating system being not available.
+When `oserimg.tar.gz` is available, it is possible to run `make rebuild` and `make render` even without internet access (provided that Docker and `make` have been installed before).
 Of course, there is still an extremely rare situation where Docker (the program) itself is no longer available [^make].
@@
 This paper presents `rang`, a solution to (re)construct R computational environments based on Docker.
-As the six examples in Section 3 shown, `rang` can be used archaeologically to rerun some old code, many of them not executable without the analytic and reconstruction processes facilitated by `rang`.
-These retrospectively use cases demonstrate how versatile `rang` is.
+As the six examples in Section 3 show, `rang` can be used archaeologically to rerun old code, many of them not executable without the analytic and reconstruction processes facilitated by `rang`.
+These retrospective use cases demonstrate how versatile `rang` is.
 `rang` is also helpful for prospective usage, as demonstrated in Section 4 whereby an executable compendium is created.
@@
 There are also other limitations.
-In the Vignette of `rang` (https://cran.r-project.org/web/packages/rang/vignettes/faq.html), we list out all of these limitations as well as possible mitigation.
+In the Vignette of `rang` (https://cran.r-project.org/web/packages/rang/vignettes/faq.html), we list all of these limitations as well as possible mitigation.
 

### 2023-03-08 paper/paper.qmd
Add arxiv version

@@
 Zenodo, for example, allows the sharing of 50G of files.
-Therefore, sharing of the last two components of the executable compendium prepared with `rang` is at least possible on Zenodo.
+Therefore, sharing of the last two components of the executable compendium prepared with `rang` is at least possible on Zenodo [^zenodo].
 However, for data repositories with more restrictions on data size, sharing the executable compendium without the last two parts could be considered sufficient.
 For that, run `make` will make the default target `all` and generate all the things needed for reproducing the analysis inside a container.
+
+[^zenodo]: The complete version of the executable compendium is available from Zenodo: https://doi.org/10.5281/zenodo.7708417
 
@@
 
-Computational reproducibility is a complex topic and as in all of these complex topic, there is no silver bullet [@canon:2019:CPR].
+Computational reproducibility is a complex topic and as in all of these complex topics, there is no silver bullet [@canon:2019:CPR].
 All solutions have their trade-offs.

### 2023-03-08 README.Rmd

@@
 
-To obtain the dependency graph of R packages, use `resolve`. Currently, this package supports both CRAN and Github packages.
+To obtain the dependency graph of R packages, use `resolve`. Currently, this package supports CRAN, Bioconductor, GitHub, and local packages.
 
@@
 library(rang)
-x <- resolve(pkgs = c("sna", "schochastics/rtoot"), snapshot_date = "2022-11-30")
+x <- resolve(pkgs = c("sna", "schochastics/rtoot", "S4Vectors"), snapshot_date = "2022-11-30")
 ```

### 2023-03-09 paper/nathaniel/fn_5.R
Paper (#109)
* Change basic content [paper] [no ci]

* Add David as an author [no ci]

* Rewrite the paper to reflect the name change

* just minor typo/grammar fixes

* Add Hilgard

* Add hilgard

* minor typo and grammar fixes

* Add political analysis and readme

* Add `maxent` example

* Add pdf

* Switch over to quarto

* Update the abstract about Bioconductor

* Further revise the abstract

* Update the paper

* in

* Add peng example

* Add all examples and compendium

* Clean up to reflect the upcoming v0.2

* Add first complete draft of the paper

* Update the paper readme

* delete r

@@ -1,0 +1,29 @@
+                                        #Before starting we route all output to a file, fn_5.txt
+sink("fn_5.txt")
+                                        #First we load library foreign and library bife
+                                        #Note the library command is hardcoded to the standard file for libraries
+                                        #If your libraries are different, or they fail to load,  look at the commented out lib.loc subcommand
+#These are the only R packages needed that are not in baseR
+library("foreign")
+#lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
+library("bife")
+#lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
+                                        # now we read in the same Stata data as used in Table_1, saved as an early version Stata dta file, besley.dta which is put into data_b
+                                        #For timing
+start.time <- Sys.time()
+ data_b<-read.dta("besley.dta")
+glm_naive<-glm(graduate~Democracy+logGDPcapita+factor(country),data=data_b,family=binomial(link=logit))
+ summary.glm(glm_naive)
+ #now we run bife to both get correct LogitFE results and to get the data set corresponding to NotAllZero groups) - we run "vanilla" bife without cluster robust standard errors
+ bife_all<-bife(graduate~Democracy+logGDPcapita|country,data=data_b,bias_corr="no")
+ summary(bife_all,fixed="TRUE")
+                                        #Now we create the NotAllZero data set for glm to show this works
+y_notall0<-bife_all$model_info$y
+x_notall0<-bife_all$model_info$X
+country_notall0<-bife_all$model_info$id
+glm_notall0 <- glm(y_notall0~x_notall0+factor(country_notall0),family=binomial(link=logit))
+summary.glm(glm_notall0)
+end.time <- Sys.time()
+time.taken <- end.time - start.time
+cat("Execution time:",  time.taken)
+closeAllConnections()

### 2023-03-09 paper/oser/oser.R

@@ -1,0 +1,9 @@
+require(rang)
+
+## all packages used in the project
+cran_pkgs <- as_pkgrefs("meta-analysis") ## dmetar is an undeclared github package: MathiasHarrer/dmetar
+
+cran_pkgs[cran_pkgs == "cran::dmetar"] <- "MathiasHarrer/dmetar"
+x <- resolve(cran_pkgs, "2021-08-11", verbose = TRUE)
+##print(x, all_pkgs = TRUE)
+dockerize(x, "oserdocker", materials_dir = "meta-analysis", cache = TRUE)

### 2023-03-09 paper/peng.R

@@ -1,0 +1,21 @@
+##require(rang)
+
+##graph <- resolve("./ptproc_1.5-1.tar.gz", "2004-07-01")
+##dockerize(graph, "~/dev/misc/ptproc", cache = TRUE)
+
+require(ptproc)
+
+set.seed(1000)
+x <- cbind(runif(100), runif(100), runif(100))
+hPois.cond.int <- function(params, eval.pts, pts = NULL, data = NULL, TT = NULL) {
+    mu <- params[1]
+    if(is.null(TT))
+        rep(mu, nrow(eval.pts))
+    else {
+        vol <- prod(apply(TT, 2, diff))
+        mu * vol
+    }
+}
+ppm <- ptproc(pts = x, cond.int = hPois.cond.int, params = 50, ranges = cbind(c(0,1), c(0,1), c(0,1)))
+fit <- ptproc.fit(ppm, optim.control = list(trace = 2), method = "BFGS")
+summary(fit)

### 2023-03-09 paper/peng.sh

@@ -1,0 +1,8 @@
+Rscript -e "require(rang); dockerize(resolve('ptproc', '2004-07-01'), 'pengdocker', cache = TRUE)"
+docker build -t pengimg ./pengdocker
+docker run -d --rm --name "pengcontainer" -ti pengimg
+docker cp peng.R pengcontainer:/peng.R
+docker exec pengcontainer R CMD BATCH peng.R
+docker exec pengcontainer cat peng.Rout
+docker cp pengcontainer:/peng.Rout peng.Rout
+docker stop pengcontainer

### 2023-03-09 paper/sushi.sh

@@ -1,0 +1,9 @@
+Rscript -e "require(rang); dockerize(resolve('Sushi', '2014-06-05'),
+'sushidocker', no_rocker = TRUE, cache = TRUE)"
+docker build -t sushiimg ./sushidocker
+docker run -d --rm --name "sushicontainer" -ti sushiimg
+docker cp PaperFigure.R sushicontainer:/PaperFigure.R
+docker exec sushicontainer mkdir vignettes
+docker exec sushicontainer R CMD BATCH PaperFigure.R
+docker cp sushicontainer:/vignettes/Figure_1.pdf sushi_figure1.pdf
+docker stop sushicontainer

### 2023-03-09 README.Rmd
Add citation information
add refinements before CRAN submission [no ci]

@@
 
-To reconstruct a historical R computational environment, this package assumes only the availability of source packages on CRAN.
+To reconstruct a historical R computational environment, this package assumes only the availability of source packages online.
 The reconstruction procedures have been tested in several vintage versions of R.
+
+Please cite this package as:
+
+Chan CH, Schoch D (2023) rang: Reconstructing reproducible R computational environments. arXiv preprint:[2303.04758](https://doi.org/10.48550/arXiv.2303.04758)
 

### 2023-03-09 R/installation.R

@@ -309,7 +309,7 @@
 #' @param check_cran_mirror logical, whether to check the CRAN mirror
 #' @param bioc_mirror character, which Bioconductor mirror to use
 #' @return `path`, invisibly
-#' @details The idea behind this is to determine the installation order of R packages locally. Then, the installation script can be depolyed to another
+#' @details The idea behind this is to determine the installation order of R packages locally. Then, the installation script can be deployed to another
 #' fresh R session to install R packages. [dockerize()] is a more reasonable way because a fresh R session with all system requirements
 #' is provided. The current approach does not work in R < 2.1.0.
 #' @export
@@ -401,11 +401,11 @@ export_renv <- function(rang, path = ".") {
     for(i in seq_len(nrow(pkg_df))){
         pkg_list[[i]][["Package"]] <- pkg_df$x[i]
         pkg_list[[i]][["Version"]] <- pkg_df$version[i]
-        if(pkg_df$source[i]=="cran"){
+        if(pkg_df$source[i] == "cran") {
             pkg_list[[i]][["Source"]] <- "Repository"
             pkg_list[[i]][["Repository"]] <- "CRAN"
         #   pkg_list[[i]][["Requirements"]] <- c()
-        } else if(pkg_df$source[i]=="github"){
+        } else if(pkg_df$source[i] == "github") {
             pkg_list[[i]][["Source"]] <- "GitHub"
             pkg_list[[i]][["RemoteType"]] <- "GitHub"
             pkg_list[[i]][["RemoteHost"]] <- "api.github.com"
@@ -414,18 +414,18 @@ export_renv <- function(rang, path = ".") {
             pkg_list[[i]][["RemoteRef"]] = "HEAD"
             pkg_list[[i]][["RemoteSha"]] = pkg_df$uid[i]
         #   pkg_list[[i]][["Requirements"]] <- c()
-        } else if(pkg_df$source[i]=="bioc"){
+        } else if(pkg_df$source[i] == "bioc") {
             pkg_list[[i]][["Source"]] <- "Bioconductor"
             # pkg_list[[i]][["git_url"]] <- paste0("https://git.bioconductor.org/packages/",pkg_df$x[i])
             # pkg_list[[i]][["git_branch"]] <- ""
             # pkg_list[[i]][["git_last_commit"]] <- ""
             # pkg_list[[i]][["git_last_commit_date"]] <- ""
             # pkg_list[[i]][["Requirements"]] <- c()
[…]

### 2023-03-10 paper/paper.qmd
Fix formatting issues #111

@@
 
-> ~ excerpt of [this README file](https://www.michelecoscia.com/?page_id=2105)
+> ~ excerpt of [the README file](https://www.michelecoscia.com/?page_id=2105) of a software for polarization estimation
 
@@
 But as computer science research has shown, this task is incredibly complex [@abate:2015:MCR;@dolstra:2010:N].
-In the realm of a usual scripting language such as R [^r], that pertains four aspects: a) operating system, b) system components such as `libxml2`, c) the exact R version, and d) what and which version of the installed R packages.
+For a usual scripting language such as R [^r], that pertains to four aspects: A) operating system, B) system components such as `libxml2`, C) the exact R version, and D) what and which version of the installed R packages.
 We will call them Component A, B, C, D in the following sections.
@@
 The case in point described in this paper, `rang`, is one of them.
-We argue that `rang` is the only easy-to-use solution available that can pin down and restore all four components without the reliance on any commercial service such as MRAN.
+We argue that `rang` is the only easy-to-use solution available that can pin down and restore all four components without the reliance on any commercial service <!-- such as MRAN -->.
 
@@
 It is also worth mentioning that MRAN is not the only archival service.
-Posit also provides a free (*gratis*) time-stamped daily backup of CRAN and Bioconductor (a series of repositories of R package for bioinformatics and computational biology) called Posit Public Package Manager (https://packagemanager.rstudio.com/client/#/repos/2/packages/).
+Posit also provides a free (*gratis*) time-stamped daily backup of CRAN and Bioconductor (a series of repositories of R package for bioinformatics and computational biology) called Posit Public Package Manager [^posit].
 It can goes as far back as October 10, 2017.
+
+[^posit]: [https://packagemanager.rstudio.com/client/#/repos/2/packages/](https://packagemanager.rstudio.com/client/#/repos/2/packages/)
 
@@
 
-## Structure of this paper
+<!-- ## Structure of this paper -->
 
-In Section 2, we will explain how to use `rang`.
+<!-- In Section 2, we will explain how to use `rang`.
 In Section 3, `rang` is used to enable the reproducibility of published literature (with increasing sophistication).
@@
 In Section 4, `rang` is used to create an executable research compendium with Docker and `Make` [@baker:2020:UGM].
+-->
 
@@
 
-[^1]: https://stat.ethz.ch/pipermail/r-announce/1997/000001.html
+[^1]: [https://stat.ethz.ch/pipermail/r-announce/1997/000001.html](https://stat.ethz.ch/pipermail/r-announce/1997/000001.html)
 
@@
 
-[^pkgrefs]: https://r-lib.github.io/pkgdepends/reference/pkg_refs.html
+[^pkgrefs]: [https://r-lib.github.io/pkgdepends/reference/pkg_refs.html](https://r-lib.github.io/pkgdepends/reference/pkg_refs.html)
 
@@
 
-[^hilgard]: https://github.com/Joe-Hilgard/vvg-2d4d
+[^hilgard]: [https://github.com/Joe-Hilgard/vvg-2d4d](https://github.com/Joe-Hilgard/vvg-2d4d)
 
@@
 
-[^groundhog]: http://datacolada.org/100
+[^groundhog]: [http://datacolada.org/100](http://datacolada.org/100)
 
@@
 
-[^checkerror]: https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html
+[^checkerror]: [https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html](https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html)
 
-[^evidence]: https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9
+[^evidence]: [https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9](https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9)
 
@@
 
-[^taskview]: https://cran.r-project.org/web/views/SpatioTemporal.html
+[^taskview]: [https://cran.r-project.org/web/views/SpatioTemporal.html](https://cran.r-project.org/web/views/SpatioTemporal.html)
 
@@
 The package is extremely challenging to install on a modern R system because the package was written before the introduction of name space management in R 1.7.0 [@RN-2003-001].
-In other words, the available tarball file from the original author's website does not contain a `NAMESPACE` file as all other modern R packages do.
+In other words, the available tarball files from the original author's website and CRAN do not contain a `NAMESPACE` file like all other modern R packages do.
 
@@
 x <- cbind(runif(100), runif(100), runif(100))
-hPois.cond.int <- function(params, eval.pts, pts = NULL, data = NULL, TT = NULL) {
+hPois.cond.int <- function(params, eval.pts, pts = NULL,
+                           data = NULL, TT = NULL) {
     mu <- params[1]
@@
 
-[^peng]: https://www.biostat.jhsph.edu/~rpeng/software/
+[^peng]: [https://www.biostat.jhsph.edu/~rpeng/software/](https://www.biostat.jhsph.edu/~rpeng/software/)
 
-[^blog]: https://blog.mathandpencil.com/installing-ptproc-on-osx and https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/
+[^blog]: [https://blog.mathandpencil.com/installing-ptproc-on-osx](https://blog.mathandpencil.com/installing-ptproc-on-osx) and [https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/](https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/)
 
@@
 
-[^sushi]: https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R
+[^sushi]: [https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R](https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R)
 
@@
 
-The above R script is saved as `oser.R`. The central piece of the executable compendium is the `Makefile`.
+The above R script is saved as `oser.R`. The central piece of the executable compendium is the `Makefile` [@baker:2020:UGM].
 
@@
 output_file=reproduced.html
-r_cmd = "rmarkdown::render('materials/README.Rmd', output_file = '${output_file}')"
+r_cmd = "rmarkdown::render('materials/README.Rmd',\
+output_file = '${output_file}')"
 handle=oser
@@
 
-[^zenodo]: The complete version of the executable compendium is available from Zenodo: https://doi.org/10.5281/zenodo.7708417
+[^zenodo]: The complete version of the executable compendium is available on Zenodo: [https://doi.org/10.5281/zenodo.7708417](https://doi.org/10.5281/zenodo.7708417)
 
@@
 
-[^singularity]: https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html
+[^singularity]: [https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html](https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html)
 
@@
 There are also other limitations.
-In the Vignette of `rang` (https://cran.r-project.org/web/packages/rang/vignettes/faq.html), we list all of these limitations as well as possible mitigation.
+In the Vignette of `rang` ([https://cran.r-project.org/web/packages/rang/vignettes/faq.html](https://cran.r-project.org/web/packages/rang/vignettes/faq.html)), we list all of these limitations as well as possible mitigation.
 

### 2023-03-10 paper/paper.qmd
Fix the confusing CRAN terminology #111

@@
 
-[^checkerror]: [https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html](https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html)
+[^checkerror]: This is the check log: [https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html](https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html) Here, we said the package `maxent` was "removed from CRAN" as per what the webpage ([https://cran.r-project.org/web/packages/maxent/index.html](https://cran.r-project.org/web/packages/maxent/index.html)) is written.
+However, the terminology as stated in the [CRAN Policies](https://cran.r-project.org/web/packages/policies.html#Source-packages) is "archived", quote: "Packages will not normally be removed from CRAN: however, they may be archived, including at the maintainer’s request." As all CRAN packages have all the submitted versions archived, we find this terminology confusing.
+Therefore, we use "removed" packages throughout this paper to indicate packages that cannot be installed by the usual method, i.e.
+`install.packages` but still have the old versions archived on CRAN.
 

### 2023-03-10 paper/paper.qmd
Fix issues related to Docker in the paper #111

@@
 `containerit` [@nuest:2019] takes the current state of the computational environment and documents it as a Dockerfile.
-`containerit` makes the assumption that Component A has a weak influence on computational reproducibility and therefore defaults to Linux-based Rocker images.
+`containerit` makes the assumption that Component A has a weak influence on computational reproducibility and therefore defaults to Linux-based Rocker base images [@boettiger:2017:IR].
 In this way, it fixes Component A.
@@
 
-For R >= 3.1, the images from the Rocker project are used [@boettiger:2017:IR].
-For R < 3.1 but >= 1.3.1, a custom image based on Debian is used.
+For R >= 3.1, the base images from the Rocker project are used [@boettiger:2017:IR].
+For R < 3.1 but >= 1.3.1, a custom base image based on Debian is used.
 As of writing, `rang` does not support R < 1.3.1, i.e.
@@
 1.
-By default, the container building process downloads source packages from their sources and then compiles them.
+By default, the Docker image building process downloads source packages from their sources and then compiles them.
 This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) [^1], Bioconductor, and Github.
@@
 
-For the sake of completeness, the instructions for building and running the Docker container on Unix-like systems are included here.
+For the sake of completeness, the instructions for building the Docker image and running the Docker container on Unix-like systems are included here.
 
@@
 ## might need to sudo
-docker build -t rang .
+docker build -t rangimg .
 ## interactive environment
-docker run --rm --name "rangtest" -ti rang
+docker run --rm --name "rangcontainer" -ti rangimg
 ```
@@
 This can be regenerated by running `make export`.
-Preserving this file insures against all the situations mentioned above, but also the situations of Docker Hub and the software repositories used by the dockerized operating system being not available.
+Preserving this file insures against all the situations mentioned above, but also the situations of Docker Hub (the hosting service provided by Docker for base images such as Rocker) and the software repositories used by the dockerized operating system being not available.
 When `oserimg.tar.gz` is available, it is possible to run `make rebuild` and `make render` even without internet access (provided that Docker and `make` have been installed before).

### 2023-03-10 paper/paper.qmd
Fix all in-text citations for PLOS One #111

@@
 Although tutorials are available [e.g.
-@nuest:2019], providing a declarative description of the computational environment in the form of Dockerfile is far from the standard code sharing practice.
-This might be due to a lack of (DevOps) skills of most scientists to create a Dockerfile [@kim:2018:E].
+@nuest:2019], providing a declarative description of the computational environment in the form of `Dockerfile` is far from the standard code sharing practice.
+This might be due to a lack of (DevOps) skills of most scientists to create a `Dockerfile` [@kim:2018:E].
 But there are many tools available to automate the process [e.g.
@@
 
-`containerit` [@nuest:2019] takes the current state of the computational environment and documents it as a Dockerfile.
+`containerit` [@nuest:2019] takes the current state of the computational environment and documents it as a `Dockerfile`.
 `containerit` makes the assumption that Component A has a weak influence on computational reproducibility and therefore defaults to Linux-based Rocker base images [@boettiger:2017:IR].
@@
 
- @cruewell:2023:WB evaluate the computational reproducibility of 14 articles published in *Psyhocological Science*.
-Among these articles, the paper by @hilgard:2019:NEG has been rated as having "package dependency issues".
+Crüwell et al.
+[@cruewell:2023:WB] evaluate the computational reproducibility of 14 articles published in *Psyhocological Science*.
+Among these articles, the paper by Hilgard et al.
+[@hilgard:2019:NEG] has been rated as having "package dependency issues".
 
@@
 
-When running `dockerize()`, one can take advantage of the `materials_dir` parameter to transfer the shared materials from @hilgard:2019:NEG into the Docker image.
+When running `dockerize()`, one can take advantage of the `materials_dir` parameter to transfer the shared materials from Hilgard et al. [@hilgard:2019:NEG] into the Docker image.
 
@@
 
-All R scripts ran fine inside the container and the figures generated are the same as the ones in @hilgard:2019:NEG.
+All R scripts ran fine inside the container and the figures generated are the same as the ones in Hilgard et al. [@hilgard:2019:NEG].
 
@@
 
-The study by @trisovic:2022 evaluates the reproducibility of R scripts shared on Dataverse.
+The study by Trisovic et al.
+[@trisovic:2022] evaluates the reproducibility of R scripts shared on Dataverse.
 They found that 75\% of R scripts cannot be successfully executed.
-Among these failed R scripts is an R script shared by @beck:2019:EGD.
+Among these failed R scripts is an R script shared by Beck [@beck:2019:EGD].
 
@@
 
-Using `rang`, we are able to reconstruct a computational environment with R 2.15.0 (2012-03-30) to run all code snippets published in @jurka:2012 [^speed].
+Using `rang`, we are able to reconstruct a computational environment with R 2.15.0 (2012-03-30) to run all code snippets published in Jurka [@jurka:2012] [^speed].
 For removed CRAN packages, we strongly recommend querying the Github read-only mirror of CRAN instead (https://github.com/cran).
@@
 before 2004-10-04).
-Similar to the case of `maxent` above, a Dockerfile to assemble a Docker image with `ptproc` installed can be generated with two lines of code.
+Similar to the case of `maxent` above, a `Dockerfile` to assemble a Docker image with `ptproc` installed can be generated with two lines of code.
 
@@
 
-Suppose we have an R script, extracted from @peng:2003:MDP, called "peng.R" like this:
+Suppose we have an R script, extracted from Peng [@peng:2003:MDP], called "peng.R" like this:
 
@@
 [^random]: It is also important to note that the random number generator (RNG) of R has been changed several times over the course of the development.
-In this case, we are using the same generation of RNG as @peng:2003:MDP.
+In this case, we are using the same generation of RNG as Peng [@peng:2003:MDP].
 
@@
 `rang` has native support for Bioconductor packages since version 0.2.
-We obtained the R script `"PaperFigure.R"` from the Github repository of `Sushi` [^sushi], which generates the figure in @phanstiel:2014:S.
+We obtained the R script `"PaperFigure.R"` from the Github repository of `Sushi` [^sushi], which generates the figure in the original paper [@phanstiel:2014:S].
 Similar to the above case of `ptproc`, we made a completely automated BASH script to run `"PaperFigure.R"` and get the generated figure out of the container (@fig-figure2).
@@
 
-To demonstrate this point, we took the recent paper by @oser:2022:HPE.
+To demonstrate this point, we took the recent paper by Oser et al.
+[@oser:2022:HPE].
 This paper was selected because 1) the paper was published in *Political Communication*, a high impact journal that awards Open Science Badges; 2) shared data and R code are available; and most importantly, 3) the shared R code is well-written.
-In the repository of this paper, we based on the materials shared by @oser:2022:HPE and prepared a research compendium that should have long-term computational reproducibility.
+In the repository of this paper, we based on the materials shared by Oser et al.
+[@oser:2022:HPE] and prepared a research compendium that should have long-term computational reproducibility.
 The research compendium is similar to the Executable Compendium suggested by the Turing way.
@@
 
-With this `Makefile`, one can create the Dockerfile with `make resolve`, build the Docker image with `make build`, render the RMarkdown file inside the container with `make render`, export the built Docker image with `make export`, and rebuild the exported Docker image with `make rebuild`.
+With this `Makefile`, one can create the `Dockerfile` with `make resolve`, build the Docker image with `make build`, render the RMarkdown file inside the container with `make render`, export the built Docker image with `make export`, and rebuild the exported Docker image with `make rebuild`.
 
@@
 In this executable compendium, only the first four elements are essential.
-The directory `oserdocker` (116 MB) contains cached R packages, a Dockerfile, and a verbatim copy of the directory `meta-analysis/` to be transferred into the Docker image.
+The directory `oserdocker` (116 MB) contains cached R packages, a `Dockerfile`, and a verbatim copy of the directory `meta-analysis/` to be transferred into the Docker image.
 That can be regenerated by running `make resolve`.

### 2023-03-10 paper/paper.qmd
Add postprocessing to clean up the paper #111

@@
     affiliation: GESIS Leibniz-Institut für Sozialwissenschaften
+    email: chung-hong.chan@gesis.org
+    corresponding: true
   - name: "David Schoch"
@@
   pdf: default
-  arxiv-pdf:
+  plos-pdf:
     keep-tex: true
+    journal:
+      id: plosone
 ---

### 2023-03-10 paper/paper.qmd
Remove figures #111

@@
 
-The above code snippet can be executed with the generated container without any problem (@fig-figure1).
+The above code snippet can be executed with the generated container without any problem `Fig~\ref{fig1}`{=latex}.
 
-```{r}
-#| fig.cap: The code snippet running in a R 3.5.1 container created with rang
-#| echo: FALSE
-#| label: fig-figure1
-knitr::include_graphics("quanteda_rstudio.png", dpi = 300)
+<!-- ```{r} -->
+<!-- #| fig.cap: The code snippet running in a R 3.5.1 container created with rang -->
+<!-- #| echo: FALSE -->
+<!-- #| label: fig-figure1 -->
+<!-- knitr::include_graphics("quanteda_rstudio.png", dpi = 300) -->
+<!-- ``` -->
+
+```{=tex}
+% Place figure captions after the first paragraph in which they are cited.
+\begin{figure}[!h]
+\caption{The code snippet running in a R 3.5.1 container created with rang.}
+\label{fig1}
+\end{figure}
 ```
@@
 We obtained the R script `"PaperFigure.R"` from the Github repository of `Sushi` [^sushi], which generates the figure in the original paper [@phanstiel:2014:S].
-Similar to the above case of `ptproc`, we made a completely automated BASH script to run `"PaperFigure.R"` and get the generated figure out of the container (@fig-figure2).
+Similar to the above case of `ptproc`, we made a completely automated BASH script to run `"PaperFigure.R"` and get the generated figure out of the container (`Fig~\ref{fig2}`{=latex}).
 We made no modification to `"PaperFigure.R"`.
@@
 
-```{r}
-#| fig.cap: The figure from the batch execution of `PaperFigure.R` inside a Docker container generated by `rang`
-#| echo: FALSE
-#| label: fig-figure2
-knitr::include_graphics("sushi_figure1.pdf", dpi = 300)
+<!-- ```{r} -->
+<!-- #| fig.cap: The figure from the batch execution of `PaperFigure.R` inside a Docker container generated by `rang` -->
+<!-- #| echo: FALSE -->
+<!-- #| label: fig-figure2 -->
+<!-- knitr::include_graphics("sushi_figure1.pdf", dpi = 300) -->
+<!-- ``` -->
+
+```{=tex}
+% Place figure captions after the first paragraph in which they are cited.
+\begin{figure}[!h]
+\caption{The figure from the batch execution of PaperFigure.R inside a Docker container generated by rang}
+\label{fig2}
+\end{figure}
 ```

### 2023-03-15 paper/paper.qmd
Make Makefile more idiomatic
add slight update

@@
 Therefore, it can pin down components A, B, C, but only a part of component D.
-`dockta` is another containerization solution that can potentially pin down all components due to the fact that MRAN is used.
+`dockta` [^dockta] is another containerization solution that can potentially pin down all components due to the fact that MRAN is used.
 But it also suffers from the same limitations mentioned above.
@@
 [^posit]: [https://packagemanager.rstudio.com/client/#/repos/2/packages/](https://packagemanager.rstudio.com/client/#/repos/2/packages/)
+
+[^dockta]: [https://stencila.github.io/dockta/](https://stencila.github.io/dockta/)
 
@@
 
+.PHONY: all resolve build render export rebuild
+
 all: resolve build render

### 2023-03-20 R/cache.R
Add tests for caching R src and Debian

@@ -111,9 +111,7 @@
             ## please note that these cached packages are not built
             .cache_pkg_local(x = x, version = version, cache_dir = cache_dir, uid = uid)
         }
-
     }
-    ## For #14, cache R source in the future here
     invisible(output_dir)
 }
 

### 2023-03-20 tests/testthat/test_resolve.R

@@ -104,6 +104,11 @@ test_that("cache for R < 3.1 and R >= 2.1", {
     expect_true(file.exists(file.path(temp_dir, "cache/rpkgs", "foreign_0.8-54.tar.gz")))
     expect_true(file.exists(file.path(temp_dir, "cache/rpkgs", "evaluate_0.4.7.tar.gz")))
     expect_true(file.exists(file.path(temp_dir, "cache/rpkgs", "testthat_0.7.1.tar.gz")))
+    ## Debian & rsrc
+    expect_true(dir.exists(file.path(temp_dir, "cache/rsrc")))
+    expect_true(file.exists(file.path(temp_dir, "cache/rsrc", "R-3.0.1.tar.gz")))
+    expect_true(dir.exists(file.path(temp_dir, "cache/debian")))
+    expect_true(file.exists(file.path(temp_dir, "cache/debian", "rootfs.tar.xz")))
 })
 
 test_that("github correct querying; also #25", {
@@ -331,6 +336,12 @@ test_that("skip_r17", {
     x <- readLines(file.path(temp_dir, "Dockerfile"))
     expect_false(any(grepl("^RUN bash compile_r\\.sh 1\\.8\\.0", x)))
     expect_true(any(grepl("^RUN bash compile_r\\.sh 1\\.7\\.0", x)))
+    ## debian & rsrc
+    expect_true(dir.exists(file.path(temp_dir, "cache/rsrc")))
+    expect_true(file.exists(file.path(temp_dir, "cache/rsrc", "R-1.7.0.tgz")))
+    expect_true(dir.exists(file.path(temp_dir, "cache/debian")))
+    expect_true(file.exists(file.path(temp_dir, "cache/debian", "rootfs.tar.xz")))
+
 })
 
 ## always keep this at the very last

### 2023-04-13 inst/turing/paper.Rmd
Add a template of Executable Compendium

---
title: "Executable Compendium (The Turing Way)"
author: Garry Gilmore, Elena Emerson, Sebastian Stewart, Isabel Ingram, Sophie Sasaki
output: pdf_document
bibliography: "bibliography.bib"
---

```{r setup, include = FALSE}
library(here)
library(knitr)
```

# Ipsum

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vivamus tempus quis justo vel elementum. Ut arcu tortor, pharetra ac mattis vel, scelerisque in metus. Morbi iaculis ante eros, sit amet iaculis dolor suscipit non. Duis lectus urna, laoreet ut nisi quis, ultricies condimentum ipsum. Nullam aliquam id ipsum eget maximus. Pellentesque vulputate felis augue, eget interdum mauris iaculis et. Mauris maximus eros dolor, ut fermentum augue suscipit faucibus.

Vivamus nulla turpis, condimentum eget urna in, fringilla auctor sapien. Pellentesque sodales lacus ac consequat sodales. In hac habitasse platea dictumst. Praesent ut elit est. Nullam quis quam dignissim, vehicula massa et, dignissim nisl. Ut commodo interdum dapibus. Integer ac sem aliquam, sollicitudin erat mollis, blandit arcu. Vestibulum finibus nisi id tincidunt consequat. Vestibulum varius massa ac dui aliquet condimentum. Duis vel mollis ipsum. Curabitur aliquet diam sit amet dignissim luctus. Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Phasellus massa risus, molestie quis rhoncus quis, varius sit amet elit.

Sed ultrices neque eget magna sagittis, et consectetur nibh finibus. Aliquam sit amet convallis massa. Sed accumsan nunc et diam lacinia feugiat. Vivamus dapibus augue vestibulum felis vestibulum ornare. Curabitur sodales leo at orci gravida, a varius ex consectetur. Maecenas in tellus sed odio varius volutpat. In vestibulum tellus vitae sapien convallis, at facilisis nibh posuere. Sed dictum leo scelerisque nisl dignissim, a tristique risus rutrum.

Suspendisse scelerisque facilisis augue, a aliquet augue tempor vel. Quisque ac vulputate sem. Nulla facilisi. Donec sagittis ante vitae enim sollicitudin dictum. Aenean in luctus nisi, ac mollis erat. Curabitur pellentesque risus consequat ante volutpat, non auctor nisi molestie. Fusce sit amet lacinia arcu, non euismod metus. Suspendisse fermentum sit amet felis sed ultrices. Suspendisse potenti. Quisque gravida purus mi, nec laoreet massa sodales in. Quisque tincidunt eu nibh vitae convallis. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.

Mauris tristique mauris et imperdiet imperdiet. Proin aliquet metus nec ligula egestas sagittis. Nulla metus lorem, cursus ut ex a, aliquam condimentum erat. Cras eget augue in est euismod varius. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia curae; Vivamus eget justo metus. Integer volutpat interdum felis blandit laoreet. Quisque et sagittis est, quis semper risus.

# Analysis

This is a visualization of the `palmerpenguins` dataset [@horst2020; @gorman2014].

```{r flipper, echo = FALSE, fig.cap = 'Flipper and bill length'}
knitr::include_graphics(here("figures", "flipper_bil.png"))
```

# Ipsum

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vivamus tempus quis justo vel elementum. Ut arcu tortor, pharetra ac mattis vel, scelerisque in metus. Morbi iaculis ante eros, sit amet iaculis dolor suscipit non. Duis lectus urna, laoreet ut nisi quis, ultricies condimentum ipsum. Nullam aliquam id ipsum eget maximus. Pellentesque vulputate felis augue, eget interdum mauris iaculis et. Mauris maximus eros dolor, ut fermentum augue suscipit faucibus.

Vivamus nulla turpis, condimentum eget urna in, fringilla auctor sapien. Pellentesque sodales lacus ac consequat sodales. In hac habitasse platea dictumst. Praesent ut elit est. Nullam quis quam dignissim, vehicula massa et, dignissim nisl. Ut commodo interdum dapibus. Integer ac sem aliquam, sollicitudin erat mollis, blandit arcu. Vestibulum finibus nisi id tincidunt consequat. Vestibulum varius massa ac dui aliquet condimentum. Duis vel mollis ipsum. Curabitur aliquet diam sit amet dignissim luctus. Orci varius natoque penatibus et magnis dis parturient montes, nascetur ridiculus mus. Phasellus massa risus, molestie quis rhoncus quis, varius sit amet elit.

Sed ultrices neque eget magna sagittis, et consectetur nibh finibus. Aliquam sit amet convallis massa. Sed accumsan nunc et diam lacinia feugiat. Vivamus dapibus augue vestibulum felis vestibulum ornare. Curabitur sodales leo at orci gravida, a varius ex consectetur. Maecenas in tellus sed odio varius volutpat. In vestibulum tellus vitae sapien convallis, at facilisis nibh posuere. Sed dictum leo scelerisque nisl dignissim, a tristique risus rutrum.

Suspendisse scelerisque facilisis augue, a aliquet augue tempor vel. Quisque ac vulputate sem. Nulla facilisi. Donec sagittis ante vitae enim sollicitudin dictum. Aenean in luctus nisi, ac mollis erat. Curabitur pellentesque risus consequat ante volutpat, non auctor nisi molestie. Fusce sit amet lacinia arcu, non euismod metus. Suspendisse fermentum sit amet felis sed ultrices. Suspendisse potenti. Quisque gravida purus mi, nec laoreet massa sodales in. Quisque tincidunt eu nibh vitae convallis. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.

Mauris tristique mauris et imperdiet imperdiet. Proin aliquet metus nec ligula egestas sagittis. Nulla metus lorem, cursus ut ex a, aliquam condimentum erat. Cras eget augue in est euismod varius. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia curae; Vivamus eget justo metus. Integer volutpat interdum felis blandit laoreet. Quisque et sagittis est, quis semper risus.

# References

### 2023-04-14 R/apptainer.R
remove debian image cache for apptainer, no easy way to do it, better to bootrstrap from docker image

@@ -25,11 +25,12 @@
     apptainer_content$POST <- append(apptainer_content$POST, paste0("bash $COMPILE_PATH ", r_version))
   }
   if (isTRUE(cache)) {
+    apptainer_content$BOOTSTRAP <- "Bootstrap: docker"
+    apptainer_content$FROM <- c(paste0("From: debian/eol:", debian_version))
     apptainer_content$FILES <- append(apptainer_content$FILES,
                                         c(paste0("cache/rpkgs ", file.path(cache_path, "rpkgs")),
                                           paste0("cache/rsrc ", file.path(cache_path, "rsrc"))))
-    apptainer_content$FROM <- c("From: scratch", paste0("ADD ", file.path(rel_dir, "cache/debian/rootfs.tar.xz"), " /"))
-    apptainer_content$ENV <- append(apptainer_content$ENV, paste0("CACHE_PATH=", cache_path))
+    apptainer_content$POST <- prepend(apptainer_content$POST, paste0("export CACHE_PATH=", cache_path))
   }
   apptainer_content$POST <- append(apptainer_content$POST, post_installation_steps)
   if (isTRUE(copy_all)) {
@@ -44,15 +45,16 @@
                                                  copy_all = FALSE) {
   rang_path <- file.path(rel_dir, "rang.R")
   cache_path <- file.path(rel_dir, "cache")
+  environment_vars <- c(paste0("export RANG_PATH=", rang_path))
   apptainer_content <- list(
     BOOTSTRAP = "Bootstrap: docker",
     FROM = c(paste0("From: rocker/", image, ":", r_version)),
     ENV_section = "\n%environment\n",
-    ENV = c(paste0("export RANG_PATH=", rang_path)),
+    ENV = environment_vars,
     FILES_section = "\n%files\n",
     FILES = c(paste0("rang.R ", rang_path)),
     POST_section = "\n%post\n",
-    POST = sysreqs_cmd,
+    POST = c(environment_vars, sysreqs_cmd),
     STARTSCRIPT_section = "\n%startscript\n",
     STARTSCRIPT = c("exec R \"${@}\""))
   if (!is.na(lib)) {

### 2023-04-14 tests/testthat/test_apptainerize.R
fix tests for apptainer with custom post install

@@ -56,7 +56,8 @@ test_that("integration of #16 in apptainerize()", {
     ## post
     apptainerize(rang = rang_ok, output_dir = temp_dir, post_installation_steps = "date")
     container_file <- readLines(file.path(temp_dir, "container.def"))
-    expect_equal(container_file[length(container_file) -2], "date")
+    last_post_line_before_start_script <- which(grepl("\\%startscript", container_file)) -2
+    expect_equal(container_file[last_post_line_before_start_script], "date")
 })
 
 test_that("integration of #18 in apptainerize()", {
@@ -143,7 +144,8 @@ test_that("apptainerize R < 3.1 and >= 2.1", {
     expect_equal(tail(container_file, 1), "exec R \"${@}\"")
     apptainerize(rang_rio, output_dir = temp_dir, lib = "abc", post_installation_step = "date")
     container_file <- readLines(file.path(temp_dir, "container.def"))
-    expect_equal(container_file[length(container_file) -2], "date")
+    last_post_line_before_start_script <- which(grepl("\\%startscript", container_file)) -2
+    expect_equal(container_file[last_post_line_before_start_script], "date")
 })
 
 test_that("Docker R < 1.3.1", {

### 2023-04-14 README.Rmd
Update README and FAQ vignette with info about Apptainer/Singularity

@@
 
+### Using Apptainer/Singularity containers for HPC
+
+A `rang` object can be used to recreate the computational environment via [Rocker](https://github.com/rocker-org/rocker).
+Instead of Docker you can also use [Apptainer/Singularity](https://apptainer.org/).
+Please note that the oldest R version one can get from Rocker is R 3.1.0.
+
+```r
+apptainerize(graph, "~/rocker_test")
+# singularize(graph, "~/rocker_test") # same function, as so far Apptainer is identical to Singularity
+```
+
+Now, you can build and run the Apptainer/Singularity container.
+
+For Apptainer installation:
+
+```bash
+cd ~/rocker_test
+sudo apptainer build container.sif container.def
+apptainer instance start container.sif rangtest
+```
+
+For Singularity installation:
+
+```bash
+cd ~/rocker_test
+sudo singularity build container.sif container.def
+singularity exec container.sif R
+```
+
+Using the above example, `sessionInfo()` outputs the following. You have successfully gone back to the pre-pandemic time on your HPC.
+
+```
+R version 3.6.2 (2019-12-12) -- "Dark and Stormy Night"
+Copyright (C) 2019 The R Foundation for Statistical Computing
+Platform: x86_64-pc-linux-gnu (64-bit)
+
+R is free software and comes with ABSOLUTELY NO WARRANTY.
+You are welcome to redistribute it under certain conditions.
+Type 'license()' or 'licence()' for distribution details.
+
+R is a collaborative project with many contributors.
+Type 'contributors()' for more information and
+'citation()' on how to cite R or R packages in publications.
+
+Type 'demo()' for some demos, 'help()' for on-line help, or
+'help.start()' for an HTML browser interface to help.
+Type 'q()' to quit R.
+
+> sessionInfo()
+R version 3.6.2 (2019-12-12)
+Platform: x86_64-pc-linux-gnu (64-bit)
+Running under: Debian GNU/Linux 10 (buster)
+
+Matrix products: default
+BLAS/LAPACK: /usr/lib/x86_64-linux-gnu/libopenblasp-r0.3.5.so
+
+locale:
+ [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C
+ [3] LC_TIME=en_US.UTF-8        LC_COLLATE=en_US.UTF-8
+ [5] LC_MONETARY=en_US.UTF-8    LC_MESSAGES=C
+ [7] LC_PAPER=en_US.UTF-8       LC_NAME=C
+ [9] LC_ADDRESS=C               LC_TELEPHONE=C
+[11] LC_MEASUREMENT=en_US.UTF-8 LC_IDENTIFICATION=C
+
+attached base packages:
+[1] stats     graphics  grDevices utils     datasets  methods   base
+
+loaded via a namespace (and not attached):
+[1] compiler_3.6.2
+```
+
+`apptainerize()`/`singularize()` functions work exactly the same as `dockerize()`, expect you cannot cache Linux distribution rootfs.
+
 ## Recreate the computational environment for R < 3.1.0

### 2023-04-14 R/installation.R
fix typos in error and warnings messages

@@ -509,17 +509,17 @@ apptainerize <- function(rang, output_dir, materials_dir = NULL, post_installati
                          insert_readme = TRUE,
                          copy_all = FALSE) {
   if (length(rang$ranglets) == 0) {
-    warning("Nothing to appteinerize/singularize.")
+    warning("Nothing to apptainerize/singularize.")
     return(invisible(NULL))
   }
   if (missing(output_dir)) {
     stop("You must provide `output_dir`.", call. = FALSE)
   }
   if (!grepl("^ubuntu", rang$os)) {
-    stop("System dependencies of ", rang$os, " can't be appteinerized/singularized.", call. = FALSE)
+    stop("System dependencies of ", rang$os, " can't be apptainerized/singularized.", call. = FALSE)
   }
   if (.is_r_version_older_than(rang, "1.3.1")) {
-    stop("`appteinerize/singularize` doesn't support this R version (yet):", rang$r_version, call. = FALSE)
+    stop("`apptainerize/singularize` doesn't support this R version (yet):", rang$r_version, call. = FALSE)
   }
   if (!is.null(materials_dir) && !(dir.exists(materials_dir))) {
     stop(paste0("The folder ", materials_dir, " does not exist"), call. = FALSE)

### 2023-04-14 tests/testthat/test_apptainerize.R
add tests for new version of .insert_materials_dir

@@ -317,3 +317,14 @@ test_that("apptainerize with inst/rang", {
     expect_true(". /" %in% container_file) ## coerced
     expect_true("export RANG_PATH=inst/rang/rang.R" %in% container_file)
 })
+
+test_that(".insert_materials_dir with actual outcome", {
+  container_content <- list()
+  container_content$COPY <- "COPY rang.R /rang.R"
+  expect_identical(names(.insert_materials_dir(container_content, container_type = "docker")), "COPY")
+  expect_false(names(.insert_materials_dir(container_content, container_type = "docker")) %in% "FILES")
+  container_content <- list()
+  container_content$FILES <- "rang.R /rang.R"
+  expect_identical(names(.insert_materials_dir(container_content, container_type = "apptainer")), "FILES")
+  expect_false(names(.insert_materials_dir(container_content, container_type = "apptainer")) %in% "COPY")
+})

### 2023-04-14 R/installation.R
add reference to apptainerize() in export_rang() docs

@@ -146,7 +146,7 @@
 #' @param bioc_mirror character, which Bioconductor mirror to use
 #' @return `path`, invisibly
 #' @details The idea behind this is to determine the installation order of R packages locally. Then, the installation script can be deployed to another
-#' fresh R session to install R packages. [dockerize()] is a more reasonable way because a fresh R session with all system requirements
+#' fresh R session to install R packages. [dockerize()] and [apptainerize()] are more reasonable ways because a fresh R session with all system requirements
 #' is provided. The current approach does not work in R < 2.1.0.
 #' @export
 #' @references

### 2023-04-14 R/apptainer.R
(1) make 4 spaces indent; (2) rewrite .generate_docker_readme() to generate_container_readme() to work with both types of containers; (3) apptainerize() inherits duplicate params from dockerize()

@@ -1,78 +1,85 @@
 .generate_debian_eol_apptainer_content <- function(r_version, lib, sysreqs_cmd, cache, debian_version = "lenny",
-                                                     post_installation_steps = NULL,
-                                                     rel_dir = "",
-                                                     copy_all = FALSE) {
-  rang_path <- file.path(rel_dir, "rang.R")
-  cache_path <- file.path(rel_dir, "cache")
-  compile_path <- file.path(rel_dir, "compile_r.sh")
-  environment_vars <- c("export TZ=UTC", paste0("export COMPILE_PATH=", compile_path), paste0("export RANG_PATH=", rang_path))
-  apptainer_content <- list(
-    BOOTSTRAP = "Bootstrap: docker",
-    FROM = c(paste0("From: debian/eol:", debian_version)),
-    ENV_section = "\n%environment\n",
-    ENV = environment_vars,
-    FILES_section = "\n%files\n",
-    FILES = c(paste0("rang.R ", rang_path), paste0("compile_r.sh ", compile_path)),
-    POST_section = "\n%post\n",
-    POST = c(environment_vars,
-             "ln -snf /usr/share/zoneinfo/$TZ /etc/localtime && echo $TZ > /etc/timezone && apt-get update -qq && apt-get install wget locales build-essential r-base-dev -y",
-             sysreqs_cmd),
-    STARTSCRIPT_section = "\n%startscript\n",
-    STARTSCRIPT = c("exec R \"${@}\""))
-  if (!is.na(lib)) {
-    apptainer_content$POST <- append(apptainer_content$POST, paste0("mkdir ", lib, " && bash $COMPILE_PATH ", r_version))
-  } else {
-    apptainer_content$POST <- append(apptainer_content$POST, paste0("bash $COMPILE_PATH ", r_version))
-  }
-  if (isTRUE(cache)) {
-    apptainer_content$BOOTSTRAP <- "Bootstrap: docker"
-    apptainer_content$FROM <- c(paste0("From: debian/eol:", debian_version))
-    apptainer_content$FILES <- append(apptainer_content$FILES,
-                                        c(paste0("cache/rpkgs ", file.path(cache_path, "rpkgs")),
-                                          paste0("cache/rsrc ", file.path(cache_path, "rsrc"))))
-    apptainer_content$POST <- prepend(apptainer_content$POST, paste0("export CACHE_PATH=", cache_path))
-  }
[…]

### 2023-04-14 R/installation.R

@@ -114,10 +114,16 @@
     utils::compareVersion(rang$r_version, r_version) == -1
 }
 
-.generate_docker_readme <- function(output_dir,image) {
+.generate_container_readme <- function(output_dir, image, container_type = c("docker", 'apptainer')) {
     file.create(file.path(output_dir,"README"))
     con <- file(file.path(output_dir,"README"), open="w")
-    readme <- readLines(system.file("readme_template.txt", package = "rang"))
+    container_type <- match.arg(container_type)
+    if (container_type == "docker") {
+        readme <- readLines(system.file("docker_readme_template.txt", package = "rang"))
+    }
+    if (container_type == "apptainer") {
+        readme <- readLines(system.file("apptainer_readme_template.txt", package = "rang"))
+    }
     readme <- gsub("__DATE__",Sys.Date(),readme)
     readme <- gsub("__OUTPUT__",output_dir,readme)
     readme <- gsub("__IMAGE__",image,readme)
@@ -414,7 +420,7 @@ dockerize <- function(rang, output_dir, materials_dir = NULL, post_installation_
     ## This should be written in the root level, not base_dir
     .write_container_file(dockerfile_content, file.path(output_dir, "Dockerfile"))
     if (isTRUE(insert_readme)) {
-        .generate_docker_readme(output_dir = output_dir, image = image)
+        .generate_container_readme(output_dir = output_dir, image = image, container_type = "docker")
     }
     invisible(output_dir)
 }
@@ -438,22 +444,21 @@ dockerise_rang <- function(...) {
 }
 
 .insert_materials_dir <- function(container_content, container_type = c("docker", "apptainer")) {
-  container_type <- match.arg(container_type)
-  if (container_type == "docker") {
-    container_content$COPY <- append(container_content$COPY, "COPY materials/ ./materials/")
-    return(container_content)
-  }
-  if (container_type == "apptainer") {
-    container_content$FILES <- append(container_content$FILES, "materials/ ./materials/")
-    return(container_content)
-  }
+    container_type <- match.arg(container_type)
+    if (container_type == "docker") {
+        container_content$COPY <- append(container_content$COPY, "COPY materials/ ./materials/")
+        return(container_content)
+    }
+    if (container_type == "apptainer") {
[…]

### 2023-04-14 R/installation.R
update docs

@@ -467,7 +467,7 @@ dockerise_rang <- function(...) {
 #' @param output_dir character, where to put the Apptainer/Singularity definition file and associated content
 #' @param materials_dir character, path to the directory containing additional resources (e.g. analysis scripts) to be copied into `output_dir` and in turn into the Apptainer/Singularity container
 #' @param post_installation_steps character, additional steps to be added before the in the end of `%post` section the Apptainer/Singularity definition file, see an example below
-#' @inheritParams source
+#' @inheritParams dockerize
 #' @param ... arguments to be passed to `apptainerize`
 #' @return `output_dir`, invisibly
 #' @inheritParams export_rang

### 2023-04-17 R/apptainer.R
fix Apptainer/Singularity startscript, as /init uses s6 overlay seemingly incompatible with Apptainer/Singularity

@@ -75,7 +75,9 @@
         apptainer_content$ENV <- append(apptainer_content$ENV, paste0("export CACHE_PATH ", cache_path))
     }
     if (image == "rstudio") {
-        apptainer_content$STARTSCRIPT <- c("exec /init \"$@\"")
+        apptainer_content$STARTSCRIPT <- c("exec /usr/lib/rstudio-server/bin/rserver \
+    --auth-none=0 --auth-pam-helper-path=pam-helper \
+    --server-user=$(whoami)")
     }
     apptainer_content$POST <- append(apptainer_content$POST, post_installation_steps)
     if (isTRUE(copy_all)) {

### 2023-04-17 README.Rmd
update readme, remove HPC references
Update README.md/Rmd and faq.Rmd with correct instructions to start RStudio IDE in Apptainer/Singularity

@@
 
-### Using Apptainer/Singularity containers for HPC
+### Using Apptainer/Singularity containers
 
@@
 cd ~/rocker_test
-sudo apptainer build container.sif container.def
-apptainer instance start container.sif rangtest
+apptainer build container.sif container.def
+apptainer run container.sif R
 ```
@@
 sudo singularity build container.sif container.def
-singularity exec container.sif R
+singularity run container.sif R
 ```
 
-Using the above example, `sessionInfo()` outputs the following. You have successfully gone back to the pre-pandemic time on your HPC.
+Using the above example, `sessionInfo()` outputs the following. You have successfully gone back to the pre-pandemic.
 
@@
 
+### Apptainer/Singularity with RStudio IDE
+
+For running RStudio IDE in Apptainer/Singulairty container, some writable folders and a config file have to be created locally:
+
+```bash
+mkdir -p run var-lib-rstudio-server
+printf 'provider=sqlite\ndirectory=/var/lib/rstudio-server\n' > database.conf
+```
+
+After that, you can run the container:
+
+```bash
+apptainer exec \
+    --env PASSWORD='set_your_password' \
+    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf \
+    container.sif \
+    /usr/lib/rstudio-server/bin/rserver \
+    --auth-none=0 --auth-pam-helper-path=pam-helper \
+    --server-user=$(whoami)
+```
+
+or
+
+```bash
+singularity exec \
+    --env PASSWORD='set_your_password' \
+    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf \
+    container.sif \
+    /usr/lib/rstudio-server/bin/rserver \
+    --auth-none=0 --auth-pam-helper-path=pam-helper \
+    --server-user=$(whoami)
+```
+
+The default port is 8787, you can also change the port by adding `--www-port=` in the end of the line above (e.g. `--www-port=8080`).
+
+Now open a browser and go to localhost:8787.
+The default username is your local username, password as specified above (in this case `set_your_password`).
+
 ## Recreate the computational environment for R < 3.1.0

### 2023-04-17 R/installation.R
Refresh `cache` ref #131

@@ -288,6 +288,7 @@ export_renv <- function(rang, path = ".") {
 #' @param skip_r17 logical, whether to skip R 1.7.x. Currently, it is not possible to compile R 1.7.x (R 1.7.0 and R 1.7.1) with the method provided by  `rang`. It affects `snapshot_date` from 2003-04-16 to 2003-10-07. When `skip_r17` is TRUE and `snapshot_date` is within the aforementioned range, R 1.8.0 is used instead
 #' @param insert_readme logical, whether to insert a README file
 #' @param copy_all logical, whether to copy everything in the current directory into the container. If `inst/rang` is detected in `output_dir`, this is coerced to TRUE.
+#' @param refresh_cache logical, if `cache` is TRUE whether to delete existing cache directory before caching
 #' @param ... arguments to be passed to `dockerize`
 #' @return `output_dir`, invisibly
 #' @inheritParams export_rang
@@ -320,7 +321,8 @@ dockerize <- function(rang, output_dir, materials_dir = NULL, post_installation_
                       debian_version = c("lenny", "squeeze", "wheezy", "jessie", "stretch"),
                       skip_r17 = TRUE,
                       insert_readme = TRUE,
-                      copy_all = FALSE) {
+                      copy_all = FALSE,
+                      refresh_cache = TRUE) {
     if (length(rang$ranglets) == 0) {
         warning("Nothing to dockerize.")
         return(invisible(NULL))
@@ -372,6 +374,9 @@ dockerize <- function(rang, output_dir, materials_dir = NULL, post_installation_
         r_version <- "1.8.0"
     } else {
         r_version <- rang$r_version
+    }
+    if (isTRUE(cache) && isTRUE(refresh_cache) && dir.exists(file.path(base_dir, "cache"))) {
+        unlink(file.path(base_dir, "cache"), recursive = TRUE)
     }
     if (isTRUE(cache)) {
         .cache_pkgs(rang = rang, base_dir = base_dir, cran_mirror = cran_mirror,

### 2023-04-17 R/apptainer.R
remove deprecated purrr::prepend()

@@ -37,7 +37,7 @@
                 paste0("cache/rsrc ", file.path(cache_path, "rsrc"))
             )
         )
-        apptainer_content$POST <- prepend(apptainer_content$POST, paste0("export CACHE_PATH=", cache_path))
+        apptainer_content$POST <- append(paste0("export CACHE_PATH=", cache_path), apptainer_content$POST)
     }
     apptainer_content$POST <- append(apptainer_content$POST, post_installation_steps)
     if (isTRUE(copy_all)) {

### 2023-04-17 README.Rmd
add info that Apptainer/Singularity containers with RStudio IDE cannot be started using `root` user
update README: (1) compatibility with old RStudio IDE (<4.0) that requires binding .rstudio to the container (2) add default port parameter so that it is obvious how to change it

@@
 ```bash
-mkdir -p run var-lib-rstudio-server
+mkdir -p run var-lib-rstudio-server .rstudio
 printf 'provider=sqlite\ndirectory=/var/lib/rstudio-server\n' > database.conf
@@
 
-After that, you can run the container:
+After that, you can run the container (do not run as `root` user, otherwise you will not be able to login to RStudio IDE):
 
@@
     --env PASSWORD='set_your_password' \
-    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf \
+    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf,.rstudio:/home/rstudio/.rstudio/ \
     container.sif \
@@
     --auth-none=0 --auth-pam-helper-path=pam-helper \
-    --server-user=$(whoami)
+    --server-user=$(whoami) --www-port=8787
 ```
@@
     --auth-none=0 --auth-pam-helper-path=pam-helper \
-    --server-user=$(whoami)
+    --server-user=$(whoami) --www-port=8787
 ```
-
-The default port is 8787, you can also change the port by adding `--www-port=` in the end of the line above (e.g. `--www-port=8080`).
 

### 2023-04-17 R/apptainer.R
Fix RStudio Apptainer/Singularity instance start. Add ENV variables that are needed for start of the instance, but not the install process

@@ -57,7 +57,8 @@
         BOOTSTRAP = "Bootstrap: docker",
         FROM = c(paste0("From: rocker/", image, ":", r_version)),
         ENV_section = "\n%environment\n",
-        ENV = environment_vars,
+        ENV = c(environment_vars, "export RPORT=${RPORT:-8787}",
+                "export USER=$(whoami)", "export PASSWORD=${PASSWORD:-set_your_password}"),
         FILES_section = "\n%files\n",
         FILES = c(paste0("rang.R ", rang_path)),
         POST_section = "\n%post\n",
@@ -75,9 +76,9 @@
         apptainer_content$ENV <- append(apptainer_content$ENV, paste0("export CACHE_PATH ", cache_path))
     }
     if (image == "rstudio") {
-        apptainer_content$STARTSCRIPT <- c("exec /usr/lib/rstudio-server/bin/rserver \
-    --auth-none=0 --auth-pam-helper-path=pam-helper \
-    --server-user=$(whoami)")
+        apptainer_content$STARTSCRIPT <- c("exec /usr/lib/rstudio-server/bin/rserver \\\
+    --auth-none=0 --auth-pam-helper-path=pam-helper \\\
+    --server-user=${USER} --www-port=${RPORT}")
     }
     apptainer_content$POST <- append(apptainer_content$POST, post_installation_steps)
     if (isTRUE(copy_all)) {

### 2023-04-17 README.Rmd
Update README, readme template for Apptainer/Singularity and FAQ with instructions in running containers

@@
 
-For running RStudio IDE in Apptainer/Singulairty container, some writable folders and a config file have to be created locally:
+To run RStudio IDE in Apptainer/Singularity container, some writeable folders and a config file have to be created locally:
 
@@
 
-After that, you can run the container (do not run as `root` user, otherwise you will not be able to login to RStudio IDE):
+After that, you can run the container (do not run as `root` user, otherwise you will not be able to login to RStudio IDE).
+
+Start instance (on default RSTUDIO port 8787):
+
+```bash
+apptainer instance start \
+    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf,.rstudio:/home/rstudio/.rstudio/ \
+    container.sif \
+    rangtest
+```
+
+Now open a browser and go to localhost:8787.
+The default username is your local username, default password is 'set_your_password' (if you are using container generated by rang).
+
+List running instances:
+
+```bash
+apptainer instance list
+```
+
+Stop instance:
+
+```bash
+apptainer instance stop rangtest
+```
+
+Start instance with custom port (e.g. 8080) and password:
+
+```bash
+apptainer instance start \
+    --enc RPORT=8080
+    --env PASSWORD='set_your_password' \
+    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf,.rstudio:/home/rstudio/.rstudio/ \
+    container.sif \
+    rangtest
+```
+
+Run container with custom `rserver` command line:
 
@@
 
-or
-
-```bash
-singularity exec \
-    --env PASSWORD='set_your_password' \
-    --bind run:/run,var-lib-rstudio-server:/var/lib/rstudio-server,database.conf:/etc/rstudio/database.conf,.rstudio:/home/rstudio/.rstudio/ \
-    container.sif \
-    /usr/lib/rstudio-server/bin/rserver \
-    --auth-none=0 --auth-pam-helper-path=pam-helper \
-    --server-user=$(whoami) --www-port=8787
-```
-
-Now open a browser and go to localhost:8787.
-The default username is your local username, password as specified above (in this case `set_your_password`).
+If you run the container using `apptainer exec` command, you will have to kill the `rserver` process manually or Cmd/Ctrl+C from the running container to stop the server.
 

### 2023-04-17 tests/testthat/test_apptainerize.R
Update test for correct Apptainer/Singularity container for RStudio image

@@ -266,7 +266,8 @@ test_that("#123 rstudio", {
     temp_dir <- .generate_temp_dir()
     apptainerize(graph, output_dir = temp_dir, image = "rstudio")
     container_file <- readLines(file.path(temp_dir, "container.def"))
-    expect_equal(tail(container_file, 1), "exec /init \"$@\"")
+    first_startscript_line <- which(grepl("\\%startscript", container_file)) + 2
+    expect_true(all(grepl("exec /usr/lib/rstudio-server/bin/rserver", container_file[first_startscript_line])))
 })
 
 test_that("apptainerize with bioc #58", {

### 2023-04-17 R/apptainer.R
CACHE_PATH should be available during container build, not in the enviroment of a running container

@@ -73,7 +73,7 @@
     }
     if (isTRUE(cache)) {
         apptainer_content$FILES <- append(apptainer_content$FILES, paste0("cache ", cache_path))
-        apptainer_content$ENV <- append(apptainer_content$ENV, paste0("export CACHE_PATH ", cache_path))
+        apptainer_content$POST <- append(paste0("export CACHE_PATH=", cache_path), apptainer_content$POST)
     }
     if (image == "rstudio") {
         apptainer_content$STARTSCRIPT <- c("exec /usr/lib/rstudio-server/bin/rserver \\\

### 2023-04-17 tests/testthat/test_apptainerize.R
remove tests from test_appteinerize.R that have nothing to do with appteinerize() and are carbon copies of tests in tests_dockerize.R

@@ -159,19 +159,6 @@ test_that("Docker R < 1.3.1", {
     expect_error(apptainerize(rang_rio, output_dir = temp_dir))
 })
 
-test_that(".group_sysreqs and issue #21", {
-    graph <- readRDS("../testdata/graph.RDS")
-    expect_equal(.group_sysreqs(graph), "apt-get update -qq && apt-get install -y libpcre3-dev zlib1g-dev pkg-config libcurl4-openssl-dev && apt-get install -y default-jdk libgsl0-dev libicu-dev libpng-dev libxml2-dev make python3 zlib1g-dev liblzma-dev libpcre3-dev libbz2-dev && R CMD javareconf")
-    graph <- readRDS("../testdata/rang_ok.RDS")
-    expect_equal(.group_sysreqs(graph), "apt-get update -qq && apt-get install -y libpcre3-dev zlib1g-dev pkg-config libcurl4-openssl-dev")
-    graph <- readRDS("../testdata/issue21.RDS")
-    expected_output <- "apt-get update -qq && apt-get install -y libpcre3-dev zlib1g-dev pkg-config && apt-get install -y software-properties-common && add-apt-repository -y ppa:cran/libgit2 && apt-get update && apt-get install -y cmake git libcurl4-openssl-dev libfontconfig1-dev libfreetype6-dev libfribidi-dev libgit2-dev libgsl0-dev libharfbuzz-dev libicu-dev libjpeg-dev libpng-dev libssh2-1-dev libssl-dev libtiff-dev libxml2-dev make pandoc pari-gp zlib1g-dev"
-    expect_warning(output <- .group_sysreqs(graph))
-    expect_equal(output, expected_output)
-    graph <- readRDS("../testdata/issue21_ubuntu2004.RDS")
-    expect_warning(output <- .group_sysreqs(graph), NA)
-    expect_true(grepl("gnutls", output))
-})
 
 test_that("apptainerize warning, issue #21", {
     graph <- readRDS("../testdata/issue21.RDS")
@@ -296,13 +283,6 @@ test_that("no_rocker #67", {
     temp_dir <- .generate_temp_dir()
     expect_error(apptainerize(rang = rang_ok, output_dir = temp_dir, no_rocker = TRUE,
               debian_version = "3.11"))
-})
-
-test_that(".check_tarball_path", {
-    expect_error(.check_tarball_path("../testdata/gesis_2.0.tar.gz", "gesis")) ##dir = FALSE
[…]

### 2023-04-18 README.Rmd
fix an error in doc

@@
 
-`apptainerize()`/`singularize()` functions work exactly the same as `dockerize()`, expect you cannot cache Linux distribution rootfs.
+`apptainerize()`/`singularize()` functions work exactly the same as `dockerize()`, except you cannot cache Linux distribution rootfs.
 
@@
 apptainer instance start \
-    --enc RPORT=8080
+    --env RPORT=8080
     --env PASSWORD='set_your_password' \

### 2023-05-16 paper/paper.qmd
First pass

@@
 abstract: |
-  A complete declarative description of the computational environment is often missing when researchers share their materials.
+  A complete declarative description of the computational environment is usually missing when researchers share their materials.
 Without such description, software obsolescence and missing system components can jeopardize computational reproducibility in the future, even when data and computer code are available.
@@
 The declarative description generated by rang satisfies the definition of a reproducible research compendium and can be shared as such.
-In this contribution, we show how rang can be used to make otherwise unexecutable code, spanning from fields such as computational social science and bioinformatics, executable again.
+In this contribution, we show how rang can be used to make otherwise unexecutable code, spanning fields such as computational social science and bioinformatics, executable again.
 We also provide instructions on how to use rang to construct reproducible and shareable research compendia of current research.
@@
 But as computer science research has shown, this task is incredibly complex [@abate:2015:MCR;@dolstra:2010:N].
-For a usual scripting language such as R [^r], that pertains to four aspects: A) operating system, B) system components such as `libxml2`, C) the exact R version, and D) what and which version of the installed R packages.
+For a usual scripting language such as R [@rcore] [^r], that pertains to four aspects: A) operating system, B) system components such as `libxml2`, C) the exact R version, and D) what and which version of the installed R packages.
 We will call them Component A, B, C, D in the following sections.
@@
 In reality, the impact of Component A is relatively weak as mainstream, open source programming languages and their software libraries are usually cross platform.
-In modern computational research, Linux is the de-facto operating system in high performance computing environments (e.g.
-Slurm).
+In modern computational research, Linux is the de-facto operating system in High Performance Computing environments.
 Instead, the impact of Components B, C, and D is much higher.
@@
 To build a container, one needs to write a plain text declarative description of the required computational environment.
-Inside this declarative description, it should pin down all four Components mentioned above.
+Inside this declarative description, it should unambiguously specify all four Components mentioned above.
 For Docker, it is in the form of a plain text file called `Dockerfile`.
@@
 The case in point described in this paper, `rang`, is one of them.
-We argue that `rang` is the only easy-to-use solution available that can pin down and restore all four components without the reliance on any commercial service <!-- such as MRAN -->.
+We argue that `rang` is the only easy-to-use solution available that can specify and restore all four components without the reliance on any commercial service <!-- such as MRAN -->.
 
@@
 
-`renv` [@renvrpkg] (and its derivatives such as `jetpack` and its predecessor `packrat`) takes a similar approach to Python's `virtualenv` and Ruby's `Gem` to pin down the exact version of R packages using a "lock file".
-Other solutions such as `checkpoint` [@checkpointrpkg] depend on the availability of The Microsoft R Application Network (MRAN, a time-stamped daily backup of CRAN), which will be shut down on July 1st, 2023.
+`renv` [@renvrpkg] (and its derivatives such as `jetpack` and its predecessor `packrat`) takes a similar approach to Python's `virtualenv` and Ruby's `Gem` to unambiguously specify the exact version of R packages using a "lock file".
+Other solutions such as `checkpoint` [@checkpointrpkg] depend on the availability of The Microsoft R Application Network (MRAN, a time-stamped daily backup of CRAN), which will be shut down on July 1, 2023.
 `groundhog` [@groundhogrpkg] used to depend on MRAN but has a plan to switch to their home-grown R package repository.
-These solution can effectively pin down Component C and D.
+These solution can effectively specify Component C and D.
 But they can only restore component D.
@@
 In this way, it fixes Component A.
-But `containerit` does not pin down the exact version of R packages.
-Therefore, it can pin down components A, B, C, but only a part of component D.
-`dockta` [^dockta] is another containerization solution that can potentially pin down all components due to the fact that MRAN is used.
-But it also suffers from the same limitations mentioned above.
+But `containerit` does not specify the exact version of R packages.
+Therefore, it can specify components A, B, C, but only a part of component D.
+`dockta` [^dockta] is another containerization solution that can potentially specify all components due to the fact that MRAN is used.
+But it also suffers from the same limitations of MRAN mentioned above, i.e.
+shutting down on July 1, 2023, going back as far as September 17, 2014, and covering only CRAN.
 
@@
 
-There are two important functions of `rang`: `resolve()` and `dockerize()`.
+There are two important functions of `rang` version 0.2.0: `resolve()` and `dockerize()`.
 
@@
 `compute.es::tes()`.
-This undocumented package can be detected by `renv::dependencies()`, which is the provider of the scanning function of `rang`.
+This undocumented package can be detected by `renv::dependencies()`, of which `as_pkgrefs()` is a wrapper.
 
@@
 @loercher:2017:D].
-Despite the covert editing of the package by a staffer of CRAN [^evidence], the package was removed from CRAN in 2019 [^checkerror].
-We attempted to install the second last (the original submitted version) and last (with covert editing) versions of `maxent` on R 4.2.2.
-Both of them didn't work.
+Despite the editing of the package by a staffer of CRAN [^evidence], the package was removed from CRAN in 2019 [^checkerror].
+We attempted to install the second last (the original submitted version) and last (with editing) versions of `maxent` on R 4.2.2.
+Neither worked.
 
@@
 ```r
-require(ptproc)
+library(ptproc)
 
@@
 ```bash
-Rscript -e "require(rang); dockerize(resolve('ptproc', '2004-07-01'),
+Rscript -e "library(rang); dockerize(resolve('ptproc', '2004-07-01'),
 'pengdocker', cache = TRUE)"
@@
 ```sh
-Rscript -e "require(rang); dockerize(resolve('Sushi', '2014-06-05'),
+Rscript -e "library(rang); dockerize(resolve('Sushi', '2014-06-05'),
 'sushidocker', no_rocker = TRUE, cache = TRUE)"
@@
 [@oser:2022:HPE] and prepared a research compendium that should have long-term computational reproducibility.
-The research compendium is similar to the Executable Compendium suggested by the Turing way.
+The research compendium is similar to the Executable Compendium suggested by the Turing way [@community:2019:TW].
 
@@
 ```r
-require(rang)
+library(rang)
 ## meta-analysis is the directory of all shared materials
@@
 That can be regenerated by running `make resolve`.
-However, having this directory preserved insures against the situations that some R packages used in the project were no longer available or any of the information providers used by `rang` for resolving the dependency relationships were not available.
-(Or in the rare circumstance of `rang` is no longer available.)
+However, preserving this directory insures against the situations where (1) some R packages used in the project are no longer available, (2) any of the information providers used by `rang` for resolving the dependency relationships are not available, or (3) in the rare circumstance of `rang` is no longer available.
 
@@
 
-There are still many features that we did not mention in this paper.
+There are still many features that we did not mention in this paper (see Table 1).
 `rang` is built with interoperability in mind.
@@
 Computational reproducibility is a complex topic and as in all of these complex topics, there is no silver bullet [@canon:2019:CPR].
-All solutions have their trade-offs.
-The (re)construction process based on `rang` takes notably more time than other solutions because all packages are compiled from source.
+All solutions have their trade-offs and cannot solve all issues.
+Specifically, `rang` can restore a specific computational environment but there might be potential barriers to reproducibility related to external dependencies that cannot be containerized.
+A common external dependency is an external data source connected via a web application programming interface (API).
+Some examples are the so-called "API packages" such as `STRINGdb` (an R interface to the STRING protein-protein database), `rtoot` (an R interface to the Mastodon API), and `googleLanguageR` (an R interface to the Google Translation API).
+Although these API packages can be containerized, the external data sources cannot.
+Therefore, `rang` is not a solution to issues related to external dependencies.
+For data collection via these API packages, we recommend caching the collected data and sharing the cached data.
+For computation via API packages (e.g.
+Google Translate), we recommend using alternative methods that do not require external dependencies (e.g.
+[@chan:2020:REC]).
+
+Moreover, the (re)construction process based on `rang` takes notably more time than other solutions because all packages are compiled from source.
 `rang` trades computational efficiency of this often one-off (re)constructing process for correctness, backward compatibility and independence from any commercial backups of software repositories such as MRAN.
-There are also other limitations.
-In the Vignette of `rang` ([https://cran.r-project.org/web/packages/rang/vignettes/faq.html](https://cran.r-project.org/web/packages/rang/vignettes/faq.html)), we list all of these limitations as well as possible mitigation.
+In the Vignette of `rang` ([https://cran.r-project.org/web/packages/rang/vignettes/faq.html](https://cran.r-project.org/web/packages/rang/vignettes/faq.html)), we list all limitations: (1) `rang` does not support R < 1.3.1, (2) `rang` can generate Ubuntu/Debian-based Docker images and therefore non-Linux features (e.g.
+WinBUGS) are not supported, (3) `dockerize()` cannot cache OS-level packages (`deb` packages), (4) R packages with changing system requirements might be incorrectly queried, (5) caching of R packages and R source must be enforced when the R version is too old, (6) R packages on CRAN, Bioconductor, and Github, as well as Rocker images, might not be available in the future.
+All caveats listed have no solution, except point (6) can be mitigated by backing up the generated Docker image (see the section on research compendia).
+
+| Function           | Purpose                                                                              |
+|--------------------|--------------------------------------------------------------------------------------|
+| `as_pkgrefs`       | Convert data structures (e.g. `sessionInfo`) into package references for `resolve()` |
+| `convert_edgelist` | Convert the resolved results to edgelist                                             |
+| `dockerize`        | Dockerize the resolved result                                                        |
+| `export_rang`      | Export the resolved result as installation script                                    |
+| `export_renv`      | Export the resolved result as renv lockfile                                          |
+| `query_sysreqs`    | Query for system requirements                                                        |
+| `resolve`          | Resolve dependencies of R packages                                                   |
+
+: List of all functions in rang version 0.2.0
 

### 2023-05-18 paper/paper.qmd

@@
 
-[^r]: In this paper, we will focus on R, a popular programming language used frequently in various computational fields (e.g. computational social science, bioinformatics).
+[^r]: In this paper, we will focus on R, a popular programming language used frequently in various computational fields (e.g. computational social science or bioinformatics).
 
-In reality, the impact of Component A is relatively weak as mainstream, open source programming languages and their software libraries are usually cross platform.
+In reality, the impact of Component A is relatively weak as mainstream open source programming languages and their software libraries are usually cross platform.
 In modern computational research, Linux is the de-facto operating system in High Performance Computing environments.
@@
 This might be due to a lack of (DevOps) skills of most scientists to create a `Dockerfile` [@kim:2018:E].
-But there are many tools available to automate the process [e.g.
+But there are many tools available to automate this process [e.g.
 @nuest:2019].
@@
 Posit also provides a free (*gratis*) time-stamped daily backup of CRAN and Bioconductor (a series of repositories of R package for bioinformatics and computational biology) called Posit Public Package Manager [^posit].
-It can goes as far back as October 10, 2017.
+It can go as far back as October 10, 2017.
 
@@
 using them now to ensure the reproducibility of the current research for future researchers.
-`rang` mostly targets retrospective usage, i.e.
-using `rang` to reconstruct historical R computational environments for which the declarative descriptions are not available.
+`rang` mostly targets retrospective usage, i.e., to reconstruct historical R computational environments for which the declarative descriptions are not available.
 One can think of `rang` as an archaeological tool.
@@
 As of writing, `rang` does not support R < 1.3.1, i.e.
-snapshot date earlier than 2001-08-31 (which is 13 years earlier than all solutions depending on MRAN).
+snapshot dates earlier than 2001-08-31 (which is 13 years earlier than all solutions depending on MRAN).
 There are two features of `dockerize()` that are important for future reproducibility.
@@
 However, it is also possible to cache (or archive) the source packages now.
-The archived R packages can then be used instead during the building process.
+The archived R packages can then be used during the building process instead.
 The significance of this step in terms of long-term computational reproducibility will be discussed in Section 4.
@@
 When `as_pkgrefs()` is applied to a single path of a directory, it scans all relevant files (`DESCRIPTION`, R scripts and R Markdown files) for all R packages used (based on `renv::dependencies()` <!-- [^gh] -->).
-How it works is demonstrated in three of the following examples below.
-But an important caveat is that it can only scan CRAN and Bioconductor packages.
+Its functionality is demonstrated in three of the following examples below.
+However, an important caveat is that it can only scan CRAN and Bioconductor packages.
 
@@
 
-The following are some examples of how `rang` can be used to make shared, but otherwise unexecutable, R code runnable again.
+In this section, we present several examples for how `rang` can be used to make shared, but otherwise unexecutable, R code runnable again.
 The examples were drawn from various fields spanning from political science, psychological science, and bioinformatics.
@@
 
-The issue is that installing `quanteda` 1.3.4 this way installs the latest dependencies from CRAN.
+The problem is that installing `quanteda` 1.3.4 this way installs the latest dependencies from CRAN and not the "period appropriate" ones.
 `quanteda` 1.3.4 uses a deprecated (but not yet removed) function of `Matrix` (`as(<dgTMatrix>, "dgCMatrix")`).
-If this function were removed in the future, the above code snippet would not be executable anymore.
+If this function is removed in the future, the above code snippet will not be executable anymore.
 
@@
 
-Inside the container, the materials are located in the `materials` directory. We used the following shell script to test the reproducibility of all R scripts.
+Inside the container, the material is located in the `materials` directory. We used the following shell script to test the reproducibility of all R scripts.
 
@@
 For removed CRAN packages, we strongly recommend querying the Github read-only mirror of CRAN instead (https://github.com/cran).
-It is because in this way, the resolved system requirements have a higher chance of being correct.
+This way, the resolved system requirements have a higher chance of being correct.
 
@@
 The package is still listed in the "Handling and Analyzing Spatio-Temporal Data" CRAN Task View [^taskview] despite being uninstallable without modification on any modern R system (see below).
-As of writing, the package, as a tarball file (tar.gz), is still downloadable from the original author's website [^peng].
+As of writing the package is still downloadable from the original author's website[^peng] as a tarball file (tar.gz).
 
@@
 As late as 2017, there are blog posts on how to install the long obsolete package on modern versions of R [^blog].
-The package is extremely challenging to install on a modern R system because the package was written before the introduction of name space management in R 1.7.0 [@RN-2003-001].
+The package is extremely challenging to install on a modern R system because it was written before the introduction of name space management in R 1.7.0 [@RN-2003-001].
 In other words, the available tarball files from the original author's website and CRAN do not contain a `NAMESPACE` file like all other modern R packages do.
@@
 
-The preparation of the research compendium is easy as `rang` can scan a materials directory for all R packages used [^dmetar].
+The preparation of the research compendium is easy since `rang` can scan a materials directory for all R packages used [^dmetar].
 
@@
 These retrospective use cases demonstrate how versatile `rang` is.
-`rang` is also helpful for prospective usage, as demonstrated in Section 4 whereby an executable compendium is created.
+`rang` is also helpful for prospective usage, as demonstrated in Section 4 where an executable compendium is introduced.
 
-There are still many features that we did not mention in this paper (see Table 1).
+There are still many features that we did not cover in this paper (see Table 1).
 `rang` is built with interoperability in mind.
@@
 All solutions have their trade-offs and cannot solve all issues.
-Specifically, `rang` can restore a specific computational environment but there might be potential barriers to reproducibility related to external dependencies that cannot be containerized.
-A common external dependency is an external data source connected via a web application programming interface (API).
-Some examples are the so-called "API packages" such as `STRINGdb` (an R interface to the STRING protein-protein database), `rtoot` (an R interface to the Mastodon API), and `googleLanguageR` (an R interface to the Google Translation API).
+Although `rang` can restore a specific computational environment,  there might be potential barriers to reproducibility related to external dependencies which cannot be containerized.
+A common external dependency is a data source connected via a web application programming interface (API).
+Some examples are so-called "API packages" such as `STRINGdb` (an R interface to the STRING protein-protein database), `rtoot` (an R interface to the Mastodon API), and `googleLanguageR` (an R interface to the Google Translation API).
 Although these API packages can be containerized, the external data sources cannot.

### 2023-05-18 paper/paper.qmd
Minor update on the paper

@@
 The archived R packages can then be used during the building process instead.
-The significance of this step in terms of long-term computational reproducibility will be discussed in Section 4.
+The significance of this step in terms of long-term computational reproducibility will be discussed in the section on executable compendia.
 
@@
 
-# Preparing research compendia with long-term computational reproducibility
+# Preparing executable compendia with long-term computational reproducibility
 
 The above six examples show how powerful `rang` is to reconstruct tricky computational environments which have not been completely declared in the literature.
-Although we position `rang` mostly as an archaeological tool, we think that `rang` can also be used to prepare research compendia of current research.
-We can't predict the future but research compendia generated by `rang` would probably have long-term computational reproducibility.
+Although we position `rang` mostly as an archaeological tool, we think that `rang` can also be used to prepare executable research compendia of current research.
+We cannot predict the future but research compendia generated by `rang` would probably have long-term computational reproducibility.
 
@@
 These retrospective use cases demonstrate how versatile `rang` is.
-`rang` is also helpful for prospective usage, as demonstrated in Section 4 where an executable compendium is introduced.
+`rang` is also helpful for prospective usage, as demonstrated in the section where an executable compendium is introduced.
 
@@
 WinBUGS) are not supported, (3) `dockerize()` cannot cache OS-level packages (`deb` packages), (4) R packages with changing system requirements might be incorrectly queried, (5) caching of R packages and R source must be enforced when the R version is too old, (6) R packages on CRAN, Bioconductor, and Github, as well as Rocker images, might not be available in the future.
-All caveats listed have no solution, except point (6) can be mitigated by backing up the generated Docker image (see the section on research compendia).
+All caveats listed have no solution, except point (6) can be mitigated by backing up the generated Docker image (see the section on executable compendia).
 
 | Function           | Purpose                                                              |
-|--------------------|--------------------------------------------------------------------------------------|
-| `as_pkgrefs`       | Convert data structures (e.g. `sessionInfo`) into package references for `resolve()` |
+|--------------------|----------------------------------------------------------------------|
+| `as_pkgrefs`       | Convert data structures (e.g. `sessionInfo`) into package references |
 | `convert_edgelist` | Convert the resolved results to edgelist                             |

### 2023-05-27 paper/paper.qmd
Remove all footnotes

@@
 But as computer science research has shown, this task is incredibly complex [@abate:2015:MCR;@dolstra:2010:N].
-For a usual scripting language such as R [@rcore] [^r], that pertains to four aspects: A) operating system, B) system components such as `libxml2`, C) the exact R version, and D) what and which version of the installed R packages.
+For a usual scripting language such as R [@rcore] (also a popular programming language used frequently in various computational fields, e.g.
+computational social science or bioinformatics) as an example, that pertains to four aspects: A) operating system, B) system components such as `libxml2`, C) the exact R version, and D) what and which version of the installed R packages.
 We will call them Component A, B, C, D in the following sections.
@@
 If the shared computer code requires the R package `xml2`, then the whole execution fails.
-
-[^r]: In this paper, we will focus on R, a popular programming language used frequently in various computational fields (e.g. computational social science or bioinformatics).
 
@@
 Therefore, it can specify components A, B, C, but only a part of component D.
-`dockta` [^dockta] is another containerization solution that can potentially specify all components due to the fact that MRAN is used.
+`dockta` <!-- [^dockta] --> is another containerization solution that can potentially specify all components due to the fact that MRAN is used.
 But it also suffers from the same limitations of MRAN mentioned above, i.e.
@@
 It is also worth mentioning that MRAN is not the only archival service.
-Posit also provides a free (*gratis*) time-stamped daily backup of CRAN and Bioconductor (a series of repositories of R package for bioinformatics and computational biology) called Posit Public Package Manager [^posit].
+Posit also provides a free (*gratis*) time-stamped daily backup of CRAN and Bioconductor (a series of repositories of R package for bioinformatics and computational biology) called Posit Public Package Manager <!-- [^posit] -->.
 It can go as far back as October 10, 2017.
 
-[^posit]: [https://packagemanager.rstudio.com/client/#/repos/2/packages/](https://packagemanager.rstudio.com/client/#/repos/2/packages/)
+<!-- [^posit]: [https://packagemanager.rstudio.com/client/#/repos/2/packages/](https://packagemanager.rstudio.com/client/#/repos/2/packages/) -->
 
-[^dockta]: [https://stencila.github.io/dockta/](https://stencila.github.io/dockta/)
+<!-- [^dockta]: [https://stencila.github.io/dockta/](https://stencila.github.io/dockta/) -->
 
@@
 By default, the Docker image building process downloads source packages from their sources and then compiles them.
-This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) [^1], Bioconductor, and Github.
+This step depends on the future availability of R packages on CRAN (which is extremely likely to be the case in the near future, given the continuous availability since 1997-04-23) <!-- [^1] -->, Bioconductor, and Github.
 However, it is also possible to cache (or archive) the source packages now.
@@
 
-[^1]: [https://stat.ethz.ch/pipermail/r-announce/1997/000001.html](https://stat.ethz.ch/pipermail/r-announce/1997/000001.html)
+<!-- [^1]: [https://stat.ethz.ch/pipermail/r-announce/1997/000001.html](https://stat.ethz.ch/pipermail/r-announce/1997/000001.html) -->
 
@@
 The first argument of `resolve()` is processed by a separate function called `as_pkgrefs()`.
-For interoperability, `rang` supports the "package references" standard [^pkgrefs] used also in other packages such as `renv` [@renvrpkg].
+For interoperability, `rang` supports the "package references" standard <!-- [^pkgrefs] --> used also in other packages such as `renv` [@renvrpkg].
 It is mostly used for converting "shorthands" (e.g.
@@
 
-[^pkgrefs]: [https://r-lib.github.io/pkgdepends/reference/pkg_refs.html](https://r-lib.github.io/pkgdepends/reference/pkg_refs.html)
+<!-- [^pkgrefs]: [https://r-lib.github.io/pkgdepends/reference/pkg_refs.html](https://r-lib.github.io/pkgdepends/reference/pkg_refs.html) -->
 
@@
 
-All data and computer code are available from GitHub with the last commit on 2019-01-17 [^hilgard].
+All data and computer code are available from GitHub with the last commit on 2019-01-17<!-- [^hilgard] -->.
 The R code contains a list of R packages used in the project as `library()` statements, including an R package on GitHub that is written by the main author of that paper.
@@
 
-[^hilgard]: [https://github.com/Joe-Hilgard/vvg-2d4d](https://github.com/Joe-Hilgard/vvg-2d4d)
+<!-- [^hilgard]: [https://github.com/Joe-Hilgard/vvg-2d4d](https://github.com/Joe-Hilgard/vvg-2d4d) -->
 
@@
 
-This R script has been "rescued" by the author of the R package `groundhog` [@groundhogrpkg], as demonstrated in a blog post [^groundhog].
+This R script has been "rescued" by the author of the R package `groundhog` [@groundhogrpkg], as demonstrated in a blog post <!-- [^groundhog] -->(http://datacolada.org/100).
 We were wondering if `rang` can also be used to "rescue" the concerned R script.
@@
 
-[^groundhog]: [http://datacolada.org/100](http://datacolada.org/100)
+<!-- [^groundhog]: [http://datacolada.org/100](http://datacolada.org/100) -->
 
@@
 @loercher:2017:D].
-Despite the editing of the package by a staffer of CRAN [^evidence], the package was removed from CRAN in 2019 [^checkerror].
-We attempted to install the second last (the original submitted version) and last (with editing) versions of `maxent` on R 4.2.2.
+Despite the editing of the package by a staffer of CRAN [^evidence], the package was removed from CRAN in 2019.
+(Here, we said the package `maxent` was "removed from CRAN" as per what the webpage ([https://cran.r-project.org/web/packages/maxent/index.html](https://cran.r-project.org/web/packages/maxent/index.html)) is written.
+However, the terminology as stated in the [CRAN Policies](https://cran.r-project.org/web/packages/policies.html#Source-packages) is "archived", quote: "Packages will not normally be removed from CRAN: however, they may be archived, including at the maintainer’s request." As all CRAN packages have all the submitted versions archived, we find this terminology confusing.
+Therefore, we use "removed" packages throughout this paper to indicate packages that cannot be installed by the usual method, i.e.
+`install.packages` but still have the old versions archived on CRAN.) We attempted to install the second last (the original submitted version) and last (with editing) versions of `maxent` on R 4.2.2.
 Neither worked.
 
-Using `rang`, we are able to reconstruct a computational environment with R 2.15.0 (2012-03-30) to run all code snippets published in Jurka [@jurka:2012] [^speed].
+Using `rang`, we are able to reconstruct a computational environment with R 2.15.0 (2012-03-30) to run all code snippets published in Jurka [@jurka:2012] <!-- [^speed] -->.
 For removed CRAN packages, we strongly recommend querying the Github read-only mirror of CRAN instead (https://github.com/cran).
@@
 
-[^checkerror]: This is the check log: [https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html](https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html) Here, we said the package `maxent` was "removed from CRAN" as per what the webpage ([https://cran.r-project.org/web/packages/maxent/index.html](https://cran.r-project.org/web/packages/maxent/index.html)) is written.
-However, the terminology as stated in the [CRAN Policies](https://cran.r-project.org/web/packages/policies.html#Source-packages) is "archived", quote: "Packages will not normally be removed from CRAN: however, they may be archived, including at the maintainer’s request." As all CRAN packages have all the submitted versions archived, we find this terminology confusing.
-Therefore, we use "removed" packages throughout this paper to indicate packages that cannot be installed by the usual method, i.e.
-`install.packages` but still have the old versions archived on CRAN.
+<!-- [^checkerror]: [https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html](https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html)  -->
 
@@
 
-[^speed]: On an interesting historical side note: The original paper reported ---based on a benchmark--- that "the algorithm is very fast; `maxent` uses only 135.4 megabytes of RAM and finishes in 53.3 seconds." On a modest computer in 2023 with a dockerized R 2.15.0, the benchmark finishes in 4 seconds.
+<!-- [^speed]: On an interesting historical side note: The original paper reported ---based on a benchmark--- that "the algorithm is very fast; `maxent` uses only 135.4 megabytes of RAM and finishes in 53.3 seconds." On a modest computer in 2023 with a dockerized R 2.15.0, the benchmark finishes in 4 seconds.
+-->
 
@@
 The only release on CRAN was on 2002-10-10.
-The package is still listed in the "Handling and Analyzing Spatio-Temporal Data" CRAN Task View [^taskview] despite being uninstallable without modification on any modern R system (see below).
-As of writing the package is still downloadable from the original author's website[^peng] as a tarball file (tar.gz).
+The package is still listed in the "Handling and Analyzing Spatio-Temporal Data" CRAN Task View <!-- [^taskview] --> despite being uninstallable without modification on any modern R system (see below).
+As of writing the package is still downloadable from the original author's website<!-- [^peng] --> as a tarball file (tar.gz).
 
-[^taskview]: [https://cran.r-project.org/web/views/SpatioTemporal.html](https://cran.r-project.org/web/views/SpatioTemporal.html)
+<!-- [^taskview]: [https://cran.r-project.org/web/views/SpatioTemporal.html](https://cran.r-project.org/web/views/SpatioTemporal.html) -->
 
 Even with this over-a-decade removal and new packages with similar functionalities have been created, there is evidence that `ptproc` is still being sought for.
-As late as 2017, there are blog posts on how to install the long obsolete package on modern versions of R [^blog].
+As late as 2017, there are blog posts on how to install the long obsolete package on modern versions of R <!-- [^blog] -->.
 The package is extremely challenging to install on a modern R system because it was written before the introduction of name space management in R 1.7.0 [@RN-2003-001].
@@
 The file `peng.Rout` contains the execution results of the script from inside the Docker container.
-As the random seed was preserved by the original author [@peng:2003:MDP], the above BASH script can perfectly reproduce the analysis [^random].
+As the random seed was preserved by the original author [@peng:2003:MDP], the above BASH script can perfectly reproduce the analysis.
+It is also important to note that the random number generator (RNG) of R has been changed several times over the course of the development.
+In this case, we are using the same generation of RNG as Peng [@peng:2003:MDP] in order to reproduce the analysis.
 
-[^peng]: [https://www.biostat.jhsph.edu/~rpeng/software/](https://www.biostat.jhsph.edu/~rpeng/software/)
+<!-- [^peng]: [https://www.biostat.jhsph.edu/~rpeng/software/](https://www.biostat.jhsph.edu/~rpeng/software/) -->
 
-[^blog]: [https://blog.mathandpencil.com/installing-ptproc-on-osx](https://blog.mathandpencil.com/installing-ptproc-on-osx) and [https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/](https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/)
-
-[^random]: It is also important to note that the random number generator (RNG) of R has been changed several times over the course of the development.
-In this case, we are using the same generation of RNG as Peng [@peng:2003:MDP].
+<!-- [^blog]: [https://blog.mathandpencil.com/installing-ptproc-on-osx](https://blog.mathandpencil.com/installing-ptproc-on-osx) and [https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/](https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/) -->
 
@@
 
-The preparation of the research compendium is easy since `rang` can scan a materials directory for all R packages used [^dmetar].
+The preparation of the research compendium is easy since `rang` can scan a materials directory for all R packages used.
+We detected a minor issue in the code base that an undeclared Github package is used.
+But it can be easily solved, as in the Psychological Science example above.
 
@@
 When `oserimg.tar.gz` is available, it is possible to run `make rebuild` and `make render` even without internet access (provided that Docker and `make` have been installed before).
-Of course, there is still an extremely rare situation where Docker (the program) itself is no longer available [^make].
-However, it is possible to convert the image file for use on other containerization solutions such as Singularity [^singularity], if Docker is really not available anymore.
+Of course, there is still an extremely rare situation where Docker (the program) itself is no longer available <!-- [^make] -->.
+However, it is possible to convert the image file for use on other containerization solutions such as Singularity <!-- [^singularity] -->, if Docker is really not available anymore.
 
@@
 Zenodo, for example, allows the sharing of 50G of files.
-Therefore, sharing of the last two components of the executable compendium prepared with `rang` is at least possible on Zenodo [^zenodo].
+Therefore, sharing of the last two components of the executable compendium prepared with `rang` is at least possible on Zenodo (https://doi.org/10.5281/zenodo.7708417).
 However, for data repositories with more restrictions on data size, sharing the executable compendium without the last two parts could be considered sufficient.
@@
 
-[^zenodo]: The complete version of the executable compendium is available on Zenodo: [https://doi.org/10.5281/zenodo.7708417](https://doi.org/10.5281/zenodo.7708417)
+<!-- [^zenodo]: [https://doi.org/10.5281/zenodo.7708417](https://doi.org/10.5281/zenodo.7708417) -->
 
@@
 
-[^make]: We can't imagine a world without `Make`, a tool that has been available since 1976.
+<!-- [^make]: We can't imagine a world without `Make`, a tool that has been available since 1976. -->
 
-[^singularity]: [https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html](https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html)
-
-[^dmetar]: We detected a minor issue in the code base that an undeclared Github package is used. But it can be easily solved, as in the Psychological Science example above.
+<!-- [^singularity]: [https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html](https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html) -->
 

### 2023-06-03 paper/paper.qmd
Fix flattening

@@
 @loercher:2017:D].
-Despite the editing of the package by a staffer of CRAN [^evidence], the package was removed from CRAN in 2019.
+Despite the editing of the package by a staffer of CRAN <!-- [^evidence] -->, the package was removed from CRAN in 2019.
 (Here, we said the package `maxent` was "removed from CRAN" as per what the webpage ([https://cran.r-project.org/web/packages/maxent/index.html](https://cran.r-project.org/web/packages/maxent/index.html)) is written.
@@
 
-[^evidence]: [https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9](https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9)
+<!-- [^evidence]: [https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9](https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9) -->
 
@@
 `rang` has native support for Bioconductor packages since version 0.2.
-We obtained the R script `"PaperFigure.R"` from the Github repository of `Sushi` [^sushi], which generates the figure in the original paper [@phanstiel:2014:S].
+We obtained the R script `"PaperFigure.R"` from the Github repository of `Sushi` <!-- [^sushi] -->, which generates the figure in the original paper [@phanstiel:2014:S].
 Similar to the above case of `ptproc`, we made a completely automated BASH script to run `"PaperFigure.R"` and get the generated figure out of the container (`Fig~\ref{fig2}`{=latex}).
@@
 
-[^sushi]: [https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R](https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R)
+<!-- [^sushi]: [https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R](https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R) -->
 

### 2023-06-09 paper/before-body.tex

$-- PLOS insert title and author within the main body
$-- Use this template to insert the necessary pieces before any content provided in .qmd file
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{$title$} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
$by-author:_authors.tex()[, ]$
\\
\bigskip
$by-affiliation:_affiliations.tex()[\\ ]$
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
$for(by-author)$
$if(by-author.attributes.corresponding)$
* $by-author.email$
$endif$
$endfor$

\end{flushleft}

$if(abstract)$
\section*{Abstract}
$abstract$
$endif$

$if(author-summary)$
\section*{Author summary}
$author-summary$
$endif$

\linenumbers

### 2023-07-17 paper/basel2023_rang/index.qmd
Add basel slides

---
title: "rang"
subtitle: "(Re)constructing reproducible R computational environment"
author: Chung-hong Chan, David Schoch
institute: GESIS
date: "2023-07-21"
date-format: "YYYY-MM-DD"
footer: "https://chainsawriot.github.io/basel2023_rang"
embed-resources: true
format: fakegesis-revealjs
---

## Computational reproducibility

|                        | **Same Data** | **Different Data** |
|------------------------|---------------|--------------------|
| **Same Analysis**      | Reproducible  | Replicable         |
| **Different Analysis** | Robust        | Generalisable      |

::: {style="font-size: 0.3em"}

Source: [The Turing Way](https://the-turing-way.netlify.app/reproducible-research/overview/overview-definitions.html)

:::

## Computational Reproducibility

A result is reproducible when the same analysis steps performed on the same dataset **consistently produce the same answer**. This reproducibility of the result can be checked by the original investigators and other researchers within a local environment.

::: {style="font-size: 0.3em"}
Schoch, Chan, Wagner, & Bleier. (2023) Computational Reproducibility in Computational Social Science. [arXiv](https://doi.org/10.48550/arXiv.2307.01918)
:::

## Case study I: dplyr

```{r}
#| echo: true
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
data
```

## Case study I: dplyr

Are these analysis steps performed on the same dataset **consistently produce the same answer**?

```r
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

## On my own machine: dplyr 1.1.2

```{r}
#| echo: true
#| error: true
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

## Also on my own machine

![](dplyr.png)

## Time machine

:::: {.columns}

:::{.column width="50%"}

![](bowie.jpg)

(1947.01.08 - 2016.01.10)
:::

:::{.column width="50%"}

![](dplyr1.png)

:::

::::

::: {.notes}
Before Tidyverse (2016-09-09). Trump and Brexit too.
:::

## Method 1

Code rewrite

```{.r code-line-numbers="|3"}
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id, .keep_all = TRUE) %>%
    select(outcome)
```

::: {.notes}
You'll need to edit your code when there are new breaking changes
:::

## Method 2

Install the old version

```{.r code-line-numbers="|1"}
remotes::install_version("dplyr", "0.4.3")
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

::: {.notes}
It installs the latest recursive dependencies
:::

## (Non-)method 3

Using ~~MRAN~~ Posit Public Package Manager (P3M)

```{.r code-line-numbers="|1|2"}
p3m <- "https://packagemanager.posit.co/cran/2016-01-10"
options(repos = c(REPO_NAME = p3m))
install.packages("dplyr") # install 0.4.3?

require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

::: {.notes}
It won't work because P3M can only go as far back as 2017 Oct 10.
It installs the "time-appropriate" recursive dependencies. But it mixes old packages with your current system library.
Of course, you can use install.packages("dplyr", lib = "xyz")
:::

## (Non-)method 4

Using `renv` 1.0.0

```{.r code-line-numbers="|3"}
require(renv)
renv::init() ## restart
renv::checkout(date = "2016-01-10") 
install.packages("dplyr")
renv::snapshot()
require(dplyr)
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
distinct(data, id) %>%
    select(outcome)
```

::: {.notes}
It won't work also because P3M can only go as far back as 2017 Oct 10.
:::

## Pit shop 1: Summary

1. Computational reproducibility is tacitly dependent on the computational environment
2. Recursive dependencies can best be tackled with time-based pinning or lockfile-based system
3. P3M has its limit (2017-10-10).

::: {.notes}
P3M also has an inconvenience limitation: What if it's no longer available (like MRAN before it)?
:::

## Case study II

![](comps.png)

::: {style="font-size: 0.3em"}
Source: @whydoesr@botsin.space
:::

## Case study II

Are these analysis steps performed on the same dataset **consistently produce the same answer**?

```r
unix.time(mean(rnorm(100)))
n _ 10
require(sna)
x _ rgraph(n, mode = "graph")
evcent(x)
```

## On my own machine: R 4.3.1

```{r}
#| echo: true
#| error: true
unix.time(mean(rnorm(100)))
n _ 10
suppressPackageStartupMessages(require(sna))
x _ rgraph(n, mode = "graph")
evcent(x)
```

Explanations:

```r
?`base-defunct`
```

## Non-method #1: P3M/renv

. . .

It's base.

::: {.notes}
in Norm Matloff's voice
:::

## Method #2?: Docker / Rocker

```bash
docker run --rm -it rocker/r-ver:3.2.0
```

. . .

```text
> unix.time(mean(rnorm(100)))
   user  system elapsed 
      0       0       0 
> n _ 10
Error: unexpected input in "n _"
> 
```

::: {.notes}
You probably will have problem installing sna
:::

## Pit shop 2: Summary

1. Computational reproducibility is tacitly dependent on the computational environment
2. It's not always about packages, the programming language changes too
3. Rocker has its limit (R 3.0.0)

::: {.notes}
(think Python 2 -> 3)

Rocker: Many old images are not usable anymore (either hardcoded MRAN or expiring Debian)
:::

## Our solution: rang

![](rang_logo.png){width=300,fig-align="left"}

```r
install.packages("rang")
## remotes::install_github("chainsawriot/rang")
```

::: {style="font-size: 0.3em"}
Chan & Schoch (2023). rang: Reconstructing reproducible R computational environments. PLOS One [doi](https://doi.org/10.1371/journal.pone.0286761)
:::

## Features (CRAN version)

* Date-based pinning (like P3M, `renv`)
* Create container images (like Rocker)
* Support CRAN, Bioconductor, GitHub, & local packages
* Resolve system dependencies, e.g. `libxml2`
* Depend on r-hub, CRAN, Bioconductor, GitHub API; doesn't use P3M
* Interoperability with `renv` (lockfile, project scanning)
* Can go as far back as **2001-08-31** (or R 1.3.1)

::: {.notes}
We are talking about Tony Blair and Bush 43 (and that's two weeks before 911)
:::

## Usage: Resolve

Resolve the dependency graph

```r
library(rang)
graph <- resolve(pkgs = c("dplyr"), snapshot_date = "2016-01-10")
graph
```

```{r}
#| echo: false
library(rang)
readRDS("dplyr.RDS")
```

## Usage: Dockerize

```r
dockerize(rang = graph,
          output_dir = "~/dplyr_example",
          no_rocker = TRUE,  # recommended
          cache = TRUE) # recommended
```

## What's inside

```text
├── cache
│   ├── debian
│   │   └── rootfs.tar.xz
│   ├── rpkgs
│   │   ├── assertthat_0.1.tar.gz
│   │   ├── BH_1.60.0-1.tar.gz
│   │   ├── DBI_0.3.1.tar.gz
│   │   ├── dplyr_0.4.3.tar.gz
│   │   ├── lazyeval_0.1.10.tar.gz
│   │   ├── magrittr_1.5.tar.gz
│   │   ├── R6_2.1.1.tar.gz
│   │   └── Rcpp_0.12.2.tar.gz
│   └── rsrc
│       └── R-3.2.3.tar.gz
├── compile_r.sh
├── Dockerfile
├── rang.R
└── README

4 directories, 14 files
```

## README

```text
This README offers some guidance on how to work with the included docker container 
which was created with the R package rang on 2023-07-17

# Installing docker

The installation depends on the OS you are running
...
```

## Dockerfile

```Dockerfile
FROM scratch
ADD /cache/debian/rootfs.tar.xz /
ENV TZ UTC
RUN ln -snf /usr/share/zoneinfo/$TZ /etc/localtime \
	&& echo $TZ > /etc/timezone \
	&& apt-get update -qq \
	&& apt-get install wget locales build-essential r-base-dev  -y
ENV RANG_PATH /rang.R
ENV COMPILE_PATH /compile_r.sh
ENV CACHE_PATH /cache
COPY rang.R /rang.R
COPY compile_r.sh /compile_r.sh
COPY cache/rpkgs /cache/rpkgs
COPY cache/rsrc /cache/rsrc
RUN apt-get update -qq \
	&& apt-get install -y libpcre3-dev zlib1g-dev pkg-config libcurl4-openssl-dev
RUN bash $COMPILE_PATH 3.2.3
CMD ["R"]
```

## Build and launch

```sh
cd ~/dplyr_example
docker build -t dplyr .
docker run --rm --name "dplyr" -ti dplyr
```

. . .

![](dplyr.png)

## Full bash

```bash
Rscript -e "rang::resolve(\"sna\", snapshot_date = \"2002-02-01\") |> 
rang::dockerize(\"~/scoms\", cache = TRUE)"
cd ~/scoms
docker build -t scoms .
docker run --rm --name "scoms" -ti scoms
```
## For compatibility with S

![](base2.png)

## Many features not demo.

* More examples in [our paper](https://doi.org/10.1371/journal.pone.0286761)

Upcoming features

* Apptainer (née Singularity) - Thanks to the contribution from Egor Kotov
* Research Compendium
* Post installation steps, e.g. installing Quarto

## Summary

## Summary

:::: {.columns}

:::{.column width="80%"}
1. Computational reproducibility is tacitly dependent on the computational environment
2. Use P3M for projects after 2017-10-10
3. Use Rocker for projects newer than R 3.0.0
4. Use `rang` for probably all projects
:::

:::{.column width="20%"}
![](rang_logo.png){width=300}
:::

::::

::: {style="font-size: 0.3em"}
Schoch, Chan, Wagner, & Bleier. (2023) Computational Reproducibility in Computational Social Science. [arXiv](https://doi.org/10.48550/arXiv.2307.01918)

Chan & Schoch (2023). rang: Reconstructing reproducible R computational environments. PLOS One [doi](https://doi.org/10.1371/journal.pone.0286761)
:::

### 2023-07-20 paper/basel2023_rang/index.qmd
Correct typos
Say more in the last slide
Trim slides

@@
 
-## Computational reproducibility
+<!-- ## Computational reproducibility -->
 
-|                        | **Same Data** | **Different Data** |
-|------------------------|---------------|--------------------|
-| **Same Analysis**      | Reproducible  | Replicable         |
-| **Different Analysis** | Robust        | Generalisable      |
+<!-- |                        | **Same Data** | **Different Data** | -->
+<!-- |------------------------|---------------|--------------------| -->
+<!-- | **Same Analysis**      | Reproducible  | Replicable         | -->
+<!-- | **Different Analysis** | Robust        | Generalisable      | -->
 
-::: {style="font-size: 0.3em"}
+<!-- ::: {style="font-size: 0.3em"} -->
 
-Source: [The Turing Way](https://the-turing-way.netlify.app/reproducible-research/overview/overview-definitions.html)
+<!-- Source: [The Turing Way](https://the-turing-way.netlify.app/reproducible-research/overview/overview-definitions.html) -->
 
-:::
+<!-- ::: -->
 
@@
 
-:::: {.columns}
+<!-- :::: {.columns} -->
 
-:::{.column width="50%"}
+<!-- :::{.column width="50%"} -->
 
-![](bowie.jpg)
+<!-- ![](bowie.jpg) -->
 
-(1947.01.08 - 2016.01.10)
-:::
+<!-- (1947.01.08 - 2016.01.10) -->
+<!-- ::: -->
 
-:::{.column width="50%"}
+<!-- :::{.column width="50%"} -->
 
@@
 
-:::
+<!-- ::: -->
 
-::::
+<!-- :::: -->
 
@@
 
-```{.r code-line-numbers="|1|2"}
+```{.r code-line-numbers="|1|2|"}
 p3m <- "https://packagemanager.posit.co/cran/2016-01-10"
@@
 
-```{.r code-line-numbers="|3"}
+```{.r code-line-numbers="|3|"}
 require(renv)
@@
 
-## Pit shop 1: Summary
+## Pit stop 1: Summary
 
@@
 
-It's base.
+It's base. You can't pin `unix.time()` with P3M/`renv`.
 
@@
 ```bash
-docker run --rm -it rocker/r-ver:3.2.0
+# work
+docker run --rm -it rocker/r-ver:3.0.0
 ```
@@
 
+. . .
+
+```bash
+# not work
+docker run --rm -it rocker/r-ver:2.15.3
+```
+
 ::: {.notes}
@@
 
-## Pit shop 2: Summary
+## Pit stop 2: Summary
 
@@
 1. Computational reproducibility is tacitly dependent on the computational environment
-2. Use P3M for projects after 2017-10-10
-3. Use Rocker for projects newer than R 3.0.0
+2. Use P3M / renv for projects after 2017-10-10
+    * only at the level of R packages
+3. Plus Rocker for projects newer than R 3.0.0
 4. Use `rang` for probably all projects

### 2023-07-21 paper/basel2023_rang/index.qmd
Finalize the slides

@@
 ---
-
-<!-- ## Computational reproducibility -->
-
-<!-- |                        | **Same Data** | **Different Data** | -->
-<!-- |------------------------|---------------|--------------------| -->
-<!-- | **Same Analysis**      | Reproducible  | Replicable         | -->
-<!-- | **Different Analysis** | Robust        | Generalisable      | -->
-
-<!-- ::: {style="font-size: 0.3em"} -->
-
-<!-- Source: [The Turing Way](https://the-turing-way.netlify.app/reproducible-research/overview/overview-definitions.html) -->
-
-<!-- ::: -->
-
 ## Computational Reproducibility
@@
 
-<!-- :::: {.columns} -->
-
-<!-- :::{.column width="50%"} -->
-
-<!-- ![](bowie.jpg) -->
-
-<!-- (1947.01.08 - 2016.01.10) -->
-<!-- ::: -->
-
-<!-- :::{.column width="50%"} -->
-
 ![](dplyr1.png)
 
-<!-- ::: -->
-
-<!-- :::: -->
+::: {style="font-size: 0.3em"}
+Source: [https://dplyr.tidyverse.org/news/index.html#dplyr-050](https://dplyr.tidyverse.org/news/index.html#dplyr-050)
+:::
 
@@
 
-```{.r code-line-numbers="|1|2|"}
+```{.r code-line-numbers="|1,2"}
 p3m <- "https://packagemanager.posit.co/cran/2016-01-10"
@@
 
-```{.r code-line-numbers="|3|"}
+```{.r code-line-numbers="|4"}
 require(renv)
 renv::init() ## restart
+install.packages("dplyr")
 renv::checkout(date = "2016-01-10")
-install.packages("dplyr")
 renv::snapshot()

### 2023-07-30 R/resolve.R
Fix #150 fix #149 fix#152 (#151)
* Fix #150 fix #149

* Retry when hit by r-hub/pkgsearch#117

To make resolve more determinstic

@@ -456,17 +456,17 @@ print.rang <- function(x, all_pkgs = FALSE, ...) {
     return(res)
 }
 
-.gh <- function(path,ref = NULL,...){
+.gh <- function(path,ref = NULL,...) {
     url <- httr::parse_url("https://api.github.com/")
     url <- httr::modify_url(url, path = path)
     token <- Sys.getenv("GITHUB_PAT", NA_character_)
-    if(is.na(token)){
+    if(is.na(token)) {
         token <- Sys.getenv("GITHUB_TOKEN", NA_character_)
     }
-    if(is.na(token)){
+    if(is.na(token)) {
         token <- ""
     }
-    config <- httr::add_headers(Accept = "application/vnd.github.v3+json",Authorization=token)
+    config <- httr::add_headers(Accept = "application/vnd.github.v3+json", Authorization = paste0("token ", token))
     params <- list(ref = ref,...)
     request_results <- httr::GET(httr::modify_url(url, path = path), config, query = params)
     status_code <- httr::status_code(request_results)

### 2023-07-30 tests/testthat/test_resolve.R

@@ -146,6 +146,8 @@ test_that("Non-cran must enforce caching ref #22", {
 ## })
 
 test_that("Integration of as_pkgrefs() in resolve() for sessionInfo()", {
+    skip_if_offline()
+    skip_on_cran()
     x <- resolve(c("cran::sna"), snapshot_date = "2020-05-01", query_sysreqs = FALSE)
     si <- readRDS("../testdata/sessionInfo2.RDS")
     expect_error(graph <- resolve(si, snapshot_date = "2020-05-01", query_sysreqs = FALSE), NA)

### 2023-07-30 paper/paper_r0.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%

\PassOptionsToPackage{table}{xcolor}

\documentclass[
  10pt,
  letterpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% marvosym package for additional characters
\usepackage{marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% Using natbib instead
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[numbers,square,comma]{natbib}
\bibliographystyle{plos2015}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={rang: Reconstructing reproducible R computational environments},
  pdfkeywords={R, reproducibility, docker},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{rang: Reconstructing reproducible R computational
environments} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
\\
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
Chung-hong Chan\textsuperscript{1*}, David Schoch\textsuperscript{1}
\\
\bigskip
\textbf{1} GESIS Leibniz-Institut für Sozialwissenschaften, 
\bigskip

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* chung-hong.chan@gesis.org

\end{flushleft}

\section*{Abstract}
A complete declarative description of the computational environment is
often missing when researchers share their materials. Without such
description, software obsolescence and missing system components can
jeopardize computational reproducibility in the future, even when data
and computer code are available. The R package rang is a complete
solution for generating the declarative description for other
researchers to automatically reconstruct the computational environment
at a specific time point. The reconstruction process, based on Docker,
has been tested for R code as old as 2001. The declarative description
generated by rang satisfies the definition of a reproducible research
compendium and can be shared as such. In this contribution, we show how
rang can be used to make otherwise unexecutable code, spanning from
fields such as computational social science and bioinformatics,
executable again. We also provide instructions on how to use rang to
construct reproducible and shareable research compendia of current
research. The package is currently available from CRAN
(https://cran.r-project.org/web/packages/rang/index.html) and GitHub
(https://github.com/chainsawriot/rang).

\linenumbers\hypertarget{background}{%
\section{Background}\label{background}}

\emph{``In some cases the polarization estimation will not work \ldots{}
This is \emph{NOT} a problem in the method, it is entirely dependent on
the numpy version (and even the OS's). If you have different versions of
numpy or even the same version of numpy on a different OS configuration,
different networks will fail randomly\ldots{} {[}F{]}or instance, the
109th Congress will fail, but will work entirely normally on a different
numpy version, which will fail on a different Congress network.''}

\begin{quote}
\textasciitilde{} excerpt of
\href{https://www.michelecoscia.com/?page_id=2105}{the README file} of a
software for polarization estimation
\end{quote}

Other than bad programming practices \citep{trisovic:2022}, the main
computing barrier to computational reproducibility is the failure to
reconstruct the computational environment like the one used by the
original researchers. This task looks trivially simple. But as computer
science research has shown, this task is incredibly complex
\citep{abate:2015:MCR, dolstra:2010:N}. For a usual scripting language
such as R \footnote{In this paper, we will focus on R, a popular
  programming language used frequently in various computational fields
  (e.g.~computational social science, bioinformatics).}, that pertains
to four aspects: A) operating system, B) system components such as
\texttt{libxml2}, C) the exact R version, and D) what and which version
of the installed R packages. We will call them Component A, B, C, D in
the following sections. Any change in these four components can possibly
affect the execution of any shared computer code. For example, the lack
of the system component \texttt{libxml2} can impact whether the R
package \texttt{xml2} can be installed on a Linux system. If the shared
computer code requires the R package \texttt{xml2}, then the whole
execution fails.

In reality, the impact of Component A is relatively weak as mainstream,
open source programming languages and their software libraries are
usually cross platform. In modern computational research, Linux is the
de-facto operating system in high performance computing environments
(e.g.~Slurm). Instead, the impact of Components B, C, and D is much
higher. Component D is the most volatile among them all as there are
many possible combinations of R packages and versions. Software updates
with breaking changes (even in a dependency) might render existing
shared code using those changed features not executable or not producing
the same result anymore. Also, software obsolescence is commonplace,
especially since academic software is often not well maintained due to
lack of incentives \citep{merow:2023:B}.

The DevOps (software development and IT operations) community is also
confronted with this problem. The issue is usually half-jokingly
referred to as ``it works on my machine''-problem \citep[a software
works on someone's local machine but is not working anymore when
deployed to the production system, indicates the software tacitly
depends on the computational environment of the local
machine]{valstar:2020:UDS}. A partial solution to this problem from the
DevOps community is called \emph{containerization}. In essence, to
containerize is to develop and deploy the software together with all the
libraries and the operating system in an OS-level virtualization
environment. In this way, software dependency issues can be resolved
inside the isolated virtualized software environment and independent of
what is installed on the local computer. Docker is a popular choice in
the DevOps world for containerization.

To build a container, one needs to write a plain text declarative
description of the required computational environment. Inside this
declarative description, it should pin down all four Components
mentioned above. For Docker, it is in the form of a plain text file
called \texttt{Dockerfile}. This \texttt{Dockerfile} is then used as the
recipe to build a Docker image, where the four Components are assembled.
Then, one can launch a container with the built Docker image.

There has been many papers written on how containerization solutions
such as Docker can be helpful also to foster computational
reproducibility of science
\citep[e.g.][]{nuest:2019, peikert:2021:RDA, boettiger:2017:IR}.
Although tutorials are available \citep[e.g.][]{nuest:2019}, providing a
declarative description of the computational environment in the form of
\texttt{Dockerfile} is far from the standard code sharing practice. This
might be due to a lack of (DevOps) skills of most scientists to create a
\texttt{Dockerfile} \citep{kim:2018:E}. But there are many tools
available to automate the process \citep[e.g.][]{nuest:2019}. The case
in point described in this paper, \texttt{rang}, is one of them. We
argue that \texttt{rang} is the only easy-to-use solution available that
can pin down and restore all four components without the reliance on any
commercial service .

\hypertarget{existing-solutions}{%
\subsection{Existing solutions}\label{existing-solutions}}

\texttt{renv} \citep{renvrpkg} (and its derivatives such as
\texttt{jetpack} and its predecessor \texttt{packrat}) takes a similar
approach to Python's \texttt{virtualenv} and Ruby's \texttt{Gem} to pin
down the exact version of R packages using a ``lock file''. Other
solutions such as \texttt{checkpoint} \citep{checkpointrpkg} depend on
the availability of The Microsoft R Application Network (MRAN, a
time-stamped daily backup of CRAN), which will be shut down on July 1st,
2023. \texttt{groundhog} \citep{groundhogrpkg} used to depend on MRAN
but has a plan to switch to their home-grown R package repository. These
solution can effectively pin down Component C and D. But they can only
restore component D. Also, for solutions depending on MRAN, there is a
limit on how far back this reproducibility can go, since MRAN can only
go back as far as September 17, 2014. Additionally, it only covers CRAN
packages.

\texttt{containerit} \citep{nuest:2019} takes the current state of the
computational environment and documents it as a \texttt{Dockerfile}.
\texttt{containerit} makes the assumption that Component A has a weak
influence on computational reproducibility and therefore defaults to
Linux-based Rocker base images \citep{boettiger:2017:IR}. In this way,
it fixes Component A. But \texttt{containerit} does not pin down the
exact version of R packages. Therefore, it can pin down components A, B,
C, but only a part of component D. \texttt{dockta} \footnote{\url{https://stencila.github.io/dockta/}}
is another containerization solution that can potentially pin down all
components due to the fact that MRAN is used. But it also suffers from
the same limitations mentioned above.

It is also worth mentioning that MRAN is not the only archival service.
Posit also provides a free (\emph{gratis}) time-stamped daily backup of
CRAN and Bioconductor (a series of repositories of R package for
bioinformatics and computational biology) called Posit Public Package
Manager \footnote{\url{https://packagemanager.rstudio.com/client/\#/repos/2/packages/}}.
It can goes as far back as October 10, 2017.

These solutions are better for prospective usage, i.e.~using them now to
ensure the reproducibility of the current research for future
researchers. \texttt{rang} mostly targets retrospective usage,
i.e.~using \texttt{rang} to reconstruct historical R computational
environments for which the declarative descriptions are not available.
One can think of \texttt{rang} as an archaeological tool. In this realm,
we could not find any existing solution targeting R specifically which
does not currently depend on MRAN.

\hypertarget{basic-usage}{%
\section{Basic usage}\label{basic-usage}}

There are two important functions of \texttt{rang}: \texttt{resolve()}
and \texttt{dockerize()}.

\texttt{resolve()} queries various web services from the r-hub project
of the R Consortium for information about R packages at a specific time
point that is necessary for reconstructing a computational environment,
e.g.~(deep) dependencies (Component D), R version (Component C), and
system requirements (Component B). For instance, if there was a
computational environment constructed on 2020-01-16 (called ``snapshot
date'') with the several natural language processing R packages,
\texttt{resolve()} can be used to resolve all the dependencies of these
R packages. Currently, \texttt{rang} supports CRAN, Bioconductor,
GitHub, and local packages.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(rang)}
\NormalTok{graph }\OtherTok{\textless{}{-}} \FunctionTok{resolve}\NormalTok{(}\AttributeTok{pkgs =} \FunctionTok{c}\NormalTok{(}\StringTok{"openNLP"}\NormalTok{, }\StringTok{"LDAvis"}\NormalTok{, }\StringTok{"topicmodels"}\NormalTok{, }\StringTok{"quanteda"}\NormalTok{),}
                 \AttributeTok{snapshot\_date =} \StringTok{"2020{-}01{-}16"}\NormalTok{)}
\NormalTok{graph}
\end{Highlighting}
\end{Shaded}

The resolved result is an S3 object called \texttt{rang}. The
information contained in a \texttt{rang} object can then be used to
construct a computational environment in a similar manner as
\texttt{containerit}, but with the packages and R versions pinned on the
snapshot date. Then, the function \texttt{dockerize()} is used to
generate the \texttt{Dockerfile} and other scripts in the
\texttt{output\_dir}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{dockerize}\NormalTok{(graph, }\AttributeTok{output\_dir =} \StringTok{"docker"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

For R \textgreater= 3.1, the base images from the Rocker project are
used \citep{boettiger:2017:IR}. For R \textless{} 3.1 but \textgreater=
1.3.1, a custom base image based on Debian is used. As of writing,
\texttt{rang} does not support R \textless{} 1.3.1, i.e.~snapshot date
earlier than 2001-08-31 (which is 13 years earlier than all solutions
depending on MRAN). There are two features of \texttt{dockerize()} that
are important for future reproducibility.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  By default, the Docker image building process downloads source
  packages from their sources and then compiles them. This step depends
  on the future availability of R packages on CRAN (which is extremely
  likely to be the case in the near future, given the continuous
  availability since 1997-04-23) \footnote{\url{https://stat.ethz.ch/pipermail/r-announce/1997/000001.html}},
  Bioconductor, and Github. However, it is also possible to cache (or
  archive) the source packages now. The archived R packages can then be
  used instead during the building process. The significance of this
  step in terms of long-term computational reproducibility will be
  discussed in Section 4.
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{dockerize}\NormalTok{(graph, }\AttributeTok{output\_dir =} \StringTok{"docker"}\NormalTok{, }\AttributeTok{cache =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  It is also possible to install R packages in a separate library during
  the building process to isolate all these R packages from the main
  library.
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{dockerize}\NormalTok{(graph, }\AttributeTok{output\_dir =} \StringTok{"docker"}\NormalTok{, }\AttributeTok{cache =} \ConstantTok{TRUE}\NormalTok{,}
          \AttributeTok{lib =} \StringTok{"anotherlibrary"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

For the sake of completeness, the instructions for building the Docker
image and running the Docker container on Unix-like systems are included
here.

\begin{Shaded}
\begin{Highlighting}[]
\BuiltInTok{cd}\NormalTok{ docker}
\CommentTok{\#\# might need to sudo}
\ExtensionTok{docker}\NormalTok{ build }\AttributeTok{{-}t}\NormalTok{ rangimg .}
\CommentTok{\#\# interactive environment}
\ExtensionTok{docker}\NormalTok{ run }\AttributeTok{{-}{-}rm} \AttributeTok{{-}{-}name} \StringTok{"rangcontainer"} \AttributeTok{{-}ti}\NormalTok{ rangimg}
\end{Highlighting}
\end{Shaded}

\hypertarget{project-scanning}{%
\subsection{Project scanning}\label{project-scanning}}

The first argument of \texttt{resolve()} is processed by a separate
function called \texttt{as\_pkgrefs()}. For interoperability,
\texttt{rang} supports the ``package references'' standard \footnote{\url{https://r-lib.github.io/pkgdepends/reference/pkg_refs.html}}
used also in other packages such as \texttt{renv} \citep{renvrpkg}. It
is mostly used for converting ``shorthands'' (e.g.~\texttt{xml2} and
\texttt{S4Vectors}) to package references (e.g.~\texttt{cran::xml2} and
\texttt{bioc::S4Vectors}).

When \texttt{as\_pkgrefs()} is applied to a single path of a directory,
it scans all relevant files (\texttt{DESCRIPTION}, R scripts and R
Markdown files) for all R packages used (based on
\texttt{renv::dependencies()} ). How it works is demonstrated in three
of the following examples below. But an important caveat is that it can
only scan CRAN and Bioconductor packages.

\hypertarget{case-studies}{%
\section{Case Studies}\label{case-studies}}

The following are some examples of how \texttt{rang} can be used to make
shared, but otherwise unexecutable, R code runnable again. The examples
were drawn from various fields spanning from political science,
psychological science, and bioinformatics.

\hypertarget{quanteda-joss-paper}{%
\subsection{quanteda JOSS paper}\label{quanteda-joss-paper}}

The software paper of the text analysis R package \texttt{quanteda} was
published on 2018-10-06 \citep{benoit:2018}. In the paper, the following
R code snippet is included.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(}\StringTok{"quanteda"}\NormalTok{)}
\CommentTok{\# construct the feature co{-}occurrence matrix}
\NormalTok{examplefcm }\OtherTok{\textless{}{-}}
\FunctionTok{tokens}\NormalTok{(data\_corpus\_irishbudget2010, }\AttributeTok{remove\_punct =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{tokens\_tolower}\NormalTok{() }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{tokens\_remove}\NormalTok{(}\FunctionTok{stopwords}\NormalTok{(}\StringTok{"english"}\NormalTok{), }\AttributeTok{padding =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
\FunctionTok{fcm}\NormalTok{(}\AttributeTok{context =} \StringTok{"window"}\NormalTok{, }\AttributeTok{window =} \DecValTok{5}\NormalTok{, }\AttributeTok{tri =} \ConstantTok{FALSE}\NormalTok{)}
\CommentTok{\# choose 30 most frequency features}
\NormalTok{topfeats }\OtherTok{\textless{}{-}} \FunctionTok{names}\NormalTok{(}\FunctionTok{topfeatures}\NormalTok{(examplefcm, }\DecValTok{30}\NormalTok{))}
\CommentTok{\# select the top 30 features only, plot the network}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{100}\NormalTok{)}
\FunctionTok{textplot\_network}\NormalTok{(}\FunctionTok{fcm\_select}\NormalTok{(examplefcm, topfeats), }\AttributeTok{min\_freq =} \FloatTok{0.8}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

On 2023-02-08, this code snippet is not executable with the current
version of \texttt{quanteda} (3.2.4). It is possible to install the
``period appropriate'' version of \texttt{quanteda} (1.3.4) using
\texttt{remotes} on the current version of R (4.2.2). And indeed, the
above code snippet can still be executed.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_version}\NormalTok{(}\StringTok{"quanteda"}\NormalTok{, }\AttributeTok{version =} \StringTok{"1.3.4"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The issue is that installing \texttt{quanteda} 1.3.4 this way installs
the latest dependencies from CRAN. \texttt{quanteda} 1.3.4 uses a
deprecated (but not yet removed) function of \texttt{Matrix}
(\texttt{as(\textless{}dgTMatrix\textgreater{},\ "dgCMatrix")}). If this
function were removed in the future, the above code snippet would not be
executable anymore.

Using \texttt{rang}, one can query the version of \texttt{quanteda} on
2018-10-06 and create a Docker container with all the ``period
appropriate'' dependencies. Here, the \texttt{rstudio} Rocker image is
selected.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(rang)}
\NormalTok{graph }\OtherTok{\textless{}{-}} \FunctionTok{resolve}\NormalTok{(}\AttributeTok{pkgs =} \StringTok{"quanteda"}\NormalTok{,}
                 \AttributeTok{snapshot\_date =} \StringTok{"2018{-}10{-}06"}\NormalTok{,}
                 \AttributeTok{os =} \StringTok{"ubuntu{-}18.04"}\NormalTok{)}
\FunctionTok{dockerize}\NormalTok{(graph, }\AttributeTok{output\_dir =} \StringTok{"quanteda\_docker"}\NormalTok{,}
          \AttributeTok{image =} \StringTok{"rstudio"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The above code snippet can be executed with the generated container
without any problem Fig~\ref{fig1}.

% Place figure captions after the first paragraph in which they are cited.
\begin{figure}[!h]
\caption{The code snippet running in a R 3.5.1 container created with rang.}
\label{fig1}
\end{figure}

\hypertarget{psychological-science}{%
\subsection{Psychological Science}\label{psychological-science}}

Crüwell et al. \citep{cruewell:2023:WB} evaluate the computational
reproducibility of 14 articles published in \emph{Psyhocological
Science}. Among these articles, the paper by Hilgard et al.
\citep{hilgard:2019:NEG} has been rated as having ``package dependency
issues''.

All data and computer code are available from GitHub with the last
commit on 2019-01-17 \footnote{\url{https://github.com/Joe-Hilgard/vvg-2d4d}}.
The R code contains a list of R packages used in the project as
\texttt{library()} statements, including an R package on GitHub that is
written by the main author of that paper. However, we identified one
package (\texttt{compute.es}) that was not written in those
\texttt{library()} statements but used with the namespace operator,
i.e.~\texttt{compute.es::tes()}. This undocumented package can be
detected by \texttt{renv::dependencies()}, which is the provider of the
scanning function of \texttt{rang}.

Based on the above information, one can run \texttt{resolve()} to obtain
the dependency graph of all R packages on 2019-01-17.

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# scan all packages}
\NormalTok{r\_pkgs }\OtherTok{\textless{}{-}} \FunctionTok{as\_pkgrefs}\NormalTok{(}\StringTok{"vvg{-}2d4d"}\NormalTok{)}
\DocumentationTok{\#\# replace cran::hilgard with Github}
\NormalTok{r\_pkgs[r\_pkgs }\SpecialCharTok{==} \StringTok{"cran::hilhard"}\NormalTok{] }\OtherTok{\textless{}{-}} \StringTok{"Joe{-}Hilgard/hilgard"}
\NormalTok{graph }\OtherTok{\textless{}{-}} \FunctionTok{resolve}\NormalTok{(r\_pkgs, }\AttributeTok{snapshot\_date =} \StringTok{"2019{-}01{-}17"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

When running \texttt{dockerize()}, one can take advantage of the
\texttt{materials\_dir} parameter to transfer the shared materials from
Hilgard et al. \citep{hilgard:2019:NEG} into the Docker image.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{dockerize}\NormalTok{(graph, }\StringTok{"hilgard"}\NormalTok{, }\AttributeTok{materials\_dir =} \StringTok{"vvg{-}2d4d"}\NormalTok{, }\AttributeTok{cache =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We then built the Docker and launch a Docker container. For this
container, we changed the entry point from R to bash so that the
container goes to the Linux command shell instead.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cd hilgard}
\NormalTok{docker build }\SpecialCharTok{{-}}\NormalTok{t hilgard .}
\NormalTok{docker run }\SpecialCharTok{{-}{-}}\NormalTok{rm }\SpecialCharTok{{-}{-}}\NormalTok{name }\StringTok{"hilgardcontainer"} \SpecialCharTok{{-}{-}}\NormalTok{entrypoint bash }\SpecialCharTok{{-}}\NormalTok{ti hilgard}
\end{Highlighting}
\end{Shaded}

Inside the container, the materials are located in the
\texttt{materials} directory. We used the following shell script to test
the reproducibility of all R scripts.

\begin{Shaded}
\begin{Highlighting}[]
\BuiltInTok{cd}\NormalTok{ materials}
\VariableTok{rfiles}\OperatorTok{=}\VariableTok{(}\NormalTok{0\_data\_aggregation.R 1\_data\_cleaning.R 2\_analysis.R 3\_plotting.R}\VariableTok{)}
\ControlFlowTok{for}\NormalTok{ i }\KeywordTok{in} \VariableTok{$\{rfiles}\OperatorTok{[@]}\VariableTok{\}}
\ControlFlowTok{do}
    \ExtensionTok{Rscript} \VariableTok{$i}
    \VariableTok{code}\OperatorTok{=}\VariableTok{$?}
    \ControlFlowTok{if} \BuiltInTok{[} \VariableTok{$code} \OtherTok{!=}\NormalTok{ 0 }\BuiltInTok{]}
    \ControlFlowTok{then}
        \BuiltInTok{exit}\NormalTok{ 1}
    \ControlFlowTok{fi}
\ControlFlowTok{done}
\end{Highlighting}
\end{Shaded}

All R scripts ran fine inside the container and the figures generated
are the same as the ones in Hilgard et al. \citep{hilgard:2019:NEG}.

\hypertarget{political-analysis}{%
\subsection{Political Analysis}\label{political-analysis}}

The study by Trisovic et al. \citep{trisovic:2022} evaluates the
reproducibility of R scripts shared on Dataverse. They found that 75\%
of R scripts cannot be successfully executed. Among these failed R
scripts is an R script shared by Beck \citep{beck:2019:EGD}.

This R script has been ``rescued'' by the author of the R package
\texttt{groundhog} \citep{groundhogrpkg}, as demonstrated in a blog post
\footnote{\url{http://datacolada.org/100}}. We were wondering if
\texttt{rang} can also be used to ``rescue'' the concerned R script. The
date of the R script, as indicated on Dataverse, is 2018-12-12. This
date is used as the snapshot date.

\begin{Shaded}
\begin{Highlighting}[]
\DocumentationTok{\#\# as\_pkgrefs is automatically run in this case}
\NormalTok{graph }\OtherTok{\textless{}{-}} \FunctionTok{resolve}\NormalTok{(}\StringTok{"nathaniel"}\NormalTok{, }\AttributeTok{snapshot\_date =} \StringTok{"2018{-}12{-}12"}\NormalTok{)}
\FunctionTok{dockerize}\NormalTok{(graph, }\AttributeTok{output\_dir =} \StringTok{"nat"}\NormalTok{, }\AttributeTok{materials\_dir =} \StringTok{"nathaniel"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cd nat}
\NormalTok{docker build }\SpecialCharTok{{-}}\NormalTok{t nat .}
\NormalTok{docker run }\SpecialCharTok{{-}{-}}\NormalTok{rm }\SpecialCharTok{{-}{-}}\NormalTok{name }\StringTok{"natcontainer"} \SpecialCharTok{{-}{-}}\NormalTok{entrypoint bash }\SpecialCharTok{{-}}\NormalTok{ti nat}
\end{Highlighting}
\end{Shaded}

Inside the container

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{cd materials}
\NormalTok{Rscript fn\_5.R}
\end{Highlighting}
\end{Shaded}

The same file can thus also be ``rescued'' by \texttt{rang}.

\hypertarget{recover-a-removed-r-package-maxent}{%
\subsection{Recover a removed R package:
maxent}\label{recover-a-removed-r-package-maxent}}

The R package \texttt{maxent} introduces a machine learning algorithm
with a small memory footprint and was available on CRAN until 2019. A
software paper was published by the original authors in 2012
\citep{jurka:2012}. The R package was also used in some subsequent
automated content analytic papers \citep[e.g.][]{loercher:2017:D}.
Despite the covert editing of the package by a staffer of CRAN
\footnote{\url{https://github.com/cran/maxent/commit/9d46c6aad27a1f41a78907b170ddd9a586192be9}},
the package was removed from CRAN in 2019 \footnote{This is the check
  log:
  \url{https://cran-archive.r-project.org/web/checks/2019/2019-03-05_check_results_maxent.html}
  Here, we said the package \texttt{maxent} was ``removed from CRAN'' as
  per what the webpage
  (\url{https://cran.r-project.org/web/packages/maxent/index.html}) is
  written. However, the terminology as stated in the
  \href{https://cran.r-project.org/web/packages/policies.html\#Source-packages}{CRAN
  Policies} is ``archived'', quote: ``Packages will not normally be
  removed from CRAN: however, they may be archived, including at the
  maintainer's request.'' As all CRAN packages have all the submitted
  versions archived, we find this terminology confusing. Therefore, we
  use ``removed'' packages throughout this paper to indicate packages
  that cannot be installed by the usual method,
  i.e.~\texttt{install.packages} but still have the old versions
  archived on CRAN.}. We attempted to install the second last (the
original submitted version) and last (with covert editing) versions of
\texttt{maxent} on R 4.2.2. Both of them didn't work.

Using \texttt{rang}, we are able to reconstruct a computational
environment with R 2.15.0 (2012-03-30) to run all code snippets
published in Jurka \citep{jurka:2012} \footnote{On an interesting
  historical side note: The original paper reported ---based on a
  benchmark--- that ``the algorithm is very fast; \texttt{maxent} uses
  only 135.4 megabytes of RAM and finishes in 53.3 seconds.'' On a
  modest computer in 2023 with a dockerized R 2.15.0, the benchmark
  finishes in 4 seconds.}. For removed CRAN packages, we strongly
recommend querying the Github read-only mirror of CRAN instead
(https://github.com/cran). It is because in this way, the resolved
system requirements have a higher chance of being correct.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{maxent }\OtherTok{\textless{}{-}} \FunctionTok{resolve}\NormalTok{(}\StringTok{"cran/maxent"}\NormalTok{, }\StringTok{"2012{-}06{-}10"}\NormalTok{)}
\FunctionTok{dockerize}\NormalTok{(maxent, }\StringTok{"maxentdir"}\NormalTok{, }\AttributeTok{cache =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{recover-a-removed-r-package-ptproc}{%
\subsection{Recover a removed R package:
ptproc}\label{recover-a-removed-r-package-ptproc}}

The software paper of the R package \texttt{ptproc} was published in
2003 and introduced multidimensional point process models
\citep{peng:2003:MDP}. But the package has been removed from CRAN for
over a decade (at least). The only release on CRAN was on 2002-10-10.
The package is still listed in the ``Handling and Analyzing
Spatio-Temporal Data'' CRAN Task View \footnote{\url{https://cran.r-project.org/web/views/SpatioTemporal.html}}
despite being uninstallable without modification on any modern R system
(see below). As of writing, the package, as a tarball file (tar.gz), is
still downloadable from the original author's website \footnote{\url{https://www.biostat.jhsph.edu/~rpeng/software/}}.

Even with this over-a-decade removal and new packages with similar
functionalities have been created, there is evidence that
\texttt{ptproc} is still being sought for. As late as 2017, there are
blog posts on how to install the long obsolete package on modern
versions of R \footnote{\url{https://blog.mathandpencil.com/installing-ptproc-on-osx}
  and
  \url{https://tomaxent.com/2017/03/16/Installing-ptproc-on-Ubuntu-16-04-LTS/}}.
The package is extremely challenging to install on a modern R system
because the package was written before the introduction of name space
management in R 1.7.0 \citep{RN-2003-001}. In other words, the available
tarball files from the original author's website and CRAN do not contain
a \texttt{NAMESPACE} file like all other modern R packages do.

The oldest version of R that \texttt{rang} can support, as of writing,
is R 1.3.1. \texttt{rang} is probably the only solution available that
can support the 1.x series of R (i.e.~before 2004-10-04). Similar to the
case of \texttt{maxent} above, a \texttt{Dockerfile} to assemble a
Docker image with \texttt{ptproc} installed can be generated with two
lines of code.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{graph }\OtherTok{\textless{}{-}} \FunctionTok{resolve}\NormalTok{(}\StringTok{"ptproc"}\NormalTok{, }\AttributeTok{snapshot\_date =} \StringTok{"2004{-}07{-}01"}\NormalTok{)}
\FunctionTok{dockerize}\NormalTok{(graph, }\StringTok{"\textasciitilde{}/dev/misc/ptproc"}\NormalTok{, }\AttributeTok{cache =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Suppose we have an R script, extracted from Peng \citep{peng:2003:MDP},
called ``peng.R'' like this:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{require}\NormalTok{(ptproc)}

\FunctionTok{set.seed}\NormalTok{(}\DecValTok{1000}\NormalTok{)}
\NormalTok{x }\OtherTok{\textless{}{-}} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{runif}\NormalTok{(}\DecValTok{100}\NormalTok{), }\FunctionTok{runif}\NormalTok{(}\DecValTok{100}\NormalTok{), }\FunctionTok{runif}\NormalTok{(}\DecValTok{100}\NormalTok{))}
\NormalTok{hPois.cond.int }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(params, eval.pts, }\AttributeTok{pts =} \ConstantTok{NULL}\NormalTok{,}
                           \AttributeTok{data =} \ConstantTok{NULL}\NormalTok{, }\AttributeTok{TT =} \ConstantTok{NULL}\NormalTok{) \{}
\NormalTok{    mu }\OtherTok{\textless{}{-}}\NormalTok{ params[}\DecValTok{1}\NormalTok{]}
    \ControlFlowTok{if}\NormalTok{(}\FunctionTok{is.null}\NormalTok{(TT))}
        \FunctionTok{rep}\NormalTok{(mu, }\FunctionTok{nrow}\NormalTok{(eval.pts))}
    \ControlFlowTok{else}\NormalTok{ \{}
\NormalTok{        vol }\OtherTok{\textless{}{-}} \FunctionTok{prod}\NormalTok{(}\FunctionTok{apply}\NormalTok{(TT, }\DecValTok{2}\NormalTok{, diff))}
\NormalTok{        mu }\SpecialCharTok{*}\NormalTok{ vol}
\NormalTok{    \}}
\NormalTok{\}}
\NormalTok{ppm }\OtherTok{\textless{}{-}} \FunctionTok{ptproc}\NormalTok{(}\AttributeTok{pts =}\NormalTok{ x, }\AttributeTok{cond.int =}\NormalTok{ hPois.cond.int, }\AttributeTok{params =} \DecValTok{50}\NormalTok{,}
              \AttributeTok{ranges =} \FunctionTok{cbind}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{), }\FunctionTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{,}\DecValTok{1}\NormalTok{)))}
\NormalTok{fit }\OtherTok{\textless{}{-}} \FunctionTok{ptproc.fit}\NormalTok{(ppm, }\AttributeTok{optim.control =} \FunctionTok{list}\NormalTok{(}\AttributeTok{trace =} \DecValTok{2}\NormalTok{), }\AttributeTok{method =} \StringTok{"BFGS"}\NormalTok{)}
\FunctionTok{summary}\NormalTok{(fit)}
\end{Highlighting}
\end{Shaded}

One can integrate \texttt{rang} into a BASH script to completely
automate the batch execution of the above R script.

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{Rscript} \AttributeTok{{-}e} \StringTok{"require(rang); dockerize(resolve(\textquotesingle{}ptproc\textquotesingle{}, \textquotesingle{}2004{-}07{-}01\textquotesingle{}),}
\StringTok{\textquotesingle{}pengdocker\textquotesingle{}, cache = TRUE)"}
\ExtensionTok{docker}\NormalTok{ build }\AttributeTok{{-}t}\NormalTok{ pengimg ./pengdocker}
\CommentTok{\#\# launching a container in daemon mode {-}d}
\ExtensionTok{docker}\NormalTok{ run }\AttributeTok{{-}d} \AttributeTok{{-}{-}rm} \AttributeTok{{-}{-}name} \StringTok{"pengcontainer"} \AttributeTok{{-}ti}\NormalTok{ pengimg}
\ExtensionTok{docker}\NormalTok{ cp peng.R pengcontainer:/peng.R}
\ExtensionTok{docker}\NormalTok{ exec pengcontainer R CMD BATCH peng.R}
\ExtensionTok{docker}\NormalTok{ exec pengcontainer cat peng.Rout}
\ExtensionTok{docker}\NormalTok{ cp pengcontainer:/peng.Rout peng.Rout}
\ExtensionTok{docker}\NormalTok{ stop pengcontainer}
\end{Highlighting}
\end{Shaded}

The file \texttt{peng.Rout} contains the execution results of the script
from inside the Docker container. As the random seed was preserved by
the original author \citep{peng:2003:MDP}, the above BASH script can
perfectly reproduce the analysis \footnote{It is also important to note
  that the random number generator (RNG) of R has been changed several
  times over the course of the development. In this case, we are using
  the same generation of RNG as Peng \citep{peng:2003:MDP}.}.

\hypertarget{recover-a-removed-bioconductor-package}{%
\subsection{Recover a removed Bioconductor
package}\label{recover-a-removed-bioconductor-package}}

Similar to CRAN, packages can also be removed over time from
Bioconductor. The Bioconductor package \texttt{Sushi} has been
deprecated by the original authors and is removed from Bioconductor
version 3.16 (2022-11-02). \texttt{Sushi} is a data visualization tool
for genomic data and was used in many online tutorials and scientific
papers, including the original paper announcing the package by the
original authors \citep{phanstiel:2014:S}.

\texttt{rang} has native support for Bioconductor packages since version
0.2. We obtained the R script \texttt{"PaperFigure.R"} from the Github
repository of \texttt{Sushi} \footnote{\url{https://github.com/PhanstielLab/Sushi/blob/master/vignettes/PaperFigure.R}},
which generates the figure in the original paper
\citep{phanstiel:2014:S}. Similar to the above case of \texttt{ptproc},
we made a completely automated BASH script to run
\texttt{"PaperFigure.R"} and get the generated figure out of the
container (Fig~\ref{fig2}). We made no modification to
\texttt{"PaperFigure.R"}.

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{Rscript} \AttributeTok{{-}e} \StringTok{"require(rang); dockerize(resolve(\textquotesingle{}Sushi\textquotesingle{}, \textquotesingle{}2014{-}06{-}05\textquotesingle{}),}
\StringTok{\textquotesingle{}sushidocker\textquotesingle{}, no\_rocker = TRUE, cache = TRUE)"}
\ExtensionTok{docker}\NormalTok{ build }\AttributeTok{{-}t}\NormalTok{ sushiimg ./sushidocker}
\ExtensionTok{docker}\NormalTok{ run }\AttributeTok{{-}d} \AttributeTok{{-}{-}rm} \AttributeTok{{-}{-}name} \StringTok{"sushicontainer"} \AttributeTok{{-}ti}\NormalTok{ sushiimg}
\ExtensionTok{docker}\NormalTok{ cp PaperFigure.R sushicontainer:/PaperFigure.R}
\ExtensionTok{docker}\NormalTok{ exec sushicontainer mkdir vignettes}
\ExtensionTok{docker}\NormalTok{ exec sushicontainer R CMD BATCH PaperFigure.R}
\ExtensionTok{docker}\NormalTok{ cp sushicontainer:/vignettes/Figure\_1.pdf sushi\_figure1.pdf}
\ExtensionTok{docker}\NormalTok{ stop sushicontainer}
\end{Highlighting}
\end{Shaded}

% Place figure captions after the first paragraph in which they are cited.
\begin{figure}[!h]
\caption{The figure from the batch execution of PaperFigure.R inside a Docker container generated by rang}
\label{fig2}
\end{figure}

\hypertarget{preparing-research-compendia-with-long-term-computational-reproducibility}{%
\section{Preparing research compendia with long-term computational
reproducibility}\label{preparing-research-compendia-with-long-term-computational-reproducibility}}

The above six examples show how powerful \texttt{rang} is to reconstruct
tricky computational environments which have not been completely
declared in the literature. Although we position \texttt{rang} mostly as
an archaeological tool, we think that \texttt{rang} can also be used to
prepare research compendia of current research. We can't predict the
future but research compendia generated by \texttt{rang} would probably
have long-term computational reproducibility.

To demonstrate this point, we took the recent paper by Oser et al.
\citep{oser:2022:HPE}. This paper was selected because 1) the paper was
published in \emph{Political Communication}, a high impact journal that
awards Open Science Badges; 2) shared data and R code are available; and
most importantly, 3) the shared R code is well-written. In the
repository of this paper, we based on the materials shared by Oser et
al. \citep{oser:2022:HPE} and prepared a research compendium that should
have long-term computational reproducibility. The research compendium is
similar to the Executable Compendium suggested by the Turing way.

The preparation of the research compendium is easy as \texttt{rang} can
scan a materials directory for all R packages used \footnote{We detected
  a minor issue in the code base that an undeclared Github package is
  used. But it can be easily solved, as in the Psychological Science
  example above.}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{require}\NormalTok{(rang)}
\DocumentationTok{\#\# meta{-}analysis is the directory of all shared materials}
\NormalTok{cran\_pkgs }\OtherTok{\textless{}{-}} \FunctionTok{as\_pkgrefs}\NormalTok{(}\StringTok{"meta{-}analysis"}\NormalTok{) }

\DocumentationTok{\#\# dmetar is an undeclared github package: MathiasHarrer/dmetar}
\NormalTok{cran\_pkgs[cran\_pkgs }\SpecialCharTok{==} \StringTok{"cran::dmetar"}\NormalTok{] }\OtherTok{\textless{}{-}} \StringTok{"MathiasHarrer/dmetar"}
\NormalTok{x }\OtherTok{\textless{}{-}} \FunctionTok{resolve}\NormalTok{(cran\_pkgs, }\StringTok{"2021{-}08{-}11"}\NormalTok{, }\AttributeTok{verbose =} \ConstantTok{TRUE}\NormalTok{)}
\DocumentationTok{\#\#print(x, all\_pkgs = TRUE)}
\FunctionTok{dockerize}\NormalTok{(x, }\StringTok{"oserdocker"}\NormalTok{, }\AttributeTok{materials\_dir =} \StringTok{"meta{-}analysis"}\NormalTok{, }\AttributeTok{cache =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The above R script is saved as \texttt{oser.R}. The central piece of the
executable compendium is the \texttt{Makefile} \citep{baker:2020:UGM}.

\begin{Shaded}
\begin{Highlighting}[]
\DataTypeTok{output\_file}\CharTok{=}\StringTok{reproduced.html}
\DataTypeTok{r\_cmd }\CharTok{=}\StringTok{ "rmarkdown::render(\textquotesingle{}materials/README.Rmd\textquotesingle{},}\CharTok{\textbackslash{}}
\StringTok{output\_file = \textquotesingle{}}\CharTok{$\{}\DataTypeTok{output\_file}\CharTok{\}}\StringTok{\textquotesingle{})"}
\DataTypeTok{handle}\CharTok{=}\StringTok{oser}
\DataTypeTok{local\_file}\CharTok{=$\{}\DataTypeTok{handle}\CharTok{\}}\StringTok{\_README.html}

\OtherTok{.PHONY:}\DataTypeTok{ all resolve build render export rebuild}

\DecValTok{all:}\DataTypeTok{ resolve build render}
\NormalTok{    echo }\StringTok{"finished"}
\DecValTok{resolve:}
\NormalTok{    Rscript }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{.R}
\DecValTok{build:}\DataTypeTok{ }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\DataTypeTok{docker}
\NormalTok{    docker build {-}t }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{img }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{docker}
\DecValTok{render:}
\NormalTok{    docker run {-}d {-}{-}rm {-}{-}name }\StringTok{"}\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\StringTok{container"}\NormalTok{ {-}ti }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{img}
\NormalTok{    docker exec }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{container Rscript {-}e }\CharTok{$\{}\DataTypeTok{r\_cmd}\CharTok{\}}
\NormalTok{    docker cp }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{container:/materials/}\CharTok{$\{}\DataTypeTok{output\_file}\CharTok{\}} \CharTok{$\{}\DataTypeTok{local\_file}\CharTok{\}}
\NormalTok{    docker stop }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{container}
\DecValTok{export:}
\NormalTok{    docker save }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{img | gzip \textgreater{} }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{img.tar.gz}
\DecValTok{rebuild:}\DataTypeTok{ }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\DataTypeTok{img.tar.gz}
\NormalTok{    docker load \textless{} }\CharTok{$\{}\DataTypeTok{handle}\CharTok{\}}\NormalTok{img.tar.gz}
\end{Highlighting}
\end{Shaded}

With this \texttt{Makefile}, one can create the \texttt{Dockerfile} with
\texttt{make\ resolve}, build the Docker image with
\texttt{make\ build}, render the RMarkdown file inside the container
with \texttt{make\ render}, export the built Docker image with
\texttt{make\ export}, and rebuild the exported Docker image with
\texttt{make\ rebuild}.

The structure of the entire executable compendium looks like this:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{Makefile}
\ExtensionTok{oser.R}
\ExtensionTok{meta{-}analysis/}
\ExtensionTok{README.md}
\ExtensionTok{oserdocker/}
\ExtensionTok{oserimg.tar.gz}
\end{Highlighting}
\end{Shaded}

In this executable compendium, only the first four elements are
essential. The directory \texttt{oserdocker} (116 MB) contains cached R
packages, a \texttt{Dockerfile}, and a verbatim copy of the directory
\texttt{meta-analysis/} to be transferred into the Docker image. That
can be regenerated by running \texttt{make\ resolve}. However, having
this directory preserved insures against the situations that some R
packages used in the project were no longer available or any of the
information providers used by \texttt{rang} for resolving the dependency
relationships were not available. (Or in the rare circumstance of
\texttt{rang} is no longer available.)

\texttt{oserimg.tar.gz} (667 MB) is a backup copy of the Docker image.
This can be regenerated by running \texttt{make\ export}. Preserving
this file insures against all the situations mentioned above, but also
the situations of Docker Hub (the hosting service provided by Docker for
base images such as Rocker) and the software repositories used by the
dockerized operating system being not available. When
\texttt{oserimg.tar.gz} is available, it is possible to run
\texttt{make\ rebuild} and \texttt{make\ render} even without internet
access (provided that Docker and \texttt{make} have been installed
before). Of course, there is still an extremely rare situation where
Docker (the program) itself is no longer available \footnote{We can't
  imagine a world without \texttt{Make}, a tool that has been available
  since 1976.}. However, it is possible to convert the image file for
use on other containerization solutions such as Singularity \footnote{\url{https://docs.sylabs.io/guides/3.0/user-guide/singularity_and_docker.html}},
if Docker is really not available anymore.

Sharing of research artifacts less than 1G is not as challenging as it
used to be. Zenodo, for example, allows the sharing of 50G of files.
Therefore, sharing of the last two components of the executable
compendium prepared with \texttt{rang} is at least possible on Zenodo
\footnote{The complete version of the executable compendium is available
  on Zenodo: \url{https://doi.org/10.5281/zenodo.7708417}}. However, for
data repositories with more restrictions on data size, sharing the
executable compendium without the last two parts could be considered
sufficient. For that, run \texttt{make} will make the default target
\texttt{all} and generate all the things needed for reproducing the
analysis inside a container.

The above \texttt{Makefile} is general enough that one can reuse it by
just modifying how the R scripts (the \texttt{r\_cmd} variable) in the
\texttt{materials} directory are executed. This can be a starting point
of a standard executable compendium format.

\hypertarget{concluding-remarks}{%
\section{Concluding remarks}\label{concluding-remarks}}

This paper presents \texttt{rang}, a solution to (re)construct R
computational environments based on Docker. As the six examples in
Section 3 show, \texttt{rang} can be used archaeologically to rerun old
code, many of them not executable without the analytic and
reconstruction processes facilitated by \texttt{rang}. These
retrospective use cases demonstrate how versatile \texttt{rang} is.
\texttt{rang} is also helpful for prospective usage, as demonstrated in
Section 4 whereby an executable compendium is created.

There are still many features that we did not mention in this paper.
\texttt{rang} is built with interoperability in mind. As of writing,
\texttt{rang} is interoperable with existing R packages such as
\texttt{renv} and R built-in \texttt{sessionInfo()}. Also, the
\texttt{rang} object can be used for network analysis with R packages
such as \texttt{igraph}.

Computational reproducibility is a complex topic and as in all of these
complex topics, there is no silver bullet \citep{canon:2019:CPR}. All
solutions have their trade-offs. The (re)construction process based on
\texttt{rang} takes notably more time than other solutions because all
packages are compiled from source. \texttt{rang} trades computational
efficiency of this often one-off (re)constructing process for
correctness, backward compatibility and independence from any commercial
backups of software repositories such as MRAN. There are also other
limitations. In the Vignette of \texttt{rang}
(\url{https://cran.r-project.org/web/packages/rang/vignettes/faq.html}),
we list all of these limitations as well as possible mitigation.

\nolinenumbers
\renewcommand\refname{References}
  \bibliography{paper.bib}

\end{document}

### 2023-07-30 paper/r1.qmd

---
title: "Re: PONE-D-23-07706"
format:
  pdf: default
---

Dear Professor Fernandez-Lozano, dear reviewers,

Thank you for your insightful feedback on our paper "rang: Reconstructing reproducible R computational environments" (PONE-D-23-07706). Following the reviewers’ suggestions, we have revised the manuscript. We respond below through a point-by-point outline.

We hope that the revised manuscript will now be suited for publication in PLOS ONE. We are looking forward to hearing back from you.

Sincerely,

The Authors

* * *

# Reviewer #1: General comments

**R1.1: This manuscript describes the ‘rang’ package for R, focusing on six examples of using rang to reconstruct the computational environments of old scripts. This is important and interesting, for reasons well described in the introduction—basically, computational reproducibility is a hard problem, and while not a panacea this package provides useful functionality to help this. The ms is fairly well written, and the core examples are very well done in their variety and step-by-step explanations.**

A1.1 We would like to thank R1 for their appreciation of our package `rang`.

**R1.2: There are some minor problems (see below), and one major one I thought: if the purpose of this ms is to describe the rang package, i.e. to be its primary peer-reviewed documentation, it needs to list all package functions and what they do (e.g. in a table; see #12 below).**

A1.2 We agreed with R1 that a list of all package functions is needed. It has been added accordingly. (See table 1)

**R1.3: In summary, this is an interesting and well-done ms that usefully lays out how to use rang to reconstruct computational environments in a wide variety of cases. It needs minor to moderate revisions for clarity in many places, but is fundamentally a strong piece worthy of publication.**

A1.3 We would like to thank R1 for their appreciation of our paper.

**R1.4 Specific comments 1. Abstract: could probably say “usually missing”**

A1.4 We adopted the suggestion by R1.

**R1.5 2. Abstract: change “spanning from” to “spanning” (grammar)**

A1.5 We adopted the suggestion by R1.

**R1.6 3. Line 13: cite R correctly – see citation()**

A1.6 We cited R accordingly.

**R1.7 4. L. 24: this is a bit odd, as slurm is a workload scheduler, not a computing environment**

A1.7 R1 correctly pointed out the problem. We removed the mention of slurm.

**R1.8 5. L. 64: “unambiguously specify” instead of “pin down”? Seems clearer**

A1.8 We adopted the R1's suggestion to say "(unambigously) specify" instead of "pin down". And we agree that it is clearer.

**R1.9 6. L. 75: what limitations, exactly? Be specific**

A1.9 We spelled out the limitations of MRAN.

**R1.10 7. L. 159: a little unclear. “which is used by the scanning function”?**

A1.10 We stated clearly that `as_pkgrefs()` is a wrapper.

**R1.11 8. L. 186: “covert editing”? Really? Clarify, expand, or remove**

A1.11 As far as we know, the editing was done without the permission from the original submitter. However, without going too detail into similar editing by the same staffer, we decided to drop the adjective "covert".

**R1.12 9. L. 189: “Neither worked.”**

A1.12 We adopted the suggestion by the reviewer.

**R1.13 10. L. 245: what does “suggested by the Turing way” mean?**

A1.13 The Turing Way (https://the-turing-way.netlify.app/reproducible-research/compendia.html) is a handbook published by the Alan Turing Institute in the UK. We provided a citation to the handbook to make it clearer.

**R1.14 11. L. 248: use “library(rang)” (cf. code on p. 3) not “require(rang)”**

A1.14 We adopted the suggestion by the reviewer.

**R1.15 12. L. 291: there are features not mentioned? Like what? A table listing all the package functions would be a useful addition**

A1.15 See A1.2

**R1.16 13. The major rang caveats, as listed in the package vignette, should be included in this article**

A1.16 We adopted the suggestion by the reviewer to include all caveats.

# Reviewer #2:

**R2.1 The article is well written and the package works as described. They have tested rang using a wide range of examples and the steps are well documented and clear. The problem they are addressing is a very important one and I commend their effort in creating the package and producing the article.**

A2.1 We would like to thank R2 for their appreciation of our package `rang` and the article.

**R2.2 My only comment is related to the discussion of limitations. Many packages access databases to obtain data. For example, the STRINGdb package has a function to obtain protein interaction data, and the API changes over time. As such, previous versions of STRINGdb cannot access the db. For packages like biomaRt, the method to access the data may be more stable, however the actual data itself can change, for example Gene Ontology data. I suggest the authors make this limitation clear and/or suggest ways to solve, or in some way minimise it.**

A2.2 We agree with R2 that this is a major reproducibility issue. In the revised version of the paper, we added this as a limitation. `rang` is not a solution to these external dependencies and we made several suggestions.

**R2.3 I would also recommend a thorough language edit - the article is very clearly written and easy to follow but there are some small errors. For example, in the following sentence, there is some mixing of tenses, and the use of brackets seems unsual: However, having this directory preserved insures against the situations that some R packages used in the project were no longer available or any of the information providers used by rang for resolving the dependency relationships were not available. (Or in the rare circumstance of rang is no longer available.)**

A2.3 We revised the language of the paper. The sentence mentioned was due to the wrong usage of past subjunctive mood ("were"). We fixed it to use present indicative mood instead.

### 2023-10-08 tests/testthat/test_expost_rang.R
Export `generate_installation_order` fix #154

@@ -132,11 +132,11 @@ test_that("empty rang export, #75", {
 test_that("prevent infinite loop, #81", {
     graph <- readRDS("../testdata/rang_ok.RDS")
     graph$ranglets[[1]]$deps[[2]] <- NULL
-    expect_error(.generate_installation_order(graph), "cran::LDAvis")
+    expect_error(generate_installation_order(graph), "cran::LDAvis")
     graph <- readRDS("../testdata/rang_ok.RDS")
     graph$ranglets[[1]]$original$y <- "S4Vectors"
     graph$ranglets[[1]]$original$y_pkgref <- "bioc::S4Vectors"
-    expect_error(.generate_installation_order(graph), "cran::LDAvis")
+    expect_error(generate_installation_order(graph), "cran::LDAvis")
 })
 
 test_that("renv export cran", {
@@ -183,10 +183,10 @@ test_that("renv export unknown source", {
 
 test_that("Super ancient special packages", {
     graph <- readRDS("../testdata/superancientsna.RDS")
-    expect_error(.generate_installation_order(graph), NA)
+    expect_error(generate_installation_order(graph), NA)
 })
 
 test_that("base as a dependency, issue 144", {
     graph <- readRDS("../testdata/dt.RDS")
-    expect_error(.generate_installation_order(graph), NA)
+    expect_error(generate_installation_order(graph), NA)
 })

### 2023-10-08 R/cache.R
Export `generate_installation_order` fix #154 [no ci] (#162)

@@ -85,7 +85,7 @@ NULL
 }
 
 .cache_pkgs <- function(rang, base_dir, cran_mirror, bioc_mirror, verbose) {
-    installation_order <- .generate_installation_order(rang)
+    installation_order <- generate_installation_order(rang)
     cache_dir <- file.path(base_dir, "cache", "rpkgs")
     if (!dir.exists(cache_dir)) {
         dir.create(cache_dir, recursive = TRUE)

### 2023-10-08 R/installation.R

@@ -7,7 +7,22 @@
     }
 }
 
-.generate_installation_order <- function(rang) {
+#' Create a Data Frame of The Resolved Result
+#' This function exports the results from [resolve()] to a data frame, which each row represents one installation step. The order of rows is the installation order. By installing packages in the specified order, one can install all the resolved packages without conflicts.
+#' @inheritParams export_rang
+#' @return A data frame ordered by installation order.
+#' @references
+#' Ripley, B. (2005) [Packages and their Management in R 2.1.0.](https://cran.r-project.org/doc/Rnews/Rnews_2005-1.pdf) R News, 5(1):8--11.
+#' @examples
+#' \donttest{
+#' if (interactive()) {
+#'     graph <- resolve(pkgs = c("openNLP", "LDAvis", "topicmodels", "quanteda"),
+#'                     snapshot_date = "2020-01-16")
+#'     generate_installation_order(graph)
+#' }
+#' }
+#' @export
+generate_installation_order <- function(rang) {
     dep <- fastmap::fastmap()
     version <- fastmap::fastmap()
     uid <- fastmap::fastmap()
@@ -133,6 +148,7 @@
 #' @param check_cran_mirror logical, whether to check the CRAN mirror
 #' @param bioc_mirror character, which Bioconductor mirror to use
 #' @return `path`, invisibly
+#' @seealso [generate_installation_order()]
 #' @details The idea behind this is to determine the installation order of R packages locally. Then, the installation script can be deployed to another
 #' fresh R session to install R packages. [dockerize()] and [apptainerize()] are more reasonable ways because a fresh R session with all system requirements
 #' is provided. The current approach does not work in R < 2.1.0.
@@ -166,7 +182,7 @@ export_rang <- function(rang, path, rang_as_comment = TRUE, verbose = TRUE, lib 
     if (.is_r_version_older_than(rang, "3.3")) { #20
         cran_mirror <- .normalize_url(cran_mirror, https = FALSE)
     }
-    installation_order <- .generate_installation_order(rang)
+    installation_order <- generate_installation_order(rang)
     file.create(path)
     con <- file(path, open="w")
     if (.is_r_version_older_than(rang, "2.1")) {
@@ -219,7 +235,7 @@ export_renv <- function(rang, path = ".") {
         warning("Nothing to export.")
[…]

### 2023-10-08 R/memo_misc.R
Bump version and update internal data [no ci]

@@ -110,7 +110,7 @@ NULL
 
 ## internal data generation
 ## ---
-## ### Supported OS Versions
+### Supported OS Versions
 ## supported_os <- c("trusty" = "ubuntu-14.04", "xenial" = "ubuntu-16.04", "bionic" = "ubuntu-18.04", "focal" = "ubuntu-20.04", "centos-6", "centos-7", "centos-8", "redhat-6", "redhat-7", "redhat-8")
 ## ### R version history
 ## cached_rver <- .rver()

### 2023-10-08 README.Rmd
Update citation information
Update handle [no ci]

@@
 <!-- badges: start -->
-[![R-CMD-check](https://github.com/chainsawriot/rang/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/chainsawriot/rang/actions/workflows/R-CMD-check.yaml)
+[![R-CMD-check](https://github.com/gesistsa/rang/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gesistsa/rang/actions/workflows/R-CMD-check.yaml)
 <!-- badges: end -->
@@
 
-Chan CH, Schoch D (2023) rang: Reconstructing reproducible R computational environments. arXiv preprint:[2303.04758](https://doi.org/10.48550/arXiv.2303.04758)
+Chan CH, Schoch D (2023) rang: Reconstructing reproducible R computational environments. PLOS ONE [https://doi.org/10.1371/journal.pone.0286761](https://doi.org/10.1371/journal.pone.0286761)
 
@@
 ``` r
-remotes::install_github("chainsawriot/rang")
+remotes::install_github("gesistsa/rang")
 ```

### 2023-10-30 paper/gesis_2023/index.qmd
Add gesis slides

--- 2023-07-21 paper/basel2023_rang/index.qmd
@@
 institute: GESIS
-date: "2023-07-21"
+date: "2023-11-02"
 date-format: "YYYY-MM-DD"
-footer: "https://chainsawriot.github.io/basel2023_rang"
+footer: "rang"
 embed-resources: true
@@
 ---
-## Computational Reproducibility
 
-A result is reproducible when the same analysis steps performed on the same dataset **consistently produce the same answer**.
-This reproducibility of the result can be checked by the original investigators and other researchers within a local environment.
-
-::: {style="font-size: 0.3em"}
-Schoch, Chan, Wagner, & Bleier. (2023) Computational Reproducibility in Computational Social Science. [arXiv](https://doi.org/10.48550/arXiv.2307.01918)
-:::
-
-## Case study I: dplyr
-
-```{r}
-#| echo: true
-data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
-data
-```
-
-## Case study I: dplyr
+## Case study
 
@@
 
-## Pit stop 1: Summary
-
-1. Computational reproducibility is tacitly dependent on the computational environment
-2. Recursive dependencies can best be tackled with time-based pinning or lockfile-based system
-3. P3M has its limit (2017-10-10).
-
-::: {.notes}
-P3M also has an inconvenience limitation: What if it's no longer available (like MRAN before it)?
-:::
-
-## Case study II
-
-![](comps.png)
-
-::: {style="font-size: 0.3em"}
-Source: @whydoesr@botsin.space
-:::
-
-## Case study II
-
-Are these analysis steps performed on the same dataset **consistently produce the same answer**?
-
-```r
-unix.time(mean(rnorm(100)))
-n _ 10
-require(sna)
-x _ rgraph(n, mode = "graph")
-evcent(x)
-```
-
-## On my own machine: R 4.3.1
-
-```{r}
-#| echo: true
-#| error: true
-unix.time(mean(rnorm(100)))
-n _ 10
-suppressPackageStartupMessages(require(sna))
-x _ rgraph(n, mode = "graph")
-evcent(x)
-```
-
-Explanations:
-
-```r
-?`base-defunct`
-```
-
-## Non-method #1: P3M/renv
-
-. . .
-
-It's base. You can't pin `unix.time()` with P3M/`renv`.
-
-::: {.notes}
-in Norm Matloff's voice
-:::
-
-## Method #2?: Docker / Rocker
-
-```bash
-# work
-docker run --rm -it rocker/r-ver:3.0.0
-```
-
-. . .
-
-```text
-> unix.time(mean(rnorm(100)))
-   user  system elapsed 
-      0       0       0 
-> n _ 10
-Error: unexpected input in "n _"
-> 
-```
-
-. . .
-
-```bash
-# not work
-docker run --rm -it rocker/r-ver:2.15.3
-```
-
-::: {.notes}
-You probably will have problem installing sna
-:::
-
-## Pit stop 2: Summary
-
-1. Computational reproducibility is tacitly dependent on the computational environment
-2. It's not always about packages, the programming language changes too
-3. Rocker has its limit (R 3.0.0)
-
-::: {.notes}
-(think Python 2 -> 3)
-
-Rocker: Many old images are not usable anymore (either hardcoded MRAN or expiring Debian)
-:::
-
 ## Our solution: rang
-
-![](rang_logo.png){width=300,fig-align="left"}
-
-```r
-install.packages("rang")
-## remotes::install_github("chainsawriot/rang")
-```
-
-::: {style="font-size: 0.3em"}
-Chan & Schoch (2023). rang: Reconstructing reproducible R computational environments. PLOS One [doi](https://doi.org/10.1371/journal.pone.0286761)
-:::
-
-## Features (CRAN version)
 
@@
 
-## README
-
-```text
-This README offers some guidance on how to work with the included docker container 
-which was created with the R package rang on 2023-07-17
-
-# Installing docker
-
-The installation depends on the OS you are running
-...
-```
-
 ## Dockerfile
@@
 
-## Full bash
+# Summary
 
-```bash
-Rscript -e "rang::resolve(\"sna\", snapshot_date = \"2002-02-01\") |> 
-rang::dockerize(\"~/scoms\", cache = TRUE)"
-cd ~/scoms
-docker build -t scoms .
-docker run --rm --name "scoms" -ti scoms
-```
-## For compatibility with S
-
-![](base2.png)
-
-## Many features not demo.
-
-* More examples in [our paper](https://doi.org/10.1371/journal.pone.0286761)
-
-Upcoming features
-
-* Apptainer (née Singularity) - Thanks to the contribution from Egor Kotov
-* Research Compendium
-* Post installation steps, e.g. installing Quarto
-
-## Summary
-
-## Summary
-
-:::: {.columns}
-
-:::{.column width="80%"}
-1. Computational reproducibility is tacitly dependent on the computational environment
-2. Use P3M / renv for projects after 2017-10-10
-    * only at the level of R packages
-3. Plus Rocker for projects newer than R 3.0.0
-4. Use `rang` for probably all projects
-:::
-
-:::{.column width="20%"}
-![](rang_logo.png){width=300}
-:::
-
-::::
-
-::: {style="font-size: 0.3em"}
-Schoch, Chan, Wagner, & Bleier. (2023) Computational Reproducibility in Computational Social Science. [arXiv](https://doi.org/10.48550/arXiv.2307.01918)
-
-Chan & Schoch (2023). rang: Reconstructing reproducible R computational environments. PLOS One [doi](https://doi.org/10.1371/journal.pone.0286761)
-:::
+* code edit
+* Install old version
+* renv
+* P3M
+* rang
+* New development: evercran (CRAN snapshot and versioned Docker images)

### 2023-12-21 README.Rmd
Update README

@@
 
-# rang <img src="man/figures/rang_logo.svg" align="right" height="200" />
+# rang <img src="man/figures/rang_logo.png" align="right" width = "120" />
 
@@
 ```r
-rang_rio <- resolve("rio", snapshot_date = "2013-08-28") ## R 3.0.1
+rang_rio <- resolve("rio", snapshot_date = "2013-08-30") ## R 3.0.1
 dockerize(rang_rio, output_dir = "~/old_renviron")
+```
+
+## `evercran` support (experimental)
+
+`rang` supports [evercran](https://github.com/r-hub/evercran). As of writing, the support is still experimental (just like `evercran` itself). In the future, `evercran` will replace the Debian method.
+
+```r
+rang_rio <- resolve("rio", snapshot_date = "2013-08-30") ## R 3.0.1
+dockerize(rang_rio, output_dir = "~/old_renviron", method = "evercran")
 ```
